---
title: "Antarctic Peninsula - Anions & Index Scores"
author: "Ellie Ryan"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: flatly
    highlight: zenburn
    toc: true
    toc_float: true
    toc_collapsed: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Created:** 9/7/2025

**Last Modified:** 2/18/2026

```{r, message=FALSE, warning=FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(stringr)
library(kableExtra)
library(scales)
library(GGally)
library(car)
library(FSA)
library(Hmisc)
library(patchwork)
library(cowplot)
```

```{r}
plot_theme <- theme(
  text = element_text(family = "Helvetica"),
  axis.title = element_text(size = 12), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  plot.margin = margin(20, 20, 20, 20)
)

plot_themePub <- theme(
  text = element_text(family = "Helvetica"),
  axis.title.x = element_text(size = 12, margin = margin(t = 10)),  # t = top margin
  axis.title.y = element_text(size = 12, margin = margin(r = 10)),  # r = right margin
  axis.text = element_text(size = 12),
  axis.line = element_blank(),
  axis.ticks.x = element_blank(),
  axis.ticks.y = element_line(color = "black", size = 0.25),  # Add tick marks
  axis.ticks.length = unit(-0.15, "cm"),  # Length of tick marks
  panel.grid = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
  legend.position = "top",
  legend.justification = c(0, 1),
  legend.text = element_text(size = 12),
  legend.title = element_text(size = 12),
  plot.title = element_text(size = 14, face = "bold"),
  plot.subtitle = element_text(size = 13, face = "italic"),
  plot.margin = margin(10, 10, 10, 10)
)


plot_theme2 <- theme(
  text = element_text(family = "Helvetica"),
  axis.title = element_text(size = 12, face = "bold"), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  axis.text.x = element_text(angle = 45, hjust = 1), 
  plot.margin = margin(20, 20, 20, 20)
)

pairsplot_theme <- theme(
  text = element_text(size = 10, family = "Helvetica"),  # Adjust variable name text size
  axis.text = element_text(size = 8),  # Adjust axis text size
  plot.title = element_text(size = 12, face = "bold"),  # Adjust plot title size
  legend.text = element_text(size = 12))  # Adjust legend text size

colorPal <- c(
  "clean" = "#769a9c",
  "red" = "#933410",
  "green" = "#326e35"
)

anionPal <- c(
  "PO43" = "#7887A4FF",
  "Cl" = "#9B6981FF",
  "SO42" = "#682C37FF",
  "Fo" = "#F6955EFF",
  "NO2NO3" = "#A8CDECFF"
)

indexPal <- c(
  "WSI" = "#3F3F44",
  "OSI" = "#728187",
  "RDSI" = "#7A5C3E"
)
```

# Overview 
The purpose of this analysis is to assess the relationship between anion concentration and index scores for my 2025 Antarctic Peninsula samples (excluding Petermann Island). Upon sample collection, index values were assigned to each sample location for three different metrics: proximity to rocky outcroppings/dust, proximity to ocean/oceanic influence, and wildlife presence. For each index, a score of 1-4 was assigned, with 4 indicating the highest influence of a given variable. 

From outlier analysis, the following outliers were identified for removal:

-   2025-02-19-SI-algae3
-   2025-02-24-DA-algae2
-   2025-02-24-DA-algae1 (anion only)
-   2025-02-24-DA-algae6
-   2025-03-07-NH-algae3

See the Thesis Figures (Preliminary) doc for justification for removal of each. Likely, more samples will be removed following visualization of pigment distribution. 

## Data Wrangling

```{r}
df <- read.csv("APIndices.csv")
str(df)
```

```{r}
# Format date column 
df <- df %>%
  mutate(
    Date = as.Date(Date, format = "%m/%d/%y"))
```

```{r}
# Add Algae column
df <- df %>% 
   mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  relocate(Algae, .after = Color)

# Convert anion concentrations to µM
df <- df %>% 
  mutate(Fo = 22.21*Fo,
         Cl = 28.21*Cl,
         NO2 = 21.73*NO2,
         NO3 = 16.13*NO3, 
         PO43 = 10.53*PO43, 
         SO42 = 10.41*SO42)
```

```{r}
outliers <- c(
  "2025-02-19-SI-algae3",
  "2025-02-24-DA-algae2",
  "2025-02-24-DA-algae1",
  "2025-02-24-DA-algae6",
  "2025-03-07-NH-algae3")

df <- df %>% 
  filter(!SampleID %in% outliers)
```

### Anions
```{r}
dfAnion <- df %>% 
  select(c("SampleID", "Date", "Location", "Site", "Color", "Algae", Fo:SO42, WildlifeScoringIndex:RocksDustScoringIndex)) %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  mutate(NO2NO3 = NO2 + NO3) %>% 
  relocate(Algae, .after = Color) %>% 
  relocate(NO2NO3, .after = Cl) %>% 
  select(!NO2) %>% 
  select(!NO3) %>% 
  na.omit(dfAnion)

# Abbreviate indices
dfAnion <- dfAnion %>% 
  mutate(WSI = WildlifeScoringIndex) %>% 
  relocate(WSI, .after = RocksDustScoringIndex) %>% 
  mutate(OSI = OceanicInfluenceScoringIndex) %>% 
  relocate(OSI, .after = WSI) %>% 
  mutate(RDSI = RocksDustScoringIndex) %>% 
  relocate(RDSI, .after = OSI)
```

### Anions (Pivoted)
```{r}
dfAnion_long <- dfAnion %>% 
  pivot_longer(cols = c(Fo, Cl, SO42, PO43, NO2NO3),
               names_to = "Anion",
               values_to = "Concentration")
```

### Scoring Index (Pivoted)
```{r}
dfIndex_long <- dfAnion %>% 
  pivot_longer(cols = c(WildlifeScoringIndex, OceanicInfluenceScoringIndex, RocksDustScoringIndex),
               names_to = "Index",
               values_to = "Score")

dfIndex_long <- dfIndex_long %>%
  mutate(Score = factor(Score))

dfIndex_long <- dfIndex_long %>% 
  mutate(IndexAbbrev = 
           case_when(
             Index == "WildlifeScoringIndex" ~ paste("WSI"),
             Index == "OceanicInfluenceScoringIndex" ~ paste("OSI"),
             Index == "RocksDustScoringIndex" ~ paste("RDSI")
           )) %>% 
  relocate(IndexAbbrev, .after = Index)
```


# Summary Tables {.tabset}

## Formate 
```{r}
# Anion Summary (formate)
FoSummary <- dfIndex_long %>% 
  group_by(Index, Score) %>% 
  summarise(
    n = n(),
    Fo_mean = round(mean(Fo, na.rm = TRUE), 2),
    Fo_median = round(median(Fo, na.rm = TRUE), 2),
    Fo_max = round(max(Fo, na.rm = TRUE), 2),
    Fo_min = round(min(Fo, na.rm = TRUE), 2),
    Fo_sd = round(sd(Fo, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Format table with kableExtra
FoSummary %>% 
  kable("html", caption = "Summary Statistics for Formate by Snow Color and Location (ppm)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Chloride 
```{r}
# Anion Summary (Chloride)
ClSummary <- dfIndex_long %>% 
  group_by(Index, Score) %>% 
  summarise(
    n = n(),
    Cl_mean = round(mean(Cl, na.rm = TRUE), 2),
    Cl_median = round(median(Cl, na.rm = TRUE), 2),
    Cl_max = round(max(Cl, na.rm = TRUE), 2),
    Cl_min = round(min(Cl, na.rm = TRUE), 2),
    Cl_sd = round(sd(Cl, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Format table with kableExtra
ClSummary %>% 
  kable("html", caption = "Summary Statistics for Chloride by Snow Color and Location (ppm)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Nitrate/Nitrite
```{r}
# Anion Summary (Nitrate/Nitrite)
NO2NO3Summary <- dfIndex_long %>% 
  group_by(Index, Score) %>% 
  summarise(
    n = n(),
    NO2NO3_mean = round(mean(NO2NO3, na.rm = TRUE), 2),
    NO2NO3_median = round(median(NO2NO3, na.rm = TRUE), 2),
    NO2NO3_max = round(max(NO2NO3, na.rm = TRUE), 2),
    NO2NO3_min = round(min(NO2NO3, na.rm = TRUE), 2),
    NO2NO3_sd = round(sd(NO2NO3, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Format table with kableExtra
NO2NO3Summary %>% 
  kable("html", caption = "Summary Statistics for Nitrate/Nitrite by Snow Color and Location (ppm)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Phosphate 
```{r}
# Anion Summary (Phosphate)
PO43Summary <- dfIndex_long %>% 
  group_by(Index, Score) %>% 
  summarise(
    n = n(),
    PO43_mean = round(mean(PO43, na.rm = TRUE), 2),
    PO43_median = round(median(PO43, na.rm = TRUE), 2),
    PO43_max = round(max(PO43, na.rm = TRUE), 2),
    PO43_min = round(min(PO43, na.rm = TRUE), 2),
    PO43_sd = round(sd(PO43, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Format table with kableExtra
PO43Summary %>% 
  kable("html", caption = "Summary Statistics for Phosphate by Snow Color and Location (ppm)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Sulfate 
```{r}
# Anion Summary (Sulfate)
SO42Summary <- dfIndex_long %>% 
  group_by(Index, Score) %>% 
  summarise(
    n = n(),
    SO42_mean = round(mean(SO42, na.rm = TRUE), 2),
    SO42_median = round(median(SO42, na.rm = TRUE), 2),
    SO42_max = round(max(SO42, na.rm = TRUE), 2),
    SO42_min = round(min(SO42, na.rm = TRUE), 2),
    SO42_sd = round(sd(SO42, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Format table with kableExtra
SO42Summary %>% 
  kable("html", caption = "Summary Statistics for Sulfate by Snow Color and Location (ppm)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

# ANOVA/Kruskal Wallis Test {.tabset}
Before proceeding with an ANOVA test to assess differences in anion concentration between index scores, I am going to check to see if these variables follow the assumptions of normality and equal variance required to run an ANOVA. I will repeat this for each index.

## Wildlife Scoring Index
```{r}
# Step 1: Group by Anion and nest data
nested <- dfAnion_long %>%
  group_by(Anion) %>%
  nest()

# Step 2: Fit ANOVA model, extract residuals, and run tests
anovaAssumptionsResults <- nested %>%
  mutate(
    # Fit ANOVA model
    aov_model = map(data, ~ aov(Concentration ~ WildlifeScoringIndex, data = .x)),
    
    # Shapiro-Wilk test for normality of residuals
    shapiro_p = map_dbl(aov_model, ~ shapiro.test(residuals(.x))$p.value),
    
    # Levene's Test for homogeneity of variance
    levene_p = map_dbl(data, ~leveneTest(Concentration ~ factor(WildlifeScoringIndex), data = .x)$`Pr(>F)`[1])
  )
```

From the below results, it appears that none of the anions are both normally distributed and have equal variances. Because of this, I am going to proceed with a Kruskal-Wallis test in lieu of an ANOVA. 

```{r}
anovaAssumptionsResults %>%
  select(Anion, shapiro_p, levene_p) %>%
  mutate(
    normality_ok = shapiro_p > 0.05,
    homogeneity_ok = levene_p > 0.05
  )
```

## Oceanic Influence Scoring Index
```{r}
# Step 1: Group by Anion and nest data
nested <- dfAnion_long %>%
  group_by(Anion) %>%
  nest()

# Step 2: Fit ANOVA model, extract residuals, and run tests
anovaAssumptionsResults <- nested %>%
  mutate(
    # Fit ANOVA model
    aov_model = map(data, ~ aov(Concentration ~ OceanicInfluenceScoringIndex, data = .x)),
    
    # Shapiro-Wilk test for normality of residuals
    shapiro_p = map_dbl(aov_model, ~ shapiro.test(residuals(.x))$p.value),
    
    # Levene's Test for homogeneity of variance
    levene_p = map_dbl(data, ~leveneTest(Concentration ~ factor(OceanicInfluenceScoringIndex), data = .x)$`Pr(>F)`[1])
  )
```

From the below results, it appears that none of the anions are both normally distributed and have equal variances. Because of this, I am going to proceed with a Kruskal-Wallis test in lieu of an ANOVA. 

```{r}
anovaAssumptionsResults %>%
  select(Anion, shapiro_p, levene_p) %>%
  mutate(
    normality_ok = shapiro_p > 0.05,
    homogeneity_ok = levene_p > 0.05
  )
```

## Rocks/Dust Scoring Index
```{r}
# Step 1: Group by Anion and nest data
nested <- dfAnion_long %>%
  group_by(Anion) %>%
  nest()

# Step 2: Fit ANOVA model, extract residuals, and run tests
anovaAssumptionsResults <- nested %>%
  mutate(
    # Fit ANOVA model
    aov_model = map(data, ~ aov(Concentration ~ RocksDustScoringIndex, data = .x)),
    
    # Shapiro-Wilk test for normality of residuals
    shapiro_p = map_dbl(aov_model, ~ shapiro.test(residuals(.x))$p.value),
    
    # Levene's Test for homogeneity of variance
    levene_p = map_dbl(data, ~leveneTest(Concentration ~ factor(RocksDustScoringIndex), data = .x)$`Pr(>F)`[1])
  )
```

From the below results, it appears that none of the anions are both normally distributed and have equal variances. Because of this, I am going to proceed with a Kruskal-Wallis test in lieu of an ANOVA. 

```{r}
anovaAssumptionsResults %>%
  select(Anion, shapiro_p, levene_p) %>%
  mutate(
    normality_ok = shapiro_p > 0.05,
    homogeneity_ok = levene_p > 0.05
  )
```

## Kruskal-Wallis Test {.tabset}
Below, I use the Kruskal-Wallis test to run a non-parametric comparison of each anion across each score for each index. Then, I follow up with a Dunn's post hoc test for the anions that show significant differences between the index scores. 

### Wildlife Scoring Index
```{r}
kruskalResultsWSI <- dfAnion_long %>%
  group_by(Anion) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ WildlifeScoringIndex, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsWSI <- kruskalResultsWSI %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(WildlifeScoringIndex), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

#### Visualize Results
```{r}
dunn_pairs <- kruskalResultsWSI %>%
  select(Anion, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlotWSI <- ggplot(dunn_pairs, aes(x = Comparison, y = Anion, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunn_pairs, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlotWSI

ggsave(dunnPlotWSI, filename = "dunnPlotWSI.png", width = 7, height = 6, bg = "white")
```

### Oceanic Influence Scoring Index
```{r}
kruskalResultsOSI <- dfAnion_long %>%
  group_by(Anion) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ OceanicInfluenceScoringIndex, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsOSI <- kruskalResultsOSI %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(OceanicInfluenceScoringIndex), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

#### Visualize Results
```{r}
dunn_pairs <- kruskalResultsOSI %>%
  select(Anion, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlotOSI <- ggplot(dunn_pairs, aes(x = Comparison, y = Anion, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunn_pairs, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlotOSI

ggsave(dunnPlotOSI, filename = "dunnPlotOSI.png", width = 7, height = 6, bg = "white")
```

### Rocks/Dust Scoring Index
```{r}
kruskalResultsRDSI <- dfAnion_long %>%
  group_by(Anion) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ RocksDustScoringIndex, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsRDSI <- kruskalResultsRDSI %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(RocksDustScoringIndex), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

#### Visualize Results
```{r}
dunn_pairs <- kruskalResultsRDSI %>%
  select(Anion, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlotRDSI <- ggplot(dunn_pairs, aes(x = Comparison, y = Anion, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunn_pairs, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlotRDSI

ggsave(dunnPlotRDSI, filename = "dunnPlotRDSI.png", width = 7, height = 6, bg = "white")
```

# Box Plots {.tabset}

## Formate
```{r}
boxplotFo <- dfIndex_long %>%
  ggplot(aes(x = Score, y = Fo)) +
  geom_boxplot(alpha = 0.8, width = 0.8, fill = "#3b3f6e", color = "#2e3157") +
  facet_wrap(~ IndexAbbrev, scales = "free_x") +
  labs(
    x = "Score",
    y = "Concentration (ppm)"
  ) +
  theme_minimal() +
  plot_theme +
  theme(panel.spacing.x = unit(0.7, "cm")) 

boxplotFo

ggsave(boxplotFo, filename = "boxplotFo.png", width = 7, height = 5, bg = "white")
```

## Chloride
```{r}
boxplotCl <- dfIndex_long %>%
  ggplot(aes(x = Score, y = Cl)) +
  geom_boxplot(alpha = 0.8, width = 0.8, fill = "#3b3f6e", color = "#2e3157") +
  facet_wrap(~ IndexAbbrev, scales = "free_x") +
  labs(
    x = "Score",
    y = "Concentration (ppm)"
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme +
  theme(panel.spacing.x = unit(0.7, "cm")) 

boxplotCl

ggsave(boxplotCl, filename = "boxplotCl.png", width = 7, height = 5, bg = "white")
```

## Nitrate/Nitrite
```{r}
boxplotNO2NO3 <- dfIndex_long %>%
  ggplot(aes(x = Score, y = NO2NO3)) +
  geom_boxplot(alpha = 0.8, width = 0.8, fill = "#3b3f6e", color = "#2e3157") +
  facet_wrap(~ IndexAbbrev, scales = "free_x") +
  labs(
    x = "Score",
    y = "Concentration (ppm)"
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme +
  theme(panel.spacing.x = unit(0.7, "cm")) 

boxplotNO2NO3

ggsave(boxplotNO2NO3, filename = "boxplotNO2NO3.png", width = 7, height = 5, bg = "white")
```

## Phosphate
```{r}
boxplotPO43 <- dfIndex_long %>%
  ggplot(aes(x = Score, y = PO43)) +
  geom_boxplot(alpha = 0.8, width = 0.8, fill = "#3b3f6e", color = "#2e3157") +
  facet_wrap(~ IndexAbbrev, scales = "free_x") +
  labs(
    x = "Score",
    y = "Concentration (ppm)"
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme +
  theme(panel.spacing.x = unit(0.7, "cm")) 

boxplotPO43

ggsave(boxplotPO43, filename = "boxplotPO43.png", width = 7, height = 5, bg = "white")
```

## Sulfate 
```{r}
boxplotSO42 <- dfIndex_long %>%
  ggplot(aes(x = Score, y = SO42)) +
  geom_boxplot(alpha = 0.8, width = 0.8, fill = "#3b3f6e", color = "#2e3157") +
  facet_wrap(~ IndexAbbrev, scales = "free_x") +
  labs(
    x = "Score",
    y = "Concentration (ppm)"
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme +
  theme(panel.spacing.x = unit(0.7, "cm")) 

boxplotSO42

ggsave(boxplotSO42, filename = "boxplotSO42.png", width = 7, height = 5, bg = "white")
```

## Snow Color and Index Score {.tabset}

### WSI
```{r}
boxplotColorWSI <- dfAnion %>% 
  ggplot(aes(x = WildlifeScoringIndex, fill = Color)) + 
  geom_bar(position = "dodge") + 
  labs( x = "WSI Score", 
        y = "Count",
        color = NULL,
        fill = NULL) + 
  scale_fill_manual(values = colorPal) +
  theme_minimal() +
  plot_theme

boxplotColorWSI

ggsave(boxplotColorWSI, filename = "boxplotColorWSI.png", width = 7, height = 5, bg = "white")
```

### OSI
```{r}
boxplotColorOSI <- dfAnion %>% 
  ggplot(aes(x = OceanicInfluenceScoringIndex, fill = Color)) + 
  geom_bar(position = "dodge") + 
  labs( x = "OSI Score", 
        y = "Count",
        color = NULL,
        fill = NULL) + 
  scale_fill_manual(values = colorPal) +
  theme_minimal() +
  plot_theme

boxplotColorOSI

ggsave(boxplotColorOSI, filename = "boxplotColorOSI.png", width = 7, height = 5, bg = "white")
```

### RDSI
```{r}
boxplotColorRDSI <- dfAnion %>% 
  ggplot(aes(x = RocksDustScoringIndex, fill = Color)) + 
  geom_bar(position = "dodge") + 
  labs( x = "RDSI Score", 
        y = "Count",
        color = NULL,
        fill = NULL) + 
  scale_fill_manual(values = colorPal) +
  theme_minimal() +
  plot_theme

boxplotColorRDSI

ggsave(boxplotColorRDSI, filename = "boxplotColorRDSI.png", width = 7, height = 5, bg = "white")
```

# Heat Maps - Anions & Index Scores
```{r}
# Define pigment and Indexvariables
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")
index_vars <- c("WSI", "OSI", "RDSI")

# Filter to just the necessary columns
df_filtered <- dfAnion %>%
  select(Location, all_of(anion_vars), all_of(index_vars)) %>%
  na.omit()

# Function to compute Anion–IndexSpearman correlations per location
get_Anion_Index_corr <- function(data, location_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")

  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Create all Anion–Indexpair combinations
  pairs <- expand.grid(Anion = anion_vars, Index = index_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Index, Anion],  # rows are Indexs, cols are Anions
      p_value = p_vals[Index, Anion],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      Location = location_name
    ) %>%
    ungroup()

  return(results)
}

# Compute correlations for each location
cor_results <- df_filtered %>%
  group_split(Location) %>%
  map_dfr(~get_Anion_Index_corr(.x, unique(.x$Location)))

# Convert to factors to preserve axis order
cor_results$Anion <- factor(cor_results$Anion, levels = anion_vars)
cor_results$Index <- factor(cor_results$Index, levels = index_vars)

# Plot
heatmapAnionIndex <- ggplot(cor_results, aes(x = Anion, y = Index, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  facet_wrap(~Location, ncol =1) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

heatmapAnionIndex

ggsave(heatmapAnionIndex, filename = "heatmapAnionIndex.png", width = 7, height = 5, bg = "white")
```



# Publication Figures

## Index-Anion Boxplots {.tabset}

### PO43-WSI
```{r}
boxplotPO43_WSI <- dfIndex_long %>%
  filter(IndexAbbrev == "WSI") %>% 
  ggplot(aes(x = Score, y = PO43, fill = IndexAbbrev, color = IndexAbbrev)) +
  geom_boxplot(alpha = 0.9, width = 0.8) +
  scale_fill_manual(values = indexPal) +
  scale_color_manual(values = indexPal) +
  labs(
    x = NULL,
    y = expression("[ "*PO[4]^{"3-"}*" ]" ~ "(µM)"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 2, 20, 50, 100, 200, 400),
               labels = function(x) ifelse(x < 1, x, as.integer(x))) +
  theme_minimal() +
  plot_themePub +
  theme(legend.position = "none")+
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    )
  )

boxplotPO43_WSI
```


### Cl-OSI & RDSI
```{r}
boxplotCl_OSI_RDSI <- dfIndex_long %>%
  filter(IndexAbbrev %in% c("OSI", "RDSI")) %>% 
  ggplot(aes(x = Score, y = Cl, fill = IndexAbbrev, color = IndexAbbrev)) +
  geom_boxplot(alpha = 0.9, width = 0.8) +
  scale_fill_manual(values = indexPal) +
  scale_color_manual(values = indexPal) +
  labs(
    x = NULL,
    y = expression("[ "*Cl^{"-"}*" ]" ~ "(µM)"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 10, 50, 100, 300, 600, 900),
               labels = function(x) ifelse(x < 1, x, as.integer(x))) +
  theme_minimal() +
  plot_themePub +
  theme(legend.position = "none")+
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    )
  )

boxplotCl_OSI_RDSI
```

### Fo-OSI
```{r}
boxplotFo_OSI <- dfIndex_long %>%
  filter(IndexAbbrev == "OSI") %>% 
  ggplot(aes(x = Score, y = Fo, fill = IndexAbbrev, color = IndexAbbrev)) +
  geom_boxplot(alpha = 0.9, width = 0.8) +
  scale_fill_manual(values = indexPal) +
  scale_color_manual(values = indexPal) +
  labs(
    x = NULL,
    y = expression("[ "*HCOO^{"-"}*" ]" ~ "(µM)"),
    fill = NULL,
    color = NULL) +
  theme_minimal() +
  plot_themePub +
  theme(legend.position = "none")+
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    )
  )

boxplotFo_OSI
```

### SO42-RDSI
```{r}
boxplotSO42_RDSI <- dfIndex_long %>%
  filter(IndexAbbrev == "RDSI") %>% 
  ggplot(aes(x = Score, y = SO42, fill = IndexAbbrev, color = IndexAbbrev)) +
  geom_boxplot(alpha = 0.9, width = 0.8) +
  scale_fill_manual(values = indexPal) +
  scale_color_manual(values = indexPal) +
  labs(
    x = NULL,
    y = expression("[ "*SO[4]^{"2-"}*" ]" ~ "(µM)"),
    fill = NULL,
    color = NULL) +
  scale_y_continuous(limits = c(0, NA)) +  
  theme_minimal() +
  plot_themePub +
  theme(legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    )
  )

boxplotSO42_RDSI
```

### Shared Legend
```{r}
legendPlot <- dfIndex_long %>%
  ggplot(aes(x = Score, y = SO42, fill = IndexAbbrev, color = IndexAbbrev)) +
  geom_boxplot(alpha = 0.9, width = 0.8) +
  scale_fill_manual(values = indexPal) +
  scale_color_manual(values = indexPal) +
  labs(
    x = "Index Score",
    y = expression("[ "*SO[4]^{"2-"}*" ]" ~ "(µM)"),
    fill = NULL,
    color = NULL) +
  scale_y_continuous(limits = c(0, NA)) +  
  theme_minimal() +
  plot_themePub +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    )) + 
  theme(
    legend.position = "bottom",
    legend.justification = "center",  # center the legend
    legend.box.just = "center",       # center the box
    text = element_text(family = "Helvetica")
  )
legendPlot

```

### Combined Plot

```{r}
# Extract legend
sharedLegend <- cowplot::get_legend(legendPlot)

# Arrange 4 plots in 2x2 with spacing controlled
grid_plot <- cowplot::plot_grid(
  boxplotPO43_WSI, boxplotCl_OSI_RDSI,
  boxplotFo_OSI, boxplotSO42_RDSI,
  ncol = 2, nrow = 2,
  rel_heights = c(1,1),
  rel_widths  = c(1,1),
  align = "hv",       # align horizontally and vertically
  axis = "tblr",      # align all axes
  labels = NULL
)

combinedPlot_AnionIndex <- cowplot::plot_grid(
  grid_plot, sharedLegend,
  ncol = 1,
  rel_heights = c(1, 0.08) # adjust vertical gap between grid and legend
)

combinedPlot_AnionIndex

ggsave(combinedPlot_AnionIndex, filename = "combinedPlot_AnionIndex.png", width = 6, height = 5.5, bg = "white")
```

