---
title: "Data Vizualization - Pigment Composition"
author: "Ellie Ryan"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: flatly
    highlight: zenburn
    toc: true
    toc_float: true
    toc_collapsed: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Created:** 7/21/2025

**Last Modified:** 1/23/2026

```{r, message=FALSE, warning=FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(stringr)
library(kableExtra)
library(scales)
library(GGally)
library(patchwork)
library(Hmisc)
library(reshape2)
library(vegan)
library(car)
library(FSA)
library(paletteer)
```

```{r}
regionPal <- c("North" = "#284C46FF",
               "South" = "#b04a25")
colorPal <- c(
  "clean" = "#769a9c",
  "red" = "#933410",
  "green" = "#326e35"
)

algaePal <- c(
  "clean" = "#a5bdc0",
  "algae" = "#735a69"
)

pigPal <- c(
  "Astaxanthin" = "#b04a25",
  "Lutein" = "#d9983b",
  "ChlA" = "#284C46FF",
  "ChlB" = "#647d6a",
  "BetaC" = "#4d1d05",
  "Pheophorbide" = "#414c66", 
  "Pheophytin" = "#95879f"
)

plot_theme <- theme(
  axis.title = element_text(size = 12, face = "bold"), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  plot.margin = margin(20, 20, 20, 20)
)

plot_themePub <- theme(
  text = element_text(family = "Helvetica"),
  axis.title.x = element_text(size = 10, margin = margin(t = 10)),  # t = top margin
  axis.title.y = element_text(size = 10, margin = margin(r = 10)),  # r = right margin
  axis.text = element_text(size = 10),
  axis.line.x = element_line(color = "black", linewidth = 0.5),
  axis.line.y = element_line(color = "black", linewidth = 0.5),,
  axis.ticks.x = element_blank(),
  axis.ticks.y = element_line(color = "black", linewidth = 0.25),  # Add tick marks
  axis.ticks.length = unit(-0.15, "cm"),  # Length of tick marks
  panel.grid = element_blank(),
  panel.background = element_blank(),
  panel.border = element_blank(),
  legend.position = "top",
  legend.justification = c(0, 1),
  legend.text = element_text(size = 10),
  legend.title = element_text(size = 10),
  plot.title = element_text(size = 14, face = "bold"),
  plot.subtitle = element_text(size = 13, face = "italic"),
  plot.margin = margin(10, 10, 10, 10)
)

plot_theme2 <- theme(
  axis.title = element_text(size = 12, face = "bold"), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  axis.text.x = element_text(angle = 45, hjust = 1), 
  plot.margin = margin(20, 20, 20, 20)
)

pairsplot_theme <- theme(
  text = element_text(size = 10),  # Adjust variable name text size
  axis.text = element_text(size = 8),  # Adjust axis text size
  plot.title = element_text(size = 12, face = "bold"),  # Adjust plot title size
  legend.text = element_text(size = 12))  # Adjust legend text size
```


# Overview

The purpose of this `.Rmd` is to visualize pigment composition (%) by color, and obtain basic summary statistics associated with each pigment in order to inform my next steps in analysis. I will additionally perform an ordination to reduce the dimensions of my pigment data for simpler visualizations. Before proceeding, I will clean the dataset so that it features only the variables I am interested in for this analysis, and calculate pigment composition by dividing the concentration of each pigment by the total pigment in each sample. In this dataset, 108 samples are used from the Antarctic Peninsula, following the removal of 5 outliers. These observations are broken down by `Color`, or the apparent snow color at time of sample collection. To keep things simple, there are only three levels to the `Color` category: red, green and clean. The breakdown of samples collected in the WAP is as follows:

-   Antarctic Peninsula:
    -   107 samples
        -   43 red
        -   46 green
        -   18 clean
        
*NOTE: Fewer samples were used for analysis, since there are NA values for pigment concentrations of some samples.*
        
From anion outlier analysis, the following outliers were identified for removal:

-   2025-02-19-SI-algae3
-   2025-02-24-DA-algae2
-   2025-02-24-DA-algae6
-   2025-03-07-NH-algae3
-   2025-01-12-PI-algae10

See the Preliminary Results and Figures doc for justification for removal of each. 

## Data Wrangling

```{r}
df <- read.csv("CompiledDataPI.csv")
str(df)
```

Adding an `Algae` column to classify samples as algae or clean snow samples and a `LocColor` column to classify AP algae samples as red or green. 

```{r}
# Format date column 
df <- df %>%
  mutate(
    Date = as.Date(Date, format = "%m/%d/%y"))

# Add Algae
df <- df %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  relocate(Algae, .after = Color)

# Add Region
df <- df %>% 
  mutate(
  Region = if_else(
    Site %in% c("Horseshoe Island", "Stonington Island"),
    "South",
    "North")) %>% 
  relocate(Region, .after = Site)

# Add LocAbbrev
df <- df %>% mutate(LocAbbrev = case_when(
  Location == "Antarctic Peninsula" ~ paste("WAP")
)) %>%  
  relocate(LocAbbrev, .after = Site)

# Add TotPig 
df <- df %>% mutate(TotPig = Astaxanthin + Lutein + BetaC + ChlA + ChlB + Pheophorbide + Pheophytin) %>% 
  relocate(TotPig, .after = Pheophytin)
```

And obtain percent pigment concentration by dividing the concentration of each pigment by `TotPig` and multiplying by 100. 

```{r}
dfComp <- df %>% 
  mutate(ChlA = (ChlA/TotPig)*100,
         ChlB = (ChlB/TotPig)*100,
         Astaxanthin = (Astaxanthin/TotPig)*100,
         Lutein = (Lutein/TotPig)*100,
         BetaC = (BetaC/TotPig)*100,
         Pheophytin = (Pheophytin/TotPig)*100,
         Pheophorbide = (Pheophorbide/TotPig)*100) 
```

```{r}
str(dfComp)
```

```{r}
outliers <- c(
  "2025-02-19-SI-algae3",
  "2025-02-24-DA-algae2",
  "2025-02-24-DA-algae6",
  "2025-03-07-NH-algae3",
  "2025-01-12-PI-algae10"
  )

dfComp <- dfComp %>% 
  filter(!SampleID %in% outliers)
```

### Pigments
```{r}
dfPig <- dfComp %>% 
  select(c("SampleID", "Date", "Location", "Site", "Region", "LocAbbrev", "Color", "Algae", Astaxanthin:Pheophytin)) %>% 
  na.omit(dfPig)
```

### Pigments/Cell Density
```{r}
dfPigCell <- df %>% 
  select(c("SampleID", "Date", "Location", "Site", "Region", "LocAbbrev", "Color", "Algae", "CellDensity", "CellLength", "TotPig", Astaxanthin:Pheophytin)) %>% 
  mutate(ChlA = (ChlA/TotPig)*100,
         ChlB = (ChlB/TotPig)*100,
         Astaxanthin = (Astaxanthin/TotPig)*100,
         Lutein = (Lutein/TotPig)*100,
         BetaC = (BetaC/TotPig)*100,
         Pheophytin = (Pheophytin/TotPig)*100,
         Pheophorbide = (Pheophorbide/TotPig)*100) %>% 
  na.omit(dfPigCell) %>% 
    filter(!SampleID %in% outliers) %>% 
    mutate(
    # Calulate volume (µm^3)
    CellVolume = (pi/6)*(CellLength^3)) %>% 
  relocate(CellVolume, .after = CellLength)
```

### Pigments & Anions 
```{r}
dfAnion <- dfComp %>% 
  select(c("SampleID", "Date", "Location", "Site", "Region", "LocAbbrev", "Color", "Algae", Astaxanthin:Pheophytin, Fo:SO42)) %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  mutate(NO2NO3 = NO2 + NO3) %>% 
  relocate(Algae, .after = Color) %>% 
  relocate(NO2NO3, .after = Cl) %>% 
  select(!NO2) %>% 
  select(!NO3) %>% 
  na.omit(dfAnion)
```

### Pigments (Pivoted)
```{r}
dfPig_long <- dfPig %>% 
  pivot_longer(cols = Astaxanthin:Pheophytin,
               names_to = "Pigment",
               values_to = "Composition")
```

### Anions (Pivoted)
```{r}
dfAnion_long <- dfAnion %>% 
  pivot_longer(cols = c(Fo, Cl, SO42, PO43, NO2NO3),
               names_to = "Anion",
               values_to = "Concentration")
```

# ANOVA/Kruskal Wallis Test
Before proceeding with an ANOVA test to assess differences in pigment Composition between colors, I am going to check to see if these variables follow the assumptions of normality and equal variance required to run an ANOVA: 
```{r}
# Step 1: Group by pigment and nest data
nested <- dfPig_long %>%
  group_by(Pigment) %>%
  nest()

# Step 2: Fit ANOVA model, extract residuals, and run tests
anovaAssumptionsResults <- nested %>%
  mutate(
    # Fit ANOVA model
    aov_model = map(data, ~ aov(Composition ~ Color, data = .x)),
    
    # Shapiro-Wilk test for normality of residuals
    shapiro_p = map_dbl(aov_model, ~ shapiro.test(residuals(.x))$p.value),
    
    # Levene's Test for homogeneity of variance
    levene_p = map_dbl(data, ~leveneTest(Composition ~ factor(Color), data = .x)$`Pr(>F)`[1])
  )
```

From the below results, it appears that the majority of the pigments are neither normally distributed or have equal variances. Because of this, I am going to proceed with a Kruskal-Wallis test in lieu of an ANOVA. 

```{r}
anovaAssumptionsResults %>%
  select(Pigment, shapiro_p, levene_p) %>%
  mutate(
    normality_ok = shapiro_p > 0.05,
    homogeneity_ok = levene_p > 0.05
  )
```

## Kruskal-Wallis Test 
Below, I use the Kruskal-Wallis test to run a non-parametric comparison of each pigment across each snow color. Then, I follow up with a Dunn's post hoc test for the pigments that show significant differences between snow colors. 

```{r}
# Run kruskal wallis test 
kruskalResultsAP <- dfPig_long %>%
  filter(Location == "Antarctic Peninsula") %>% 
  group_by(Pigment) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Composition ~ Color, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsAP <- kruskalResultsAP %>%
  mutate(
    dunn = map(data, ~ dunnTest(Composition ~ factor(Color), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

#### Visualize Results
```{r}
dunnPairsAP <- kruskalResultsAP %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_AP <- ggplot(dunnPairsAP, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsAP, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_AP

ggsave(dunnPlot_AP, filename = "dunnPlot_AP.png", width = 7, height = 6, bg = "white")
```

# Summary Tables {.tabset}

## Primary Pigments
```{r}
# Pigment Summary (ChlA and ChlB)
ChlSummary <- dfPig %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    ChlA_mean = round(mean(ChlA, na.rm = TRUE), 2),
    ChlA_median = round(median(ChlA, na.rm = TRUE), 2),
    ChlA_max = round(max(ChlA, na.rm = TRUE), 2),
    ChlA_min = round(min(ChlA, na.rm = TRUE), 2),
    ChlA_sd = round(sd(ChlA, na.rm = TRUE), 2),
    ChlB_mean = round(mean(ChlB, na.rm = TRUE), 2),
    ChlB_median = round(median(ChlB, na.rm = TRUE), 2),
    ChlB_max = round(max(ChlB, na.rm = TRUE), 2),
    ChlB_min = round(min(ChlB, na.rm = TRUE), 2),
    ChlB_sd = round(sd(ChlB, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Add "all" for each site
ChlSummaryAll <- dfPig %>% 
  summarise(
    Color = "all",
    n = n(),
    ChlA_mean = round(mean(ChlA, na.rm = TRUE), 2),
    ChlA_median = round(median(ChlA, na.rm = TRUE), 2),
    ChlA_max = round(max(ChlA, na.rm = TRUE), 2),
    ChlA_min = round(min(ChlA, na.rm = TRUE), 2),
    ChlA_sd = round(sd(ChlA, na.rm = TRUE), 2),
    ChlB_mean = round(mean(ChlB, na.rm = TRUE), 2),
    ChlB_median = round(median(ChlB, na.rm = TRUE), 2),
    ChlB_max = round(max(ChlB, na.rm = TRUE), 2),
    ChlB_min = round(min(ChlB, na.rm = TRUE), 2),
    ChlB_sd = round(sd(ChlB, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Filter Antarctic red + green
ChlSummaryAPRG <- dfPig %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Color = "red+green",
    n = n(),
    ChlA_mean = round(mean(ChlA, na.rm = TRUE), 2),
    ChlA_median = round(median(ChlA, na.rm = TRUE), 2),
    ChlA_max = round(max(ChlA, na.rm = TRUE), 2),
    ChlA_min = round(min(ChlA, na.rm = TRUE), 2),
    ChlA_sd = round(sd(ChlA, na.rm = TRUE), 2),
    ChlB_mean = round(mean(ChlB, na.rm = TRUE), 2),
    ChlB_median = round(median(ChlB, na.rm = TRUE), 2),
    ChlB_max = round(max(ChlB, na.rm = TRUE), 2),
    ChlB_min = round(min(ChlB, na.rm = TRUE), 2),
    ChlB_sd = round(sd(ChlB, na.rm = TRUE), 2)
  )

# Combine and order rows
ChlSummary <- bind_rows(ChlSummaryAll, ChlSummary, ChlSummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
ChlSummary %>% 
  kable("html", caption = "Summary Statistics for ChlA and ChlB by Snow Color (ng/cell)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Carotenoids
```{r}
# Pigment Summary (Astaxanthin, Beta Carotene and Lutein)
CarotenoidSummary <- dfPig %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    Astaxanthin_mean = round(mean(Astaxanthin, na.rm = TRUE), 2),
    Astaxanthin_median = round(median(Astaxanthin, na.rm = TRUE), 2),
    Astaxanthin_max = round(max(Astaxanthin, na.rm = TRUE), 2),
    Astaxanthin_min = round(min(Astaxanthin, na.rm = TRUE), 2),
    Astaxanthin_sd = round(sd(Astaxanthin, na.rm = TRUE), 2),
    Lutein_mean = round(mean(Lutein, na.rm = TRUE), 2),
    Lutein_median = round(median(Lutein, na.rm = TRUE), 2),
    Lutein_max = round(max(Lutein, na.rm = TRUE), 2),
    Lutein_min = round(min(Lutein, na.rm = TRUE), 2),
    Lutein_sd = round(sd(Lutein, na.rm = TRUE), 2),
    BetaC_mean = round(mean(BetaC, na.rm = TRUE), 2),
    BetaC_median = round(median(BetaC, na.rm = TRUE), 2),
    BetaC_max = round(max(BetaC, na.rm = TRUE), 2),
    BetaC_min = round(min(BetaC, na.rm = TRUE), 2),
    BetaC_sd = round(sd(BetaC, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Add "all" for each site
CarotenoidSummaryAll <- dfPig %>% 
  summarise(
    Color = "all",
    n = n(),
    Astaxanthin_mean = round(mean(Astaxanthin, na.rm = TRUE), 2),
    Astaxanthin_median = round(median(Astaxanthin, na.rm = TRUE), 2),
    Astaxanthin_max = round(max(Astaxanthin, na.rm = TRUE), 2),
    Astaxanthin_min = round(min(Astaxanthin, na.rm = TRUE), 2),
    Astaxanthin_sd = round(sd(Astaxanthin, na.rm = TRUE), 2),
    Lutein_mean = round(mean(Lutein, na.rm = TRUE), 2),
    Lutein_median = round(median(Lutein, na.rm = TRUE), 2),
    Lutein_max = round(max(Lutein, na.rm = TRUE), 2),
    Lutein_min = round(min(Lutein, na.rm = TRUE), 2),
    Lutein_sd = round(sd(Lutein, na.rm = TRUE), 2),
    BetaC_mean = round(mean(BetaC, na.rm = TRUE), 2),
    BetaC_median = round(median(BetaC, na.rm = TRUE), 2),
    BetaC_max = round(max(BetaC, na.rm = TRUE), 2),
    BetaC_min = round(min(BetaC, na.rm = TRUE), 2),
    BetaC_sd = round(sd(BetaC, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Filter Antarctic red + green
CarotenoidSummaryAPRG <- dfPig %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Color = "red+green",
    n = n(),
    Astaxanthin_mean = round(mean(Astaxanthin, na.rm = TRUE), 2),
    Astaxanthin_median = round(median(Astaxanthin, na.rm = TRUE), 2),
    Astaxanthin_max = round(max(Astaxanthin, na.rm = TRUE), 2),
    Astaxanthin_min = round(min(Astaxanthin, na.rm = TRUE), 2),
    Astaxanthin_sd = round(sd(Astaxanthin, na.rm = TRUE), 2),
    Lutein_mean = round(mean(Lutein, na.rm = TRUE), 2),
    Lutein_median = round(median(Lutein, na.rm = TRUE), 2),
    Lutein_max = round(max(Lutein, na.rm = TRUE), 2),
    Lutein_min = round(min(Lutein, na.rm = TRUE), 2),
    Lutein_sd = round(sd(Lutein, na.rm = TRUE), 2),
    BetaC_mean = round(mean(BetaC, na.rm = TRUE), 2),
    BetaC_median = round(median(BetaC, na.rm = TRUE), 2),
    BetaC_max = round(max(BetaC, na.rm = TRUE), 2),
    BetaC_min = round(min(BetaC, na.rm = TRUE), 2),
    BetaC_sd = round(sd(BetaC, na.rm = TRUE), 2)
  )

# Combine and order rows
CarotenoidSummary <- bind_rows(CarotenoidSummaryAll, CarotenoidSummary, CarotenoidSummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
CarotenoidSummary %>% 
  kable("html", caption = "Summary Statistics for Astaxanthin and Lutein by Snow Color (ng/cell)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```


## Degradation Products 
```{r}
# Pigment Summary (Pheophytin and Pheophorbide)
DegSummary <- dfPig %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    Pheophytin_mean = round(mean(Pheophytin, na.rm = TRUE), 2),
    Pheophytin_median = round(median(Pheophytin, na.rm = TRUE), 2),
    Pheophytin_max = round(max(Pheophytin, na.rm = TRUE), 2),
    Pheophytin_min = round(min(Pheophytin, na.rm = TRUE), 2),
    Pheophytin_sd = round(sd(Pheophytin, na.rm = TRUE), 2),
    Pheophorbide_mean = round(mean(Pheophorbide, na.rm = TRUE), 2),
    Pheophorbide_median = round(median(Pheophorbide, na.rm = TRUE), 2),
    Pheophorbide_max = round(max(Pheophorbide, na.rm = TRUE), 2),
    Pheophorbide_min = round(min(Pheophorbide, na.rm = TRUE), 2),
    Pheophorbide_sd = round(sd(Pheophorbide, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Add "all" for each site
DegSummaryAll <- dfPig %>% 
  summarise(
    Color = "all",
    n = n(),
    Pheophytin_mean = round(mean(Pheophytin, na.rm = TRUE), 2),
    Pheophytin_median = round(median(Pheophytin, na.rm = TRUE), 2),
    Pheophytin_max = round(max(Pheophytin, na.rm = TRUE), 2),
    Pheophytin_min = round(min(Pheophytin, na.rm = TRUE), 2),
    Pheophytin_sd = round(sd(Pheophytin, na.rm = TRUE), 2),
    Pheophorbide_mean = round(mean(Pheophorbide, na.rm = TRUE), 2),
    Pheophorbide_median = round(median(Pheophorbide, na.rm = TRUE), 2),
    Pheophorbide_max = round(max(Pheophorbide, na.rm = TRUE), 2),
    Pheophorbide_min = round(min(Pheophorbide, na.rm = TRUE), 2),
    Pheophorbide_sd = round(sd(Pheophorbide, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Filter Antarctic red + green
DegSummaryAPRG <- dfPig %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Location = "Antarctic Peninsula",
    Color = "red+green",
    n = n(),
    Pheophytin_mean = round(mean(Pheophytin, na.rm = TRUE), 2),
    Pheophytin_median = round(median(Pheophytin, na.rm = TRUE), 2),
    Pheophytin_max = round(max(Pheophytin, na.rm = TRUE), 2),
    Pheophytin_min = round(min(Pheophytin, na.rm = TRUE), 2),
    Pheophytin_sd = round(sd(Pheophytin, na.rm = TRUE), 2),
    Pheophorbide_mean = round(mean(Pheophorbide, na.rm = TRUE), 2),
    Pheophorbide_median = round(median(Pheophorbide, na.rm = TRUE), 2),
    Pheophorbide_max = round(max(Pheophorbide, na.rm = TRUE), 2),
    Pheophorbide_min = round(min(Pheophorbide, na.rm = TRUE), 2),
    Pheophorbide_sd = round(sd(Pheophorbide, na.rm = TRUE), 2)
  )

# Combine and order rows
DegSummary <- bind_rows(DegSummaryAll, DegSummary, DegSummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
DegSummary %>% 
  kable("html", caption = "Summary Statistics for Pheophytin and Pheophorbide by Snow Color (ng/cell)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

# Box Plots {.tabset}

## All Pigments (All Samples)
```{r}
boxplotAll <- dfPig_long %>%
  ggplot(aes(x = Pigment, y = Composition, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Composition (%)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme2

boxplotAll

ggsave(boxplotAll, filename = "boxplotAll.png", width = 7, height = 5, bg = "white")
```

## All Pigments (Algae Samples)
```{r}
boxplotAlgae <- dfPig_long %>%
  filter(Algae == "algae") %>% 
  ggplot(aes(x = Pigment, y = Composition, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Composition (%)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme2

boxplotAlgae

ggsave(boxplotAlgae, filename = "boxplotAlgae.png", width = 7, height = 5, bg = "white")
```


## Astaxanthin 
```{r}
boxplotAstax <- dfPig %>%
  ggplot(aes(x = Region, y = Astaxanthin, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Composition (%)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme

boxplotAstax

ggsave(boxplotAstax, filename = "boxplotAstax.png", width = 7, height = 5, bg = "white")
```

## Beta Carotene
```{r}
boxplotBetaC <- dfPig %>%
  ggplot(aes(x = Region, y = BetaC, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Composition (%)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme

boxplotBetaC

ggsave(boxplotBetaC, filename = "boxplotBetaC.png", width = 7, height = 5, bg = "white")
```

## ChlA
```{r}
boxplotChlA <- dfPig %>%
  ggplot(aes(x = Region, y = ChlA, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Composition (%)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme

boxplotChlA

ggsave(boxplotChlA, filename = "boxplotChlA.png", width = 7, height = 5, bg = "white")
```

## ChlB
```{r}
boxplotChlB <- dfPig %>%
  ggplot(aes(x = Region, y = ChlB, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Composition (%)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme

boxplotChlB

ggsave(boxplotChlB, filename = "boxplotChlB.png", width = 7, height = 5, bg = "white")
```

## Lutein
```{r}
boxplotLutein <- dfPig %>%
  ggplot(aes(x = Region, y = Lutein, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Composition (%)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme

boxplotLutein

ggsave(boxplotLutein, filename = "boxplotLutein.png", width = 7, height = 5, bg = "white")
```

## Pheophorbide 
```{r}
boxplotPheophorbide <- dfPig %>%
  ggplot(aes(x = Region, y = Pheophorbide, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Composition (%)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme

boxplotPheophorbide

ggsave(boxplotPheophorbide, filename = "boxplotPheophorbide.png", width = 7, height = 5, bg = "white")
```

## Pheophytin
```{r}
boxplotPheophytin <- dfPig %>%
  ggplot(aes(x = Region, y = Pheophytin, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Composition (%)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme

boxplotPheophytin

ggsave(boxplotPheophytin, filename = "boxplotPheophytin.png", width = 7, height = 5, bg = "white")
```

# Pairs Plots - Pigments

```{r, warning = FALSE, message = FALSE}
corplotAll <- ggpairs(
  dfPig,  
  columns = c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE, method = "spearman"))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAll

ggsave(corplotAll, filename = "corplotAll.png", width = 7, height = 7, bg = "white")
```


# Heat Maps - Pigments {.tabset}

## All Sites (All Samples)
```{r}
cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
pigment_vars <- c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin")

# List of locations
locations <- unique(dfPig$LocAbbrev)

# Function to calculate correlation matrices for each location
get_corr_data <- function(location) {
  data_sub <- dfPig %>%
    filter(LocAbbrev == location) %>%
    select(all_of(pigment_vars)) %>%
    na.omit()

  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p

  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)

  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Location <- location

  return(merged)
}

# Apply across locations
heatmap_data <- do.call(rbind, lapply(locations, get_corr_data))

# Factor levels for correct ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = pigment_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = pigment_vars)

# Create label
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Remove upper triangle + diagonal if desired (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot
heatmapPigAll <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Location) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank()) 

heatmapPigAll

ggsave(heatmapPigAll, filename = "heatmapPigAll.png", width = 11.5, height = 7, bg = "white")

```

## All Sites (Algae Samples)
```{r}
cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
pigment_vars <- c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin")

# List of locations
locations <- unique(dfPig$LocAbbrev)

# Function to calculate correlation matrices for each location
get_corr_data <- function(location) {
  data_sub <- dfPig %>%
    filter(Algae == "algae") %>% 
    filter(LocAbbrev == location) %>%
    select(all_of(pigment_vars)) %>%
    na.omit()

  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p

  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)

  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Location <- location

  return(merged)
}

# Apply across locations
heatmap_data <- do.call(rbind, lapply(locations, get_corr_data))

# Factor levels for correct ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = pigment_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = pigment_vars)

# Create label
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Remove upper triangle + diagonal if desired (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot
heatmapPigAlgae <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Location) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank()) 

heatmapPigAlgae

ggsave(heatmapPigAlgae, filename = "heatmapPigAlgae.png", width = 11.5, height = 7, bg = "white")

```

## All Sites (Clean Samples)
```{r}
cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
pigment_vars <- c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin")

# List of locations
locations <- unique(dfPig$LocAbbrev)

# Function to calculate correlation matrices for each location
get_corr_data <- function(location) {
  data_sub <- dfPig %>%
    filter(Algae == "clean") %>% 
    filter(LocAbbrev == location) %>%
    select(all_of(pigment_vars)) %>%
    na.omit()

  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p

  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)

  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Location <- location

  return(merged)
}

# Apply across locations
heatmap_data <- do.call(rbind, lapply(locations, get_corr_data))

# Factor levels for correct ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = pigment_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = pigment_vars)

# Create label
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Remove upper triangle + diagonal if desired (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot
heatmapPigClean <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Location) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank()) 

heatmapPigClean

ggsave(heatmapPigClean, filename = "heatmapPigClean.png", width = 11.5, height = 7, bg = "white")

```

## Antarctic Peninsula (Red vs. Green)
```{r}

cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
pigment_vars <- c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin")

# Filter data once for Antarctic Peninsula red and green samples
df_filtered <- dfPig %>%
  filter(Location == "Antarctic Peninsula",
         Color %in% c("red", "green"))

# Unique colors to iterate over
colors <- unique(df_filtered$Color)

# Function to calculate correlation matrices for each color group
get_corr_data <- function(color_group) {
  data_sub <- df_filtered %>%
    filter(Color == color_group) %>%
    select(all_of(pigment_vars)) %>%
    na.omit()
  
  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p
  
  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)
  
  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Color <- color_group
  
  return(merged)
}

# Get correlation data for each color
heatmap_data_list <- lapply(colors, get_corr_data)
heatmap_data <- do.call(rbind, heatmap_data_list)

# Set factor levels for ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = pigment_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = pigment_vars)

# Create labels for plot
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Keep only lower triangle (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot faceted by Color
heatmapPigAPRG <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Color) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank())
  
heatmapPigAPRG
ggsave(heatmapPigAPRG, filename = "heatmapPigAPRG.png", width = 9, height = 5, bg = "white")
```

# Principal Components Analysis - Pigments
From the above plots, it seems as though pigment data are relatively linear. I am going to proceed with a PCA ordination over a nonparametric ordination, like nMDS. 

## Standardize Data & Run PCA
Below, I use prcomp to standardize pigment data and run a PCA. Even though the pigment compositions are on the same scale, some pigments are present in much higher percent compositions than others. Because of this, I will standardize pigments so they are all on the same scale.

PC1 explains 43.92% of the variance in pigment composition. PC2 explains 20.13% of the variance. Collectively, these two explain 64.05% of the variance. Ideally, I would use the amount of PCs that bring cumulative variance up to 80% (4 in this case), but I am going to proceed with using the first two PCs so as not to overcomplicate the interpretation. From the scree plot below, it looks like the plot elbow is at PC2, although not very distinct. 

```{r}
Pigments <- dfPig[9:15]

# Standardize
PigPCA <- prcomp(Pigments, scale = TRUE)

summary(PigPCA)
```


```{r}
# Calculate the proportion of variance explained by each eigenvector
foo <- eigenvals(PigPCA)/sum(eigenvals(PigPCA))
# Screeplot
    plot(foo, type = 'l',
         ylab = 'proportion of variance explained',
         xlab = 'principal components')
    points(foo, pch = 21, bg = 'white')
```

## Biplots

```{r}
biplot(PigPCA)
```

## PC Interpretation and Exploration 

```{r}
PigPCA$rotation[,1:3]
```

## Ordination Plots {.tabset}
```{r}
## Add PC1, 2 &3 to dataset
dfPig$PC1 <- PigPCA$x[,1]
dfPig$PC2 <- PigPCA$x[,2]
dfPig$PC3 <- PigPCA$x[,3]
```

### Region 
```{r}
ordplotPigRegion <- dfPig %>% 
  ggplot(aes(x = PC1, y = PC2, color = Region, fill = Region)) + 
  geom_point(size = 2.5, alpha = 0.7, shape = 21) + 
  scale_fill_manual(values = regionPal) +
  scale_color_manual(values = regionPal) +
  labs(x = "PC1 (43.92%)", 
       y = "PC2 (20.13%)", 
       fill = NULL,
       color = NULL) + 
  theme_minimal() + 
  
  plot_theme

ordplotPigRegion

ggsave(ordplotPigRegion, filename = "ordplotPigRegion.png", width = 7, height = 5, bg = "white")
```

### Color 
```{r}
ordplotPigColor <- dfPig %>% 
  ggplot(aes(x = PC1, y = PC2, color = Color, fill = Color)) + 
  geom_point(size = 2.5, alpha = 0.7, shape = 21) + 
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(x = "PC1 (43.92%)", 
       y = "PC2 (20.13%)", 
       fill = NULL,
       color = NULL) + 
  theme_minimal() + 
  
  plot_theme

ordplotPigColor

ggsave(ordplotPigColor, filename = "ordplotPigColor.png", width = 7, height = 5, bg = "white")
```

# Pairs Plots - Pigments and Anions {.tabset}

## Photosynthetic
```{r, warning = FALSE, message = FALSE}
corplotAPChl <- dfAnion %>% 
  ggpairs(  
  columns = c("ChlA", "ChlB", "Fo", "Cl", "NO2NO3", "PO43", "SO42"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE, method = "spearman"))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAPChl

ggsave(corplotAPChl, filename = "corplotAPChl.png", width = 7, height = 7, bg = "white")
```

## Photoprotective 
```{r, warning = FALSE, message = FALSE}
corplotAPPhotoprotective <- dfAnion %>% 
  ggpairs(  
  columns = c("Astaxanthin", "Lutein", "BetaC", "Fo", "Cl", "NO2NO3", "PO43", "SO42"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE, method = "spearman"))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAPPhotoprotective

ggsave(corplotAPPhotoprotective, filename = "corplotAPPhotoprotective.png", width = 7, height = 7, bg = "white")
```

## Degradation 
```{r, warning = FALSE, message = FALSE}
corplotAPDegradation <- dfAnion %>% 
  ggpairs(  
  columns = c("Pheophorbide", "Pheophytin", "Fo", "Cl", "NO2NO3", "PO43", "SO42"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE, method = "spearman"))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAPDegradation

ggsave(corplotAPDegradation, filename = "corplotAPDegradation.png", width = 7, height = 7, bg = "white")
```

# Heat Maps - Pigments & Anions

```{r}
pigment_vars <- c("ChlA", "ChlB", "Astaxanthin", "Lutein", "BetaC", "Pheophorbide", "Pheophytin")

anion_vars <- c("Fo", "PO43", "SO42", "NO2NO3", "Cl")

# Correlation function
get_pigment_anion_corr <- function(data, group_name) {

  cor_mat <- rcorr(as.matrix(data), type = "spearman")

  r_vals <- cor_mat$r
  p_vals <- cor_mat$P

  expand.grid(
    Pigment = pigment_vars,
    Anion   = anion_vars,
    stringsAsFactors = FALSE
  ) %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Anion, Pigment],
      p_value    = p_vals[Anion, Pigment],
      significant = p_value < 0.05,
      label = paste0(
        round(Spearman_r, 2),
        "\n(p=", round(p_value, 3), ")"
      ),
      Group = group_name
    ) %>%
    ungroup()
}

# Red & Green (separate) 
cor_color <- dfAnion %>%
  filter(Color %in% c("red", "green")) %>%
  select(Color, Algae, all_of(pigment_vars), all_of(anion_vars)) %>%
  na.omit() %>%
  group_split(Color) %>%
  map_dfr(~get_pigment_anion_corr(
    data = .x %>% select(-Color, -Algae),
    group_name = unique(.x$Color)
  ))

# Algae (red + green pooled)
cor_algae <- dfAnion %>%
  filter(Algae == "algae") %>%
  select(all_of(pigment_vars), all_of(anion_vars)) %>%
  na.omit() %>%
  get_pigment_anion_corr("algae")

# All samples (red + green + clean) 
cor_all <- dfAnion %>%
  select(all_of(pigment_vars), all_of(anion_vars)) %>%
  na.omit() %>%
  get_pigment_anion_corr("all")

# Combine 
cor_results_all <- bind_rows(cor_color, cor_algae, cor_all)

# Factor ordering
cor_results_all$Pigment <- factor(cor_results_all$Pigment, levels = pigment_vars)
cor_results_all$Anion   <- factor(cor_results_all$Anion, levels = anion_vars)
cor_results_all$Group   <- factor(
  cor_results_all$Group,
  levels = c("red", "green", "algae", "all")
)

# Plot 
heatmapPigAnionAP <- ggplot(
  cor_results_all,
  aes(x = Pigment, y = Anion, fill = Spearman_r)
) +
  geom_tile() +
  geom_tile(
    data = filter(cor_results_all, significant),
    color = "black",
    linewidth = 0.75
  ) +
  geom_text(aes(label = label), size = 3) +
  facet_wrap(~Group, ncol = 1) +
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0,
    name = "Spearman r"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

heatmapPigAnionAP

# Save 
ggsave(
  heatmapPigAnionAP,
  filename = "heatmapPigAnionAP.png",
  width = 7,
  height = 10,
  bg = "white"
)
```

# Cell Density & Cellular Pigment Concentration

```{r, warning = FALSE, message = FALSE}
corplotAPCell <- dfPigCell %>% 
  ggpairs(  
  columns = c("CellDensity", "CellVolume", "Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE, method = "spearman"))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAPCell

ggsave(corplotAPCell, filename = "corplotAPCell.png", width = 7, height = 7, bg = "white")
```

## Heat Maps {.tabset}

### All Samples
```{r}
# Define pigment and Cell variables
pigment_vars <- c("ChlA", "ChlB", "Astaxanthin", "Lutein", "BetaC", "Pheophorbide", "Pheophytin")
cell_vars <- c("CellDensity", "CellVolume")

# Filter to just the necessary columns
df_filtered <- dfPigCell %>%
  select(Location, all_of(pigment_vars), all_of(cell_vars)) %>%
  na.omit()

# Function to compute pigment–Cell Spearman correlations per location
get_pigment_Cell_corr <- function(data, location_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")

  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Create all pigment–Cell pair combinations
  pairs <- expand.grid(Pigment = pigment_vars, Cell = cell_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Cell, Pigment],  # rows are Cells, cols are pigments
      p_value = p_vals[Cell, Pigment],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      Location = location_name
    ) %>%
    ungroup()

  return(results)
}

# Compute correlations for each location
cor_results <- df_filtered %>%
  group_split(Location) %>%
  map_dfr(~get_pigment_Cell_corr(.x, unique(.x$Location)))

# Convert to factors to preserve axis order
cor_results$Pigment <- factor(cor_results$Pigment, levels = pigment_vars)
cor_results$Cell <- factor(cor_results$Cell, levels = cell_vars)

# Plot
heatmapPigCell <- ggplot(cor_results, aes(x = Pigment, y = Cell, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  facet_wrap(~Location, ncol =1) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

heatmapPigCell

ggsave(heatmapPigCell, filename = "heatmapPigCell.png", width = 7, height = 9, bg = "white")
```

### Algae Samples
```{r}
# Define pigment and Cell variables
pigment_vars <- c("ChlA", "ChlB", "Astaxanthin", "Lutein", "BetaC", "Pheophorbide", "Pheophytin")
cell_vars <- c("CellDensity", "CellVolume")

# Filter to just the necessary columns
df_filtered <- dfPigCell %>%
  filter(Algae == "algae") %>% 
  select(Location, all_of(pigment_vars), all_of(cell_vars)) %>%
  na.omit()

# Function to compute pigment–Cell Spearman correlations per location
get_pigment_Cell_corr <- function(data, location_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")

  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Create all pigment–Cell pair combinations
  pairs <- expand.grid(Pigment = pigment_vars, Cell = cell_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Cell, Pigment],  # rows are Cells, cols are pigments
      p_value = p_vals[Cell, Pigment],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      Location = location_name
    ) %>%
    ungroup()

  return(results)
}

# Compute correlations for each location
cor_results <- df_filtered %>%
  group_split(Location) %>%
  map_dfr(~get_pigment_Cell_corr(.x, unique(.x$Location)))

# Convert to factors to preserve axis order
cor_results$Pigment <- factor(cor_results$Pigment, levels = pigment_vars)
cor_results$Cell <- factor(cor_results$Cell, levels = cell_vars)

# Plot
heatmapPigCellAlgae <- ggplot(cor_results, aes(x = Pigment, y = Cell, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  facet_wrap(~Location, ncol =1) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

heatmapPigCellAlgae

ggsave(heatmapPigCellAlgae, filename = "heatmapPigCellAlgae.png", width = 7, height = 9, bg = "white")
```

### Antarctic Peninsula (Red & Green)
```{r}
# Filter to Antarctic Peninsula samples with algae, and select necessary columns
df_filtered <- dfPigCell %>%
  filter(Algae == "algae", Location == "Antarctic Peninsula") %>%
  select(Color, all_of(pigment_vars), all_of(cell_vars)) %>%
  na.omit()

# Function to compute pigment–Cell Spearman correlations per snow color
get_pigment_Cell_corr <- function(data, color_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")

  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Create all pigment–Cell pair combinations
  pairs <- expand.grid(Pigment = pigment_vars, Cell = cell_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Cell, Pigment],  # rows are Cells, cols are pigments
      p_value = p_vals[Cell, Pigment],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      Color = color_name
    ) %>%
    ungroup()

  return(results)
}

# Compute correlations for each snow color
cor_results_color <- df_filtered %>%
  group_split(Color) %>%
  map_dfr(~get_pigment_Cell_corr(.x, unique(.x$Color)))

# Convert to factors to preserve axis order
cor_results_color$Pigment <- factor(cor_results_color$Pigment, levels = pigment_vars)
cor_results_color$Cell <- factor(cor_results_color$Cell, levels = cell_vars)

# Plot heatmaps by snow color
heatmapPigCellAP <- ggplot(cor_results_color, aes(x = Pigment, y = Cell, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results_color, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  facet_wrap(~Color, ncol = 1) + 
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

heatmapPigCellAP

# Save
ggsave(heatmapPigCellAP, filename = "heatmapPigCellAP.png", width = 7, height = 7, bg = "white")
```

# Publication Figures

## Stacked Bar Plot

```{r}
# Summarize by median composition
dfPig_median <- dfPig_long %>%
  # AP only, red + green only
  filter(Color %in% c("red", "green")) %>%
  mutate(
    Region  = factor(Region, levels = c("North", "South")),
    Color   = factor(Color, levels = c("green", "red")),
    Pigment = factor(
      Pigment,
      levels = c("Pheophorbide", "Pheophytin", "BetaC",
                 "Lutein", "Astaxanthin", "ChlB", "ChlA")
    )
  ) %>%
  group_by(Region, Color, Pigment) %>%
  dplyr::summarize(
    MedianComp = median(Composition, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Convert to proportion within each color group per region
  group_by(Region, Color) %>%
  mutate(MedianProp = MedianComp / sum(MedianComp, na.rm = TRUE)) %>%
  ungroup()

# Define a spacing factor (<1 reduces space between regions)
region_spacing <- 0.7

dfPig_plot <- dfPig_median %>%
  mutate(
    x = as.numeric(Region) * region_spacing +
      case_when(
        Color == "green" ~ -0.3,
        Color == "red"   ~  0
      )
  )

# Recalculate region centers for x-axis labels
region_centers <- dfPig_plot %>%
  group_by(Region) %>%
  dplyr::summarize(center = mean(x, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(match(Region, levels(dfPig_plot$Region)))

# Define pigment labels
pigLabels <- c(
  "Pheophorbide" = "Pheophorbide",
  "Pheophytin"  = "Pheophytin",
  "BetaC"       = "β-Carotene",
  "Lutein"      = "Lutein",
  "ChlB"        = "Chlorophyll b",
  "ChlA"        = "Chlorophyll a"
)

# Plot
barplotPigStacked <- ggplot(
  dfPig_plot,
  aes(x = x, y = MedianProp, fill = Pigment)
) +
  geom_bar(stat = "identity", width = 0.27) +
  scale_x_continuous(
    breaks = region_centers$center,
    labels = region_centers$Region,
    expand = expansion(mult = c(0.05, 0.05))
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = NULL,
    y = "Pigment Composition",
    fill = NULL
  ) +
  scale_fill_manual(values = pigPal, labels = pigLabels) +
  theme_minimal() +
  plot_themePub +
  theme(
    axis.text.y   = element_text(margin = margin(r = 10)),
    axis.text.x   = element_text(margin = margin(t = 5)),
    axis.ticks.x  = element_line(),
    axis.ticks.y  = element_blank(),
    legend.position = "none"
  ) +
  coord_flip()

barplotPigStacked

# Save
ggsave(
  barplotPigStacked,
  filename = "barplotPigStacked.png",
  width = 2.75,
  height = 4.5,
  bg = "white"
)
```

```{r}
# Summarize by median composition (Color only)
dfPig_median <- dfPig_long %>%
  # red + green only
  filter(Color %in% c("red", "green")) %>%
  mutate(
    Color   = factor(Color, levels = c("green", "red")),
    Pigment = factor(
      Pigment,
      levels = c("Pheophorbide", "Pheophytin", "BetaC",
                 "Lutein", "Astaxanthin", "ChlB", "ChlA")
    )
  ) %>%
  group_by(Color, Pigment) %>%
  dplyr::summarize(
    MedianComp = median(Composition, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Convert to proportion within each color group
  group_by(Color) %>%
  mutate(MedianProp = MedianComp / sum(MedianComp, na.rm = TRUE)) %>%
  ungroup()

# Pigment labels
pigLabels <- c(
  "Pheophorbide" = "Pheophorbide",
  "Pheophytin"  = "Pheophytin",
  "BetaC"       = "β-Carotene",
  "Lutein"      = "Lutein",
  "ChlB"        = "Chlorophyll b",
  "ChlA"        = "Chlorophyll a"
)

# Plot
barplotPigStacked <- ggplot(
  dfPig_median,
  aes(x = Color, y = MedianProp, fill = Pigment)
) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_x_discrete(
  expand = expansion(
    mult = c(0.05, 0.05),  # tighter between bars
    add  = c(1.0, 1.0)       # extra space at plot edges
  )
) +
  scale_fill_manual(values = pigPal, labels = pigLabels) +
  labs(
    x = NULL,
    y = "Pigment Composition",
    fill = NULL
  ) +
  theme_minimal() +
  plot_themePub +
  theme(
    axis.text.x   = element_text(margin = margin(t = 5)),
    axis.text.y   = element_text(margin = margin(r = 10)),
    axis.ticks.y  = element_blank(),
    legend.position = "none"
  )

barplotPigStacked

# Save
ggsave(
  barplotPigStacked,
  filename = "barplotPigStacked_color.png",
  width = 3,
  height = 4.5,
  bg = "white"
)
```


