---
title: "Data Visualization & Analysis - Anion Concentration"
author: "Ellie Ryan"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: flatly
    highlight: zenburn
    toc: true
    toc_float: true
    toc_collapsed: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Created:** 7/23/2025

**Last Modified:** 2/18/2026

```{r, message=FALSE, warning=FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(stringr)
library(kableExtra)
library(scales)
library(GGally)
library(patchwork)
library(Hmisc)
library(reshape2)
library(vegan)
library(purrr)
library(car)
library(FSA)
library(viridis)
library(ggrepel)
library(paletteer)
library(ggforce)
```

```{r}
colorPal <- c(
  "clean" = "#95879f",
  "red" = "#b04a25",
  "green" = "#647d6a"
)

algaePal <- c(
  "clean" = "#a5bdc0",
  "algae" = "#735a69"
)

pigPal <- c(
  "Astaxanthin" = "#b43f1f",
  "Lutein" = "#f2bc1a",
  "ChlB" = "#439468",
  "ChlA" = "#37a133",
  "BetaC" = "#bf7915",
  "Pheophorbide" = "#534f8d", 
  "Pheophytin" = "#95879f"
)

anionPal <- c(
  "PO43" = "#7887A4FF",
  "Cl" = "#9B6981FF",
  "SO42" = "#682C37FF",
  "Fo" = "#F6955EFF",
  "NO2NO3" = "#A8CDECFF"
)

paletteer_d("nationalparkcolors::Arches")

plot_theme <- theme(
  text = element_text(family = "Helvetica"),
  axis.title = element_text(size = 12, face = "bold"), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  plot.margin = margin(20, 20, 20, 20)
)

plot_themePub <- theme(
  text = element_text(family = "Helvetica"),
  axis.title.x = element_text(size = 12, margin = margin(t = 10)),  # t = top margin
  axis.title.y = element_text(size = 12, margin = margin(r = 10)),  # r = right margin
  axis.text = element_text(size = 12),
  axis.line = element_blank(),
  axis.ticks.x = element_blank(),
  axis.ticks.y = element_line(color = "black", linewidth = 0.25),  # Add tick marks
  axis.ticks.length = unit(-0.15, "cm"),  # Length of tick marks
  panel.grid = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
  legend.position = "top",
  legend.justification = c(0, 1),
  legend.text = element_text(size = 12),
  legend.title = element_text(size = 12),
  plot.title = element_text(size = 14, face = "bold"),
  plot.subtitle = element_text(size = 13, face = "italic"),
  plot.margin = margin(20, 20, 20, 20)
)

plot_theme2 <- theme(
  text = element_text(family = "Helvetica"),
  axis.title = element_text(size = 12, face = "bold"), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  axis.text.x = element_text(angle = 45, hjust = 1), 
  plot.margin = margin(20, 20, 20, 20)
)

pairsplot_theme <- theme(
  text = element_text(family = "Helvetica", size = 10),
  axis.text = element_text(size = 8),  # Adjust axis text size
  plot.title = element_text(size = 12, face = "bold"),  # Adjust plot title size
  legend.text = element_text(size = 12))  # Adjust legend text size
```

# Overview

The purpose of this `.Rmd` is to visualize anion concentration (µM) by snow color, and obtain basic summary statistics associated with each anion. I will additionally perform an ordination to reduce the dimensions of my anion data for simpler visualizations. Before proceeding, I will clean the dataset so that it features only the variables I am interested in for this analysis. In this dataset, 107 samples are used from the Antarctic Peninsula, following the removal of 6 outliers. These observations are broken down by `Color`, or the apparent snow color at time of sample collection. To keep things simple, there are only three levels to the `Color` category: red, green and clean. The breakdown of samples collected in the WAP is as follows:

-   Antarctic Peninsula:
    -   107 samples
        -   43 red
        -   46 green
        -   18 clean
        
From anion outlier analysis, the following outliers were identified for removal:

-   2025-02-19-SI-algae3 (green)
-   2025-02-24-DA-algae1 (green)
-   2025-02-24-DA-algae2 (red)
-   2025-02-24-DA-algae6 (green)
-   2025-03-07-NH-algae3 (clean)
-   2025-01-12-PI-algae10 (red)


See the Preliminary Results and Figures doc for justification for removal of each. 

## Data Wrangling

```{r}
df <- read.csv("CompiledDataPI.csv")
str(df)
```

Adding an `Algae` column to classify samples as algae or clean snow samples and a `LocColor` column to classify AP algae samples as red or green. 

```{r}
# Format date column 
df <- df %>%
  mutate(
    Date = as.Date(Date, format = "%m/%d/%y"))

# Add Algae column
df <- df %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  relocate(Algae, .after = Color)

# Add LocAbbrev
df <- df %>% mutate(LocAbbrev = case_when(
  Location == "Antarctic Peninsula" ~ paste("WAP")
)) %>%  
  relocate(LocAbbrev, .after = Location)

# Convert anion concentrations to µM
df <- df %>% 
  mutate(Fo = 22.21*Fo,
         Cl = 28.21*Cl,
         NO2 = 21.73*NO2,
         NO3 = 16.13*NO3, 
         PO43 = 10.53*PO43, 
         SO42 = 10.41*SO42)
```

Remove outliers from analysis: 
```{r}
outliers <- c(
  "2025-02-19-SI-algae3",
  "2025-02-24-DA-algae2",
  "2025-02-24-DA-algae6",
  "2025-03-07-NH-algae3",
  "2025-01-12-PI-algae10",
  "2025-02-24-DA-algae1"
  )

df <- df %>% 
  filter(!SampleID %in% outliers)
```

### Anions Dataset
```{r}
dfAnion <- df %>% 
  select(c("SampleID", "Date", "Location", "Site", "LocAbbrev", "Color", "Algae", Fo:SO42)) %>% 
  mutate(NO2NO3 = NO2 + NO3) %>% 
  relocate(NO2NO3, .after = Cl) %>% 
  select(!NO2) %>% 
  select(!NO3) %>% 
  na.omit(dfAnion)

# Add N:P Ratio
dfAnion <- dfAnion %>% 
  mutate(NP_Ratio = ifelse(PO43 > 0, NO2NO3 / PO43, NA_real_))
```

### Anions Dataset (Separate NO2, NO3)
```{r}
dfAnionN <- df %>% 
  select(c("SampleID", "Date", "Location", "Site", "LocAbbrev", "Color", "Algae", Fo:SO42)) %>% 
  na.omit(dfAnion)

```

### Anions (Pivoted)
```{r}
dfAnion_long <- dfAnion %>% 
  pivot_longer(cols = c(Fo, Cl, SO42, PO43, NO2NO3),
               names_to = "Anion",
               values_to = "Concentration")
```

# Summary Tables {.tabset}
```{r}
# Helper function: significant figures + commas + preserve .0 for integers
fmt_sigfig <- function(x, sig = 4) {
  # Round to significant figures
  y <- signif(x, sig)
  
  # Convert to string with commas
  out <- prettyNum(y, big.mark = ",", scientific = FALSE)
  
  # Add ".0" if number is an integer after rounding
  out[grepl("^\\d+$", out)] <- paste0(out[grepl("^\\d+$", out)], ".0")
  
  return(out)
}
```


## Formate
```{r}
# Anion Summary (formate)
FoSummary <- dfAnion %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    Fo_mean = fmt_sigfig(mean(Fo, na.rm = TRUE)),
    Fo_median = fmt_sigfig(median(Fo, na.rm = TRUE)),
    Fo_max = fmt_sigfig(max(Fo, na.rm = TRUE)),
    Fo_min = fmt_sigfig(min(Fo, na.rm = TRUE)),
    Fo_sd = fmt_sigfig(sd(Fo, na.rm = TRUE)),
    .groups = "drop"
  )

# Add "all" for each site
FoSummaryAll <- dfAnion %>% 
  summarise(
    Color = "all",
    n = n(),
    Fo_mean = fmt_sigfig(mean(Fo, na.rm = TRUE)),
    Fo_median = fmt_sigfig(median(Fo, na.rm = TRUE)),
    Fo_max = fmt_sigfig(max(Fo, na.rm = TRUE)),
    Fo_min = fmt_sigfig(min(Fo, na.rm = TRUE)),
    Fo_sd = fmt_sigfig(sd(Fo, na.rm = TRUE)),
    .groups = "drop"
  )

# Filter Antarctic red + green
FoSummaryAPRG <- dfAnion %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Color = "red+green",
    n = n(),
    Fo_mean = fmt_sigfig(mean(Fo, na.rm = TRUE)),
    Fo_median = fmt_sigfig(median(Fo, na.rm = TRUE)),
    Fo_max = fmt_sigfig(max(Fo, na.rm = TRUE)),
    Fo_min = fmt_sigfig(min(Fo, na.rm = TRUE)),
    Fo_sd = fmt_sigfig(sd(Fo, na.rm = TRUE)),
  )

# Combine and order rows
FoSummary <- bind_rows(FoSummaryAll, FoSummary, FoSummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
FoSummary %>% 
  kable("html", caption = "Summary Statistics for Formate by Snow Color (µM)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Chloride
```{r}
# Anion Summary (Chloride)
ClSummary <- dfAnion %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    Cl_mean = fmt_sigfig(mean(Cl, na.rm = TRUE)),
    Cl_median = fmt_sigfig(median(Cl, na.rm = TRUE)),
    Cl_max = fmt_sigfig(max(Cl, na.rm = TRUE)),
    Cl_min = fmt_sigfig(min(Cl, na.rm = TRUE)),
    Cl_sd = fmt_sigfig(sd(Cl, na.rm = TRUE)),
    .groups = "drop"
  )

# Add "all" for each site
ClSummaryAll <- dfAnion %>% 
  summarise(
    Color = "all",
    n = n(),
    Cl_mean = fmt_sigfig(mean(Cl, na.rm = TRUE)),
    Cl_median = fmt_sigfig(median(Cl, na.rm = TRUE)),
    Cl_max = fmt_sigfig(max(Cl, na.rm = TRUE)),
    Cl_min = fmt_sigfig(min(Cl, na.rm = TRUE)),
    Cl_sd = fmt_sigfig(sd(Cl, na.rm = TRUE)),
    .groups = "drop"
  )

# Filter Antarctic red + green
ClSummaryAPRG <- dfAnion %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Color = "red+green",
    n = n(),
    Cl_mean = fmt_sigfig(mean(Cl, na.rm = TRUE)),
    Cl_median = fmt_sigfig(median(Cl, na.rm = TRUE)),
    Cl_max = fmt_sigfig(max(Cl, na.rm = TRUE)),
    Cl_min = fmt_sigfig(min(Cl, na.rm = TRUE)),
    Cl_sd = fmt_sigfig(sd(Cl, na.rm = TRUE)),
  )

# Combine and order rows
ClSummary <- bind_rows(ClSummaryAll, ClSummary, ClSummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
ClSummary %>% 
  kable("html", caption = "Summary Statistics for Chloride by Snow Color (µM)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Nitrate/Nitrite {.tabset}

### NO2
```{r}
# Anion Summary (NO2)
NO2Summary <- dfAnionN %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    NO2_mean = fmt_sigfig(mean(NO2, na.rm = TRUE)),
    NO2_median = fmt_sigfig(median(NO2, na.rm = TRUE)),
    NO2_max = fmt_sigfig(max(NO2, na.rm = TRUE)),
    NO2_min = fmt_sigfig(min(NO2, na.rm = TRUE)),
    NO2_sd = fmt_sigfig(sd(NO2, na.rm = TRUE)),
    .groups = "drop"
  )

# Add "all" for each site
NO2SummaryAll <- dfAnionN %>% 
  summarise(
    Color = "all",
    n = n(),
    NO2_mean = fmt_sigfig(mean(NO2, na.rm = TRUE)),
    NO2_median = fmt_sigfig(median(NO2, na.rm = TRUE)),
    NO2_max = fmt_sigfig(max(NO2, na.rm = TRUE)),
    NO2_min = fmt_sigfig(min(NO2, na.rm = TRUE)),
    NO2_sd = fmt_sigfig(sd(NO2, na.rm = TRUE)),
    .groups = "drop"
  )

# Filter Antarctic red + green
NO2SummaryAPRG <- dfAnionN %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Color = "red+green",
    n = n(),
    NO2_mean = fmt_sigfig(mean(NO2, na.rm = TRUE)),
    NO2_median = fmt_sigfig(median(NO2, na.rm = TRUE)),
    NO2_max = fmt_sigfig(max(NO2, na.rm = TRUE)),
    NO2_min = fmt_sigfig(min(NO2, na.rm = TRUE)),
    NO2_sd = fmt_sigfig(sd(NO2, na.rm = TRUE)),
  )

# Combine and order rows
NO2Summary <- bind_rows(NO2SummaryAll, NO2Summary, NO2SummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
NO2Summary %>% 
  kable("html", caption = "Summary Statistics for NO2 by Snow Color (µM)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

### NO3
```{r}
# Anion Summary (NO3)
NO3Summary <- dfAnionN %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    NO3_mean = fmt_sigfig(mean(NO3, na.rm = TRUE)),
    NO3_median = fmt_sigfig(median(NO3, na.rm = TRUE)),
    NO3_max = fmt_sigfig(max(NO3, na.rm = TRUE)),
    NO3_min = fmt_sigfig(min(NO3, na.rm = TRUE)),
    NO3_sd = fmt_sigfig(sd(NO3, na.rm = TRUE)),
    .groups = "drop"
  )

# Add "all" for each site
NO3SummaryAll <- dfAnionN %>% 
  summarise(
    Color = "all",
    n = n(),
    NO3_mean = fmt_sigfig(mean(NO3, na.rm = TRUE)),
    NO3_median = fmt_sigfig(median(NO3, na.rm = TRUE)),
    NO3_max = fmt_sigfig(max(NO3, na.rm = TRUE)),
    NO3_min = fmt_sigfig(min(NO3, na.rm = TRUE)),
    NO3_sd = fmt_sigfig(sd(NO3, na.rm = TRUE)),
    .groups = "drop"
  )

# Filter Antarctic red + green
NO3SummaryAPRG <- dfAnionN %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Color = "red+green",
    n = n(),
    NO3_mean = fmt_sigfig(mean(NO3, na.rm = TRUE)),
    NO3_median = fmt_sigfig(median(NO3, na.rm = TRUE)),
    NO3_max = fmt_sigfig(max(NO3, na.rm = TRUE)),
    NO3_min = fmt_sigfig(min(NO3, na.rm = TRUE)),
    NO3_sd = fmt_sigfig(sd(NO3, na.rm = TRUE)),
  )

# Combine and order rows
NO3Summary <- bind_rows(NO3SummaryAll, NO3Summary, NO3SummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
NO3Summary %>% 
  kable("html", caption = "Summary Statistics for NO3 by Snow Color (µM)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

### NOx
```{r}
# Anion Summary (NOx)
NO2NO3Summary <- dfAnion %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    NOx_mean = fmt_sigfig(mean(NO2NO3, na.rm = TRUE)),
    NOx_median = fmt_sigfig(median(NO2NO3, na.rm = TRUE)),
    NOx_max = fmt_sigfig(max(NO2NO3, na.rm = TRUE)),
    NOx_min = fmt_sigfig(min(NO2NO3, na.rm = TRUE)),
    NOx_sd = fmt_sigfig(sd(NO2NO3, na.rm = TRUE)),
    .groups = "drop"
  )

# Add "all" for each site
NO2NO3SummaryAll <- dfAnion %>% 
  summarise(
    Color = "all",
    n = n(),
    NOx_mean = fmt_sigfig(mean(NO2NO3, na.rm = TRUE)),
    NOx_median = fmt_sigfig(median(NO2NO3, na.rm = TRUE)),
    NOx_max = fmt_sigfig(max(NO2NO3, na.rm = TRUE)),
    NOx_min = fmt_sigfig(min(NO2NO3, na.rm = TRUE)),
    NOx_sd = fmt_sigfig(sd(NO2NO3, na.rm = TRUE)),
    .groups = "drop"
  )

# Filter Antarctic red + green
NO2NO3SummaryAPRG <- dfAnion %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Color = "red+green",
    n = n(),
    NOx_mean = fmt_sigfig(mean(NO2NO3, na.rm = TRUE)),
    NOx_median = fmt_sigfig(median(NO2NO3, na.rm = TRUE)),
    NOx_max = fmt_sigfig(max(NO2NO3, na.rm = TRUE)),
    NOx_min = fmt_sigfig(min(NO2NO3, na.rm = TRUE)),
    NOx_sd = fmt_sigfig(sd(NO2NO3, na.rm = TRUE)),
  )

# Combine and order rows
NO2NO3Summary <- bind_rows(NO2NO3SummaryAll, NO2NO3Summary, NO2NO3SummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
NO2NO3Summary %>% 
  kable("html", caption = "Summary Statistics for NOx by Snow Color (µM)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Sulfate
```{r}
# Anion Summary (Sulfate)
SO42Summary <- dfAnion %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    SO42_mean = fmt_sigfig(mean(SO42, na.rm = TRUE)),
    SO42_median = fmt_sigfig(median(SO42, na.rm = TRUE)),
    SO42_max = fmt_sigfig(max(SO42, na.rm = TRUE)),
    SO42_min = fmt_sigfig(min(SO42, na.rm = TRUE)),
    SO42_sd = fmt_sigfig(sd(SO42, na.rm = TRUE)),
    .groups = "drop"
  )

# Add "all" for each site
SO42SummaryAll <- dfAnion %>% 
  summarise(
    Color = "all",
    n = n(),
    SO42_mean = fmt_sigfig(mean(SO42, na.rm = TRUE)),
    SO42_median = fmt_sigfig(median(SO42, na.rm = TRUE)),
    SO42_max = fmt_sigfig(max(SO42, na.rm = TRUE)),
    SO42_min = fmt_sigfig(min(SO42, na.rm = TRUE)),
    SO42_sd = fmt_sigfig(sd(SO42, na.rm = TRUE)),
    .groups = "drop"
  )

# Filter Antarctic red + green
SO42SummaryAPRG <- dfAnion %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Color = "red+green",
    n = n(),
    SO42_mean = fmt_sigfig(mean(SO42, na.rm = TRUE)),
    SO42_median = fmt_sigfig(median(SO42, na.rm = TRUE)),
    SO42_max = fmt_sigfig(max(SO42, na.rm = TRUE)),
    SO42_min = fmt_sigfig(min(SO42, na.rm = TRUE)),
    SO42_sd = fmt_sigfig(sd(SO42, na.rm = TRUE)),
  )

# Combine and order rows
SO42Summary <- bind_rows(SO42SummaryAll, SO42Summary, SO42SummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
SO42Summary %>% 
  kable("html", caption = "Summary Statistics for Sulfate by Snow Color (µM)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Phosphate
```{r}
# Anion Summary (Phosphate)
PO43Summary <- dfAnion %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    PO43_mean = fmt_sigfig(mean(PO43, na.rm = TRUE)),
    PO43_median = fmt_sigfig(median(PO43, na.rm = TRUE)),
    PO43_max = fmt_sigfig(max(PO43, na.rm = TRUE)),
    PO43_min = fmt_sigfig(min(PO43, na.rm = TRUE)),
    PO43_sd = fmt_sigfig(sd(PO43, na.rm = TRUE)),
    .groups = "drop"
  )

# Add "all" for each site
PO43SummaryAll <- dfAnion %>% 
  summarise(
    Color = "all",
    n = n(),
    PO43_mean = fmt_sigfig(mean(PO43, na.rm = TRUE)),
    PO43_median = fmt_sigfig(median(PO43, na.rm = TRUE)),
    PO43_max = fmt_sigfig(max(PO43, na.rm = TRUE)),
    PO43_min = fmt_sigfig(min(PO43, na.rm = TRUE)),
    PO43_sd = fmt_sigfig(sd(PO43, na.rm = TRUE)),
    .groups = "drop"
  )

# Filter Antarctic red + green
PO43SummaryAPRG <- dfAnion %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Color = "red+green",
    n = n(),
    PO43_mean = fmt_sigfig(mean(PO43, na.rm = TRUE)),
    PO43_median = fmt_sigfig(median(PO43, na.rm = TRUE)),
    PO43_max = fmt_sigfig(max(PO43, na.rm = TRUE)),
    PO43_min = fmt_sigfig(min(PO43, na.rm = TRUE)),
    PO43_sd = fmt_sigfig(sd(PO43, na.rm = TRUE)),
  )

# Combine and order rows
PO43Summary <- bind_rows(PO43SummaryAll, PO43Summary, PO43SummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
PO43Summary %>% 
  kable("html", caption = "Summary Statistics for Phosphate by Snow Color (µM)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

# ANOVA/Kruskal Wallis Test 

## Check ANOVA Assumptions
Before proceeding with an ANOVA test to assess differences in anion concentration between snow colors, I am going to check to see if these variables follow the assumptions of normality and equal variance required to run an ANOVA: 
```{r}
# Step 1: Group by pigment and nest data
nested <- dfAnion_long %>%
  group_by(Anion) %>%
  nest()

# Step 2: Fit ANOVA model, extract residuals, and run tests
anovaAssumptionsResults <- nested %>%
  mutate(
    # Fit ANOVA model
    aov_model = map(data, ~ aov(Concentration ~ Color, data = .x)),
    
    # Shapiro-Wilk test for normality of residuals
    shapiro_p = map_dbl(aov_model, ~ shapiro.test(residuals(.x))$p.value),
    
    # Levene's Test for homogeneity of variance
    levene_p = map_dbl(data, ~leveneTest(Concentration ~ factor(Color), data = .x)$`Pr(>F)`[1])
  )
```

From the below results, it appears that none of the anions are both normally distributed and have equal variances. Because of this, I am going to proceed with a Kruskal-Wallis test in lieu of an ANOVA. 

```{r}
anovaAssumptionsResults %>%
  select(Anion, shapiro_p, levene_p) %>%
  mutate(
    normality_ok = shapiro_p > 0.05,
    homogeneity_ok = levene_p > 0.05
  )
```

## Kruskal-Wallis Test {.tabset}

### Anions

```{r}
# Run kruskal wallis test 
kruskalResultsAP <- dfAnion_long %>%
  filter(Location == "Antarctic Peninsula") %>% 
  group_by(Anion) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ Color, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsAP <- kruskalResultsAP %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(Color), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

#### Visualize Results
```{r}
dunnPairsAP <- kruskalResultsAP %>%
  select(Anion, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_AP <- ggplot(dunnPairsAP, aes(x = Comparison, y = Anion, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsAP, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_AP

ggsave(dunnPlot_AP, filename = "dunnPlot_AP.png", width = 7, height = 6, bg = "white")
```

### N:P Ratio
```{r}
dfRatio <- dfAnion %>%
  filter(!is.na(NP_Ratio), !is.na(Color)) %>%
  mutate(Color = factor(Color))

# Run kruskal wallis test 
kruskalResultsNP <- kruskal.test(NP_Ratio ~ Color, data = dfRatio)

kruskalResultsNP$statistic
kruskalResultsNP$p.value
```

```{r}
dunnResultsNP <- dunnTest(NP_Ratio ~ Color,
                        data = dfRatio,
                        method = "bonferroni")

dunnResultsNP
```

# Box Plots {.tabset}

## All Anions (All Samples)
```{r}
boxplotAll <- dfAnion_long %>%
  ggplot(aes(x = Anion, y = Concentration, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Conecntration (µM)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotAll

ggsave(boxplotAll, filename = "boxplotAll.png", width = 7, height = 5, bg = "white")
```

## Formate
```{r}
boxplotFo <- dfAnion %>%
  ggplot(aes(x = Color, y = Fo, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (µM)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotFo

ggsave(boxplotFo, filename = "boxplotFo.png", width = 7, height = 5, bg = "white")
```

## Chloride
```{r}
boxplotCl <- dfAnion %>%
  ggplot(aes(x = Color, y = Cl, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (µM)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotCl

ggsave(boxplotCl, filename = "boxplotCl.png", width = 7, height = 5, bg = "white")
```

## Nitrate/Nitrite
```{r}
boxplotNO2NO3 <- dfAnion %>%
  ggplot(aes(x = Color, y = NO2NO3, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (µM)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotNO2NO3

ggsave(boxplotNO2NO3, filename = "boxplotNO2NO3.png", width = 7, height = 5, bg = "white")
```

## Phosphate
```{r}
boxplotPO43 <- dfAnion %>%
  ggplot(aes(x = Color, y = PO43, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (µM)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotPO43

ggsave(boxplotPO43, filename = "boxplotPO43.png", width = 7, height = 5, bg = "white")
```

## Sulfate
```{r}
boxplotSO42 <- dfAnion %>%
  ggplot(aes(x = Color, y = SO42, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (µM)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotSO42

ggsave(boxplotSO42, filename = "boxplotSO42.png", width = 7, height = 5, bg = "white")
```

# Pairs Plots - Pigments {.tabset}

```{r, warning = FALSE, message = FALSE}
corplotAll <- ggpairs(
  dfAnion,  
  columns = c("Fo", "Cl", "NO2NO3", "PO43", "SO42"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAll

ggsave(corplotAll, filename = "corplotAll.png", width = 7, height = 7, bg = "white")
```

# Heat Maps - Anions {.tabset}

## All Samples
```{r}
cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
anion_vars <- c("Fo", "Cl", "NO2NO3", "PO43", "SO42")

# List of locations
locations <- unique(dfAnion$LocAbbrev)

# Function to calculate correlation matrices for each location
get_corr_data <- function(location) {
  data_sub <- dfAnion %>%
    filter(LocAbbrev == location) %>%
    select(all_of(anion_vars)) %>%
    na.omit()

  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p

  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)

  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Location <- location

  return(merged)
}

# Apply across locations
heatmap_data <- do.call(rbind, lapply(locations, get_corr_data))

# Factor levels for correct ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = anion_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = anion_vars)

# Create label
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Remove upper triangle + diagonal if desired (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot
heatmapAnionAll <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Location) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank()) 

heatmapAnionAll

ggsave(heatmapAnionAll, filename = "heatmapAnionAll.png", width = 11.5, height = 7, bg = "white")

```

## Algae Samples
```{r}
cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
anion_vars <- c("Fo", "Cl", "NO2NO3", "PO43", "SO42")

# List of locations
locations <- unique(dfAnion$LocAbbrev)

# Function to calculate correlation matrices for each location
get_corr_data <- function(location) {
  data_sub <- dfAnion %>%
    filter(Algae == "algae") %>% 
    filter(LocAbbrev == location) %>%
    select(all_of(anion_vars)) %>%
    na.omit()

  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p

  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)

  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Location <- location

  return(merged)
}

# Apply across locations
heatmap_data <- do.call(rbind, lapply(locations, get_corr_data))

# Factor levels for correct ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = anion_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = anion_vars)

# Create label
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Remove upper triangle + diagonal if desired (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot
heatmapAnionAlgae <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Location) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank()) 

heatmapAnionAlgae

ggsave(heatmapAnionAlgae, filename = "heatmapAnionAlgae.png", width = 11.5, height = 7, bg = "white")

```

## Clean Samples
```{r}
cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
anion_vars <- c("Fo", "Cl", "NO2NO3", "PO43", "SO42")

# List of locations
locations <- unique(dfAnion$LocAbbrev)

# Function to calculate correlation matrices for each location
get_corr_data <- function(location) {
  data_sub <- dfAnion %>%
    filter(Algae == "clean") %>% 
    filter(LocAbbrev == location) %>%
    select(all_of(anion_vars)) %>%
    na.omit()

  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p

  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)

  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Location <- location

  return(merged)
}

# Apply across locations
heatmap_data <- do.call(rbind, lapply(locations, get_corr_data))

# Factor levels for correct ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = anion_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = anion_vars)

# Create label
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Remove upper triangle + diagonal if desired (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot
heatmapAnionClean <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Location) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank()) 

heatmapAnionClean

ggsave(heatmapAnionClean, filename = "heatmapAnionClean.png", width = 11.5, height = 7, bg = "white")

```

## Red vs. Green
```{r}
cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
anion_vars <- c("Fo", "Cl", "NO2NO3", "PO43", "SO42")

# Filter data once for Antarctic Peninsula red and green samples
df_filtered <- dfAnion %>%
  filter(Location == "Antarctic Peninsula",
         Color %in% c("red", "green"))

# Unique colors to iterate over
colors <- unique(df_filtered$Color)

# Function to calculate correlation matrices for each color group
get_corr_data <- function(color_group) {
  data_sub <- df_filtered %>%
    filter(Color == color_group) %>%
    select(all_of(anion_vars)) %>%
    na.omit()
  
  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p
  
  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)
  
  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Color <- color_group
  
  return(merged)
}

# Get correlation data for each color
heatmap_data_list <- lapply(colors, get_corr_data)
heatmap_data <- do.call(rbind, heatmap_data_list)

# Set factor levels for ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = anion_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = anion_vars)

# Create labels for plot
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Keep only lower triangle (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot faceted by Color
heatmapAnionAPRG <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Color) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank())
  
heatmapAnionAPRG
ggsave(heatmapAnionAPRG, filename = "heatmapAnionAPRG.png", width = 9, height = 5, bg = "white")
```

# Principal Components Analysis - Anions
From the above plots, it seems as though pigment data are relatively linear. I am going to proceed with a PCA ordination over a nonparametric ordination, like nMDS. 

## Standardize Data & Run PCA 
Below, I use prcomp to standardize panion data and run a PCA. Even though the anion concentrations are on the same scale, some anions are present in much higher concentrations than others. Because of this, I will standardize anions so they are all on the same scale.

PC1 explains 52.1% of the variance in pigment concentration. PC2 explains 20.49% of the variance in pigment concentration. Collectively, these two components explain 72.6% of the variance, and I will likely only include these two PCs in my analysis.
```{r}
Anions <- dfAnion[8:12]

AnionPCA <- prcomp(Anions, scale = TRUE)

summary(AnionPCA)
```

## Biplots
```{r}
biplot(AnionPCA)
```

## Ordination Plots {.tabset}

```{r}
## Add PC1 &2 to dataset
dfAnion$PC1 <- AnionPCA$x[,1]
dfAnion$PC2 <- AnionPCA$x[,2]
```

### By Location {.tabset}

#### Original

```{r}
ordplotAnion <- dfAnion %>% 
  ggplot(aes(x = PC1, y = PC2, color = Color, fill = Color)) + 
  geom_point(size = 2.5, alpha = 0.7) + 
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(x = "PC1 (52.1%)", 
       y = "PC2 (20.5%)", 
       fill = NULL,
       color = NULL) + 
  theme_minimal() + 
  plot_theme

ordplotAnion

ggsave(ordplotAnion, filename = "ordplotAnion.png", width = 7, height = 7, bg = "white")
```

#####  Plot loadings
```{r}
# Extract loadings (rotation = variable loadings on PCs)
loadings <- as.data.frame(AnionPCA$rotation[,1:2])

# Rescale them a bit to fit on the plot (optional)
loadings <- loadings * 5  # tweak multiplier as needed for arrow length
loadings$Anion <- rownames(loadings)

# Ordination with arrows
ordplotAnion + 
  geom_segment(data = loadings, aes(x = 0, y = 0, xend = PC1, yend = PC2),
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "black",
               inherit.aes = FALSE) +
  geom_text_repel(data = loadings, aes(x = PC1, y = PC2, label = Anion),
                  color = "black", size = 4,
                  inherit.aes = FALSE,
                  max.overlaps = Inf) 
```


#### Sized (PO43)

```{r}
ordplotAnionColor_PO43 <- dfAnion %>% 
  ggplot(aes(x = PC1, y = PC2, size = PO43, color = Color, fill = Color)) + 
  geom_point(alpha = 0.7) + 
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  scale_size_continuous(labels = label_comma()) +
  labs(x = "PC1 (52.1%)", 
       y = "PC2 (20.5%)",
       size = "Phosphate (µM)",
       fill = NULL,
       color = NULL) + 
  guides(size = guide_legend(order = 1),
         color = guide_legend(order = 2),
         fill = guide_legend(order = 2)) +
  theme_minimal() + 
  plot_theme +
  theme(legend.position = "right")

ordplotAnionColor_PO43

ggsave(ordplotAnionColor_PO43, filename = "ordplotAnionColor_PO43.png", width = 7, height = 5, bg = "white")
```

# Publication Figures 

## Stacked Bar Plot
```{r}
dfAnion_median <- dfAnion_long %>%
  mutate(
    Color = factor(Color, levels = c("clean", "red", "green")),
    Anion = factor(Anion, levels = c("Cl", "PO43", "SO42", "NO2NO3", "Fo"))
  ) %>%
  group_by(Anion, Color) %>%
  summarise(
    MedianConc = median(Concentration, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
anion_labels <- c(
  "Cl\u207B",              # Cl−
  "PO\u2084\u00B3\u207B",   # PO4³−
  "SO\u2084\u00B2\u207B",  # SO4²−
  "NO\u2093\u207B",        # NO3−
  "HCOO\u207B"           # Fo−
  )
```

```{r}
barplotAnionStacked <- ggplot(
  dfAnion_median,
  aes(x = Anion, y = MedianConc, fill = Color)
) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_x_discrete(labels = anion_labels) +
  labs(
    x = NULL,
    y = expression(Concentration ~ "(µM)"),
    fill = NULL
  ) +
  scale_fill_manual(values = colorPal) +   # your existing color palette
  scale_y_continuous(
    limits = c(0, 160),
    breaks = seq(0, 160, by = 20),
    expand = c(0, 0)) +
  theme_minimal() +
  plot_themePub +
  theme(
    axis.text.x = element_text(margin = margin(t = 10)),
    legend.position = "bottom"
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0),
      keywidth = unit(0.75, "lines"),
      keyheight = unit(0.75, "lines")
    )
  )

barplotAnionStacked

ggsave(
  barplotAnionStacked,
  filename = "barplotAnionStacked.png",
  width = 6,
  height = 5,
  bg = "white"
)
```

## Grouped Bar Plot

```{r}
dfAnion_median <- dfAnion_median %>%
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))
  )

barplotAnionGrouped <- ggplot(
  dfAnion_median,
  aes(x = Anion, y = MedianConc, fill = Color)
) +
  geom_bar(
    stat = "identity",
    width = 0.65,
    position = position_dodge(width = 0.7)
  ) +
  scale_x_discrete(labels = anion_labels) +
  labs(
    x = NULL,
    y = expression(Concentration ~ "(µM)"),
    fill = NULL
  ) +
  scale_fill_manual(values = colorPal) +
  scale_y_continuous(
    limits = c(0, 120),
    breaks = seq(0, 120, by = 20),
    expand = c(0, 0)) + 
  theme_minimal() +
  plot_themePub +
  theme(
    axis.text.x = element_text(margin = margin(t = 10)),
    legend.position = "bottom"
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0),
      keywidth = unit(0.75, "lines"),
      keyheight = unit(0.75, "lines")
    )
  )

barplotAnionGrouped

ggsave(
  barplotAnionGrouped,
  filename = "barplotAnionGrouped.png",
  width = 6,
  height = 6,
  bg = "white"
)
```

## Grouped Box Plot

```{r}
dfAnion_box <- dfAnion_long %>%
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean")),
    Anion = factor(Anion, levels = c("Cl", "PO43", "SO42", "NO2NO3", "Fo"))
  )
```

```{r}
boxplotAnionGrouped <- ggplot(
  dfAnion_box,
  aes(x = Anion, y = Concentration, color = Color, fill = Color)
) +
  geom_boxplot(
    width = 0.6,
    position = position_dodge(width = 0.7),
    outlier.shape = 21,
    outlier.size = 1.5, 
    alpha = 0.85
  ) +
  scale_x_discrete(labels = anion_labels) +
  labs(
    x = NULL,
    y = expression(Concentration ~ "(µM)"),
    fill = NULL
  ) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  scale_y_sqrt(
    breaks = c(0, 1, 4, 10, 50, 200, 800, 1200)
  ) +
  theme_minimal() +
  plot_themePub +
  theme(
    axis.text.x = element_text(margin = margin(t = 10)),
    legend.position = "bottom"
  ) +
  guides(
  fill = guide_legend(
    title = NULL,
    override.aes = list(linetype = 0),
    keywidth = unit(0.75, "lines"),
    keyheight = unit(0.75, "lines")
  ),
  color = "none"
)

boxplotAnionGrouped

ggsave(
  boxplotAnionGrouped,
  filename = "boxplotAnionGrouped.png",
  width = 6,
  height = 6,
  bg = "white"
)
```

