---
title: "Data Vizualization - Pigment Concentration"
author: "Ellie Ryan"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: flatly
    highlight: zenburn
    toc: true
    toc_float: true
    toc_collapsed: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Created:** 7/1/2025

**Last Modified:** 2/18/2026

```{r, message=FALSE, warning=FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(stringr)
library(kableExtra)
library(scales)
library(GGally)
library(patchwork)
library(Hmisc)
library(reshape2)
library(car)
library(FSA)
library(vegan)
library(DT)
library(scales)
library(ggpmisc)
```

```{r}
regionPal <- c(
  "North" = "#284C46FF",
  "South" = "#b04a25"
) 

colorPal <- c(
  "clean" = "#95879f",
  "red" = "#b04a25",
  "green" = "#647d6a"
)

algaePal <- c(
  "clean" = "#a5bdc0",
  "algae" = "#735a69"
)

pigPal <- c(
  "Astaxanthin" = "#b04a25",
  "Lutein" = "#d9983b",
  "ChlA" = "#284C46FF",
  "ChlB" = "#647d6a",
  "BetaC" = "#4d1d05",
  "Pheophorbide" = "#414c66", 
  "Pheophytin" = "#95879f"
)

plot_theme <- theme(
  axis.title = element_text(size = 12, face = "bold"), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  plot.margin = margin(20, 20, 20, 20)
)

plot_themePub <- theme(
  text = element_text(family = "Helvetica"),
  axis.title.x = element_text(size = 10, margin = margin(t = 2)),  # t = top margin
  axis.title.y = element_text(size = 10, margin = margin(r = 4)),  # r = right margin
  axis.text = element_text(size = 10),
  axis.line = element_blank(),
  axis.ticks.x = element_blank(),
  axis.ticks.y = element_line(color = "black", linewidth = 0.25),  # Add tick marks
  axis.ticks.length = unit(-0.15, "cm"),  # Length of tick marks
  panel.grid = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
  legend.position = "top",
  legend.justification = c(0, 1),
  legend.text = element_text(size = 10),
  legend.title = element_text(size = 10),
  plot.title = element_text(size = 14, face = "bold"),
  plot.subtitle = element_text(size = 13, face = "italic"),
  plot.margin = margin(3, 3, 3, 3)
)

plot_theme2 <- theme(
  axis.title = element_text(size = 12, face = "bold"), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  axis.text.x = element_text(angle = 45, hjust = 1), 
  plot.margin = margin(20, 20, 20, 20)
)

pairsplot_theme <- theme(
  text = element_text(size = 10),  # Adjust variable name text size
  axis.text = element_text(size = 8),  # Adjust axis text size
  plot.title = element_text(size = 12, face = "bold"),  # Adjust plot title size
  legend.text = element_text(size = 12))  # Adjust legend text size
```

# Overview

The purpose of this `.Rmd` is to visualize raw, non-transformed pigment concentration by site and color, and obtain basic summary statistics associated with each pigment in order to inform my next steps in analysis. I will additionally perform an ordination to reduce the dimensions of my pigment data for simpler visualizations. Before proceeding, I will clean the dataset so that it features only the variables I am interested in for this analysis. In this dataset, 108 samples are used from the Antarctic Peninsula, following the removal of 5 outliers. These observations are broken down by `Color`, or the apparent snow color at time of sample collection. To keep things simple, there are only three levels to the `Color` category: red, green and clean. The breakdown of samples collected in the WAP is as follows:

-   Antarctic Peninsula:
    -   107 samples
        -   43 red
        -   46 green
        -   18 clean

*NOTE: Fewer samples were used for analysis, since there are NA values for pigment concentrations of some samples.*

From anion outlier analysis, the following outliers were identified for removal:

-   2025-02-19-SI-algae3
-   2025-02-24-DA-algae2
-   2025-02-24-DA-algae6
-   2025-03-07-NH-algae3
-   2025-01-12-PI-algae10

See the Thesis Figures (Preliminary) doc for justification for removal of each. Likely, more samples will be removed following visualization of pigment distribution.

## Data Wrangling

```{r}
df <- read.csv("CompiledDataPI.csv")
str(df)
```

Adding an `Algae` column to classify samples as algae or clean snow samples and a `Region` column to separate North and South samples.

```{r}
# Format date column 
df <- df %>%
  mutate(
    Date = as.Date(Date, format = "%m/%d/%y"))

# Add Algae
df <- df %>%  
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  relocate(Algae, .after = Color)

# Add Region
df <- df %>% 
  mutate(
  Region = if_else(
    Site %in% c("Horseshoe Island", "Stonington Island"),
    "South",
    "North")) %>% 
  relocate(Region, .after = Site)

# Add LocAbbrev
df <- df %>% mutate(LocAbbrev = case_when(
  Location == "Antarctic Peninsula" ~ paste("WAP")
)) %>%  
  relocate(LocAbbrev, .after = Site)
```

```{r}
str(df)
```

```{r}
outliers <- c(
  "2025-02-19-SI-algae3",
  "2025-02-24-DA-algae2",
  "2025-02-24-DA-algae6",
  "2025-03-07-NH-algae3",
  "2025-01-12-PI-algae10")

df <- df %>% 
  filter(!SampleID %in% outliers)

# Add BetaC:Astax Ratio
df <- df %>% 
  mutate(BetaC_Astax_Ratio = ifelse(Astaxanthin > 0, BetaC/Astaxanthin, NA_real_)) %>% 
  relocate(BetaC_Astax_Ratio, .after = Pheophytin)

# Add Chla:Chlb Ratio
df <- df %>% 
  mutate(ChlA_ChlB_Ratio = ifelse(ChlB > 0, ChlA/ChlB, NA_real_)) %>% 
  relocate(ChlA_ChlB_Ratio, .after = BetaC_Astax_Ratio)

# Add Chla:Astax Ratio
df <- df %>% 
  mutate(ChlA_Astax_Ratio = ifelse(Astaxanthin > 0, ChlA/Astaxanthin, NA_real_)) %>% 
  relocate(ChlA_Astax_Ratio, .after = ChlA_ChlB_Ratio)


```

### Pigments

```{r}
dfPig <- df %>% 
  select(c("SampleID", "Date", "Location", "Site", "Region", "LocAbbrev", "Color", "Algae", Astaxanthin:Pheophytin)) %>% 
  na.omit(dfPig)
```


### Pigments & Cells

```{r}
dfPigCell <- df %>% 
  select(c("SampleID", "Date", "Location", "Site", "Region", "LocAbbrev", "Color", "Algae", "CellDensity", "CellLength", Astaxanthin:Pheophytin)) %>% 
  na.omit(dfPig)

dfPigCell <- dfPigCell %>% 
  mutate(
    # Calulate volume (µm^3)
    CellVolume = (pi/6)*(CellLength^3)) %>% 
  relocate(CellVolume, .after = CellLength)
```

### Pigments & Anions

```{r}
dfAnion <- df %>% 
  select(c("SampleID", "Date", "Location", "Site", "Region", "LocAbbrev", "Color", "Algae", Astaxanthin:Pheophytin, Fo:SO42)) %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  mutate(NO2NO3 = NO2 + NO3) %>% 
  relocate(Algae, .after = Color) %>% 
  relocate(NO2NO3, .after = Cl) %>% 
  select(!NO2) %>% 
  select(!NO3) %>% 
  na.omit(dfAnion)

outliersAnion <- c("2025-02-24-DA-algae1")

dfAnion <- dfAnion %>% 
  filter(!SampleID %in% outliersAnion)

```

### Pigment Ratios & Anions

```{r}
dfAnion_Ratio <- df %>% 
  select(c("SampleID", "Date", "Location", "Site", "Region", "LocAbbrev", "Color", "Algae", BetaC_Astax_Ratio:ChlA_Astax_Ratio, Fo:SO42)) %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  mutate(NO2NO3 = NO2 + NO3) %>% 
  relocate(Algae, .after = Color) %>% 
  relocate(NO2NO3, .after = Cl) %>% 
  select(!NO2) %>% 
  select(!NO3) %>% 
  na.omit(dfAnion_Ratio)

outliersAnion <- c("2025-02-24-DA-algae1")

dfAnion_Ratio <- dfAnion_Ratio %>% 
  filter(!SampleID %in% outliersAnion)

```

### Cell & Anion
```{r}
dfAnionCell <- df %>% 
  select(c("SampleID", "Date", "Location", "Site", "Region", "LocAbbrev", "Color", "Algae", "CellDensity", "CellLength", Fo:SO42)) %>% 
  na.omit(dfAnionCell)

dfAnionCell <- dfAnionCell %>% 
  mutate(
    # Calulate volume (µm^3)
    CellVolume = (pi/6)*(CellLength^3)) %>% 
  relocate(CellVolume, .after = CellLength)
```

### Normalized Pigment & Anions
```{r}
dfNorm <- df %>% 
  mutate(ChlA = (ChlA/CellDensity)*1000,
         ChlB = (ChlB/CellDensity)*1000,
         Astaxanthin = (Astaxanthin/CellDensity)*1000,
         Lutein = (Lutein/CellDensity)*1000,
         BetaC = (BetaC/CellDensity)*1000,
         Pheophytin = (Pheophytin/CellDensity)*1000,
         Pheophorbide = (Pheophorbide/CellDensity)*1000) 

dfNorm <- dfNorm %>% 
  select(c("SampleID", "Date", "Location", "Site", "Region", "LocAbbrev", "Color", "Algae", Astaxanthin:Pheophytin, Fo:SO42)) %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  mutate(NO2NO3 = NO2 + NO3) %>% 
  relocate(Algae, .after = Color) %>% 
  relocate(NO2NO3, .after = Cl) %>% 
  select(!NO2) %>% 
  select(!NO3) %>% 
  na.omit(dfNorm)
```

### Pigments (Pivoted)

```{r}
dfPig_long <- dfPig %>% 
  pivot_longer(cols = Astaxanthin:Pheophytin,
               names_to = "Pigment",
               values_to = "Concentration")
```

### Pigment Ratios (Pivoted)
```{r}
dfPigRatio_long <- dfAnion_Ratio %>% 
  pivot_longer(cols = BetaC_Astax_Ratio:ChlA_Astax_Ratio,
               names_to = "Pigments",
               values_to = "Ratio")
```

### Anions (Pivoted)

```{r}
dfAnion_long <- dfAnion %>% 
  pivot_longer(cols = c(Fo, Cl, SO42, PO43, NO2NO3),
               names_to = "Anion",
               values_to = "Concentration")
```

# Summary Tables {.tabset}

## Primary Pigments

```{r}
# Helper function: significant figures + commas + preserve .0 for integers
fmt_sigfig <- function(x, sig = 4) {
  # Round to significant figures
  y <- signif(x, sig)
  
  # Convert to string with commas
  out <- prettyNum(y, big.mark = ",", scientific = FALSE)
  
  # Add ".0" if number is an integer after rounding
  out[grepl("^\\d+$", out)] <- paste0(out[grepl("^\\d+$", out)], ".0")
  
  return(out)
}

# Pigment Summary (ChlA and ChlB)
ChlSummary <- dfPig %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    ChlA_mean   = fmt_sigfig(mean(ChlA, na.rm = TRUE)),
    ChlA_median = fmt_sigfig(median(ChlA, na.rm = TRUE)),
    ChlA_max    = fmt_sigfig(max(ChlA, na.rm = TRUE)),
    ChlA_min    = fmt_sigfig(min(ChlA, na.rm = TRUE)),
    ChlA_sd     = fmt_sigfig(sd(ChlA, na.rm = TRUE)),
    ChlB_mean   = fmt_sigfig(mean(ChlB, na.rm = TRUE)),
    ChlB_median = fmt_sigfig(median(ChlB, na.rm = TRUE)),
    ChlB_max    = fmt_sigfig(max(ChlB, na.rm = TRUE)),
    ChlB_min    = fmt_sigfig(min(ChlB, na.rm = TRUE)),
    ChlB_sd     = fmt_sigfig(sd(ChlB, na.rm = TRUE)),
    .groups = "drop"
  )

# Add "all" for each site
ChlSummaryAll <- dfPig %>% 
  summarise(
    Color = "all",
    n = n(),
    ChlA_mean   = fmt_sigfig(mean(ChlA, na.rm = TRUE)),
    ChlA_median = fmt_sigfig(median(ChlA, na.rm = TRUE)),
    ChlA_max    = fmt_sigfig(max(ChlA, na.rm = TRUE)),
    ChlA_min    = fmt_sigfig(min(ChlA, na.rm = TRUE)),
    ChlA_sd     = fmt_sigfig(sd(ChlA, na.rm = TRUE)),
    ChlB_mean   = fmt_sigfig(mean(ChlB, na.rm = TRUE)),
    ChlB_median = fmt_sigfig(median(ChlB, na.rm = TRUE)),
    ChlB_max    = fmt_sigfig(max(ChlB, na.rm = TRUE)),
    ChlB_min    = fmt_sigfig(min(ChlB, na.rm = TRUE)),
    ChlB_sd     = fmt_sigfig(sd(ChlB, na.rm = TRUE)),
    .groups = "drop"
  )

# Filter Antarctic red + green
ChlSummaryAPRG <- dfPig %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Color = "red+green",
    n = n(),
    ChlA_mean   = fmt_sigfig(mean(ChlA, na.rm = TRUE)),
    ChlA_median = fmt_sigfig(median(ChlA, na.rm = TRUE)),
    ChlA_max    = fmt_sigfig(max(ChlA, na.rm = TRUE)),
    ChlA_min    = fmt_sigfig(min(ChlA, na.rm = TRUE)),
    ChlA_sd     = fmt_sigfig(sd(ChlA, na.rm = TRUE)),
    ChlB_mean   = fmt_sigfig(mean(ChlB, na.rm = TRUE)),
    ChlB_median = fmt_sigfig(median(ChlB, na.rm = TRUE)),
    ChlB_max    = fmt_sigfig(max(ChlB, na.rm = TRUE)),
    ChlB_min    = fmt_sigfig(min(ChlB, na.rm = TRUE)),
    ChlB_sd     = fmt_sigfig(sd(ChlB, na.rm = TRUE))
  )

# Combine and order rows
ChlSummary <- bind_rows(ChlSummaryAll, ChlSummary, ChlSummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
ChlSummary %>% 
  kable("html", caption = "Summary Statistics for ChlA and ChlB by Snow Color (ppb)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```


## Carotenoids

```{r}
# Pigment Summary (Astaxanthin, Beta Carotene and Lutein)
CarotenoidSummary <- dfPig %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    Astaxanthin_mean = fmt_sigfig(mean(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_median = fmt_sigfig(median(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_max = fmt_sigfig(max(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_min = fmt_sigfig(min(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_sd = fmt_sigfig(sd(Astaxanthin, na.rm = TRUE)),
    Lutein_mean = fmt_sigfig(mean(Lutein, na.rm = TRUE)),
    Lutein_median = fmt_sigfig(median(Lutein, na.rm = TRUE)),
    Lutein_max = fmt_sigfig(max(Lutein, na.rm = TRUE)),
    Lutein_min = fmt_sigfig(min(Lutein, na.rm = TRUE)),
    Lutein_sd = fmt_sigfig(sd(Lutein, na.rm = TRUE)),
    BetaC_mean = fmt_sigfig(mean(BetaC, na.rm = TRUE)),
    BetaC_median = fmt_sigfig(median(BetaC, na.rm = TRUE)),
    BetaC_max = fmt_sigfig(max(BetaC, na.rm = TRUE)),
    BetaC_min = fmt_sigfig(min(BetaC, na.rm = TRUE)),
    BetaC_sd = fmt_sigfig(sd(BetaC, na.rm = TRUE)),
    .groups = "drop"
  )

# Add "all" for each site
CarotenoidSummaryAll <- dfPig %>% 
  summarise(
    Color = "all",
    n = n(),
    Astaxanthin_mean = fmt_sigfig(mean(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_median = fmt_sigfig(median(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_max = fmt_sigfig(max(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_min = fmt_sigfig(min(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_sd = fmt_sigfig(sd(Astaxanthin, na.rm = TRUE)),
    Lutein_mean = fmt_sigfig(mean(Lutein, na.rm = TRUE)),
    Lutein_median = fmt_sigfig(median(Lutein, na.rm = TRUE)),
    Lutein_max = fmt_sigfig(max(Lutein, na.rm = TRUE)),
    Lutein_min = fmt_sigfig(min(Lutein, na.rm = TRUE)),
    Lutein_sd = fmt_sigfig(sd(Lutein, na.rm = TRUE)),
    BetaC_mean = fmt_sigfig(mean(BetaC, na.rm = TRUE)),
    BetaC_median = fmt_sigfig(median(BetaC, na.rm = TRUE)),
    BetaC_max = fmt_sigfig(max(BetaC, na.rm = TRUE)),
    BetaC_min = fmt_sigfig(min(BetaC, na.rm = TRUE)),
    BetaC_sd = fmt_sigfig(sd(BetaC, na.rm = TRUE)),
    .groups = "drop"
  )

# Filter Antarctic red + green
CarotenoidSummaryAPRG <- dfPig %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Location = "Antarctic Peninsula",
    Color = "red+green",
    n = n(),
    Astaxanthin_mean = fmt_sigfig(mean(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_median = fmt_sigfig(median(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_max = fmt_sigfig(max(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_min = fmt_sigfig(min(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_sd = fmt_sigfig(sd(Astaxanthin, na.rm = TRUE)),
    Lutein_mean = fmt_sigfig(mean(Lutein, na.rm = TRUE)),
    Lutein_median = fmt_sigfig(median(Lutein, na.rm = TRUE)),
    Lutein_max = fmt_sigfig(max(Lutein, na.rm = TRUE)),
    Lutein_min = fmt_sigfig(min(Lutein, na.rm = TRUE)),
    Lutein_sd = fmt_sigfig(sd(Lutein, na.rm = TRUE)),
    BetaC_mean = fmt_sigfig(mean(BetaC, na.rm = TRUE)),
    BetaC_median = fmt_sigfig(median(BetaC, na.rm = TRUE)),
    BetaC_max = fmt_sigfig(max(BetaC, na.rm = TRUE)),
    BetaC_min = fmt_sigfig(min(BetaC, na.rm = TRUE)),
    BetaC_sd = fmt_sigfig(sd(BetaC, na.rm = TRUE))
  )

# Combine and order rows
CarotenoidSummary <- bind_rows(CarotenoidSummaryAll, CarotenoidSummary, CarotenoidSummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
CarotenoidSummary %>% 
  kable("html", caption = "Summary Statistics for Astaxanthin and Lutein by Snow Color (ppb)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Degradation Products

```{r}
# Pigment Summary (Pheophytin and Pheophorbide)
DegSummary <- dfPig %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    Pheophytin_mean = fmt_sigfig(mean(Pheophytin, na.rm = TRUE)),
    Pheophytin_median = fmt_sigfig(median(Pheophytin, na.rm = TRUE)),
    Pheophytin_max = fmt_sigfig(max(Pheophytin, na.rm = TRUE)),
    Pheophytin_min = fmt_sigfig(min(Pheophytin, na.rm = TRUE)),
    Pheophytin_sd = fmt_sigfig(sd(Pheophytin, na.rm = TRUE)),
    Pheophorbide_mean = fmt_sigfig(mean(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_median = fmt_sigfig(median(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_max = fmt_sigfig(max(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_min = fmt_sigfig(min(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_sd = fmt_sigfig(sd(Pheophorbide, na.rm = TRUE)),
    .groups = "drop"
  )

# Add "all" for each site
DegSummaryAll <- dfPig %>% 
  summarise(
    Color = "all",
    n = n(),
    Pheophytin_mean = fmt_sigfig(mean(Pheophytin, na.rm = TRUE)),
    Pheophytin_median = fmt_sigfig(median(Pheophytin, na.rm = TRUE)),
    Pheophytin_max = fmt_sigfig(max(Pheophytin, na.rm = TRUE)),
    Pheophytin_min = fmt_sigfig(min(Pheophytin, na.rm = TRUE)),
    Pheophytin_sd = fmt_sigfig(sd(Pheophytin, na.rm = TRUE)),
    Pheophorbide_mean = fmt_sigfig(mean(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_median = fmt_sigfig(median(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_max = fmt_sigfig(max(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_min = fmt_sigfig(min(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_sd = fmt_sigfig(sd(Pheophorbide, na.rm = TRUE)),
    .groups = "drop"
  )

# Filter Antarctic red + green
DegSummaryAPRG <- dfPig %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Color = "red+green",
    n = n(),
    Pheophytin_mean = fmt_sigfig(mean(Pheophytin, na.rm = TRUE)),
    Pheophytin_median = fmt_sigfig(median(Pheophytin, na.rm = TRUE)),
    Pheophytin_max = fmt_sigfig(max(Pheophytin, na.rm = TRUE)),
    Pheophytin_min = fmt_sigfig(min(Pheophytin, na.rm = TRUE)),
    Pheophytin_sd = fmt_sigfig(sd(Pheophytin, na.rm = TRUE)),
    Pheophorbide_mean = fmt_sigfig(mean(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_median = fmt_sigfig(median(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_max = fmt_sigfig(max(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_min = fmt_sigfig(min(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_sd = fmt_sigfig(sd(Pheophorbide, na.rm = TRUE))
  )

# Combine and order rows
DegSummary <- bind_rows(DegSummaryAll, DegSummary, DegSummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
DegSummary %>% 
  kable("html", caption = "Summary Statistics for Pheophytin and Pheophorbide by Snow Color (ppb)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

# ANOVA/Kruskal Wallis Test

Before proceeding with an ANOVA test to assess differences in pigment concentration between locations, I am going to check to see if these variables follow the assumptions of normality and equal variance required to run an ANOVA:

```{r}
# Step 1: Group by pigment and nest data
nested <- dfPig_long %>%
  group_by(Pigment) %>%
  nest()

# Step 2: Fit ANOVA model, extract residuals, and run tests
anovaAssumptionsResults <- nested %>%
  mutate(
    # Fit ANOVA model
    aov_model = map(data, ~ aov(Concentration ~ Color, data = .x)),
    
    # Shapiro-Wilk test for normality of residuals
    shapiro_p = map_dbl(aov_model, ~ shapiro.test(residuals(.x))$p.value),
    
    # Levene's Test for homogeneity of variance
    levene_p = map_dbl(data, ~leveneTest(Concentration ~ factor(Color), data = .x)$`Pr(>F)`[1])
  )
```

From the below results, it appears that none of the pigments are both normally distributed and have equal variances. Because of this, I am going to proceed with a Kruskal-Wallis test in lieu of an ANOVA.

```{r}
anovaAssumptionsResults %>%
  select(Pigment, shapiro_p, levene_p) %>%
  mutate(
    normality_ok = shapiro_p > 0.05,
    homogeneity_ok = levene_p > 0.05
  )
```

## Kruskal-Wallis Test {.tabset}

### Pigments
Below, I use the Kruskal-Wallis test to run a non-parametric comparison of each pigment across each snow color. Then, I follow up with a Dunn's post hoc test for the pigments that show significant differences between the snow colors.

```{r}
# Run kruskal wallis test 
kruskalResultsAP <- dfPig_long %>%
  filter(Location == "Antarctic Peninsula") %>% 
  group_by(Pigment) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ Color, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsAP <- kruskalResultsAP %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(Color), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

#### Visualize Results

```{r}
dunnPairsAP <- kruskalResultsAP %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_AP <- ggplot(dunnPairsAP, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsAP, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_AP

ggsave(dunnPlot_AP, filename = "dunnPlot_AP.png", width = 7, height = 6, bg = "white")
```

### Pigment Ratios
Below, I use the Kruskal-Wallis test to run a non-parametric comparison of each pigment ratio across each snow color. Then, I follow up with a Dunn's post hoc test for the ratios that show significant differences between the snow colors.

```{r}
# Run kruskal wallis test 
KruskalResultsRatio <- dfPigRatio_long %>%
  filter(Location == "Antarctic Peninsula") %>% 
  group_by(Pigments) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Ratio ~ Color, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
KruskalResultsRatio <- KruskalResultsRatio %>%
  mutate(
    dunn = map(data, ~ dunnTest(Ratio ~ factor(Color), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

#### Visualize Results

```{r}
dunnPairsRatio <- KruskalResultsRatio %>%
  select(Pigments, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_Ratio <- ggplot(dunnPairsRatio, aes(x = Comparison, y = Pigments, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsRatio, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_Ratio

ggsave(dunnPlot_Ratio, filename = "dunnPlot_Ratio.png", width = 7, height = 6, bg = "white")
```

# Box Plots {.tabset}

## All Pigments

```{r}
boxplotAll <- dfPig_long %>%
  ggplot(aes(x = Pigment, y = Concentration, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotAll

ggsave(boxplotAll, filename = "boxplotAll.png", width = 7, height = 5, bg = "white")
```

## Astaxanthin

```{r}
boxplotAstax <- dfPig %>%
  ggplot(aes(x = Region, y = Astaxanthin, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotAstax

ggsave(boxplotAstax, filename = "boxplotAstax.png", width = 7, height = 5, bg = "white")
```

## Beta Carotene

```{r}
boxplotBetaC <- dfPig %>%
  ggplot(aes(x = Region, y = BetaC, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotBetaC

ggsave(boxplotBetaC, filename = "boxplotBetaC.png", width = 7, height = 5, bg = "white")
```

## ChlA

```{r}
boxplotChlA <- dfPig %>%
  ggplot(aes(x = Region, y = ChlA, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotChlA

ggsave(boxplotChlA, filename = "boxplotChlA.png", width = 7, height = 5, bg = "white")
```

## ChlB

```{r}
boxplotChlB <- dfPig %>%
  ggplot(aes(x = Region, y = ChlB, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotChlB

ggsave(boxplotChlB, filename = "boxplotChlB.png", width = 7, height = 5, bg = "white")
```

## Lutein

```{r}
boxplotLutein <- dfPig %>%
  ggplot(aes(x = Region, y = Lutein, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotLutein

ggsave(boxplotLutein, filename = "boxplotLutein.png", width = 7, height = 5, bg = "white")
```

## Pheophorbide

```{r}
boxplotPheophorbide <- dfPig %>%
  ggplot(aes(x = Region, y = Pheophorbide, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotPheophorbide

ggsave(boxplotPheophorbide, filename = "boxplotPheophorbide.png", width = 7, height = 5, bg = "white")
```

## Pheophytin

```{r}
boxplotPheophytin <- dfPig %>%
  ggplot(aes(x = Region, y = Pheophytin, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotPheophytin

ggsave(boxplotPheophytin, filename = "boxplotPheophytin.png", width = 7, height = 5, bg = "white")
```

# Pairs Plots - Pigments

```{r, warning = FALSE, message = FALSE}
corplotAll <- ggpairs(
  dfPig,  
  columns = c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAll

ggsave(corplotAll, filename = "corplotAll.png", width = 7, height = 7, bg = "white")
```


# Heat Maps - Pigments {.tabset}

## All Sites (All Samples)

```{r, message=FALSE, warning=FALSE}
cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
pigment_vars <- c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin")

# List of locations
locations <- unique(dfPig$LocAbbrev)

# Function to calculate correlation matrices for each location
get_corr_data <- function(location) {
  data_sub <- dfPig %>%
    filter(LocAbbrev == location) %>%
    select(all_of(pigment_vars)) %>%
    na.omit()

  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p

  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)

  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Location <- location

  return(merged)
}

# Apply across locations
heatmap_data <- do.call(rbind, lapply(locations, get_corr_data))

# Factor levels for correct ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = pigment_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = pigment_vars)

# Create label
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Remove upper triangle + diagonal if desired (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot
heatmapPigAll <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Location) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank()) 

heatmapPigAll

ggsave(heatmapPigAll, filename = "heatmapPigAll.png", width = 8, height = 5, bg = "white")

```


## Antarctic Peninsula (Red vs. Green)

```{r, message=FALSE, warning=FALSE}
cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
pigment_vars <- c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin")

# Filter data once for Antarctic Peninsula red and green samples
df_filtered <- dfPig %>%
  filter(Location == "Antarctic Peninsula",
         Color %in% c("red", "green"))

# Unique colors to iterate over
colors <- unique(df_filtered$Color)

# Function to calculate correlation matrices for each color group
get_corr_data <- function(color_group) {
  data_sub <- df_filtered %>%
    filter(Color == color_group) %>%
    select(all_of(pigment_vars)) %>%
    na.omit()
  
  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p
  
  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)
  
  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Color <- color_group
  
  return(merged)
}

# Get correlation data for each color
heatmap_data_list <- lapply(colors, get_corr_data)
heatmap_data <- do.call(rbind, heatmap_data_list)

# Set factor levels for ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = pigment_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = pigment_vars)

# Create labels for plot
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Keep only lower triangle (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot faceted by Color
heatmapPigAPRG <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Color) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank())
  
heatmapPigAPRG
ggsave(heatmapPigAPRG, filename = "heatmapPigAPRG.png", width = 8, height = 4.5, bg = "white")
```

# Principal Components Analysis - Pigments

From the above plots, it seems as though pigment data are relatively linear. I am going to proceed with a PCA ordination over a nonparametric ordination, like nMDS.

## Standardize Data & Run PCA

Below, I use prcomp to standardize pigment data and run a PCA. Even though the pigment concentrations are on the same scale, some pigments are present in much higher concentrations than others. Because of this, I will standardize pigments so they are all on the same scale.

PC1 explains 79.01% of the variance in pigment concentration. PC2 explains 13.26% of the variance in pigment concentration. Collectively, these two components explain 92.27% of the variance, and I will likely only include these two PCs in my analysis.

```{r}
Pigments <- dfAnion[9:15]

PigPCA <- prcomp(Pigments, scale = TRUE)

summary(PigPCA)
```

## Biplots

Biplots allow you to assess the influence and contribution of individual variables for each of the two compared PCs. They also allow you to explore relationships between variables and get a view of the positionings of each observation. The magnitude of loadings of each variable seem very similar to one another, meaning each variable explains a similar amount of the variability in the data. On the PC1 axis, the arrows are all pointing in a positive direction, meaning they are all positively correlated with one another along this axis. There is a bit more variation in vector direction on the PC2 axis. Pheophorbide, beta carotene, ChlA, and ChlB appear not to be explained at all on the second PC, while negative values on the PC2 axis are associated with higher values of Lutein and Astaxantin. Positive values on the PC2 axis are associated with higher concentrations of Pheophytin.

```{r}
biplot(PigPCA)
```

## PC Interpretation and Exploration

The scores for each pigment in PC1 are all relatively the same size, meaning each variable carries similar weight in the PC1 equation. These scores are all positive, meaning the pigments are all positively correlated with one another. The only exception here is Astaxanthin, which has a lower score than the other pigments for PC1.

The scores for PC2 are more varied. Astaxanthin has the largest score of all the pigments on this axis, it is also negative, as is the score for Lutein. The other pigments all have positive scores. This suggests that, on this axis, astaxanthin and lutein have contrasting correlations with the other pigments.

```{r}
PigPCA$rotation[,1:2]
```

## Ordination Plots {.tabset}

```{r}
## Add PC1 &2 to dataset
dfAnion$PC1 <- PigPCA$x[,1]
dfAnion$PC2 <- PigPCA$x[,2]
```

```{r}
ordplotPigColor <- dfAnion %>% 
  ggplot(aes(x = PC1, y = PC2, color = Color, fill = Color)) + 
  geom_point(alpha = 0.7, size = 2.5) + 
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(x = "PC1 (81.0%)", 
       y = "PC2 (11.76%)",
       fill = NULL,
       color = NULL) + 
  theme_minimal() + 
  plot_theme

ordplotPigColor

ggsave(ordplotPigColor, filename = "ordplotPigColor.png", width = 7, height = 5, bg = "white")
```

## Fit Anions to Ordination

In this section, I am going to fit the nutrient variables onto the PCA ordination.

The following variables significantly align with PCA axes: Fo, Cl, PO43, and SO42.

```{r}
anions <- dfAnion[16:20]

anionsScaled <- scale(anions)

pigAnionEnvfit <- envfit(PigPCA, anionsScaled)
pigAnionEnvfit
```

### Extract Significant Nutrient Vectors

```{r}
# Extract p-values
anionPvals <- as.data.frame(pigAnionEnvfit$vectors$pvals)
colnames(anionPvals) <- "pvals"

sigAnions <- rownames(anionPvals)[anionPvals$pvals < 0.05]
```

```{r}
# Extract only significant vectors from envfit
sigVectors <- as.data.frame(scores(pigAnionEnvfit, display = "vectors"))

sigVectors <- sigVectors[rownames(sigVectors) %in% sigAnions, ]

sigVectors$Anion <- rownames(sigVectors)
```

### Scale Vectors

```{r}
# Calculate scaling factor based on both PC axes
range_x <- range(dfAnion$PC1)
range_y <- range(dfAnion$PC2)

# Use a conservative multiplier to keep arrows within the plot range
scale_factor <- min(diff(range_x) / (2 * max(abs(sigVectors$PC1))),
                    diff(range_y) / (2 * max(abs(sigVectors$PC2)))) * 0.9

# Apply scaling
sigVectors_scaled <- sigVectors
sigVectors_scaled[, c("PC1", "PC2")] <- sigVectors_scaled[, c("PC1", "PC2")] * scale_factor
```

### Plot PCA with Significant Anion Vectors 

```{r}
ordplotColorAnion <- ordplotPigColor + 
  geom_segment(data = sigVectors_scaled, 
               aes(x = 0, y = 0, xend = PC1, yend = PC2),
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "black", 
               inherit.aes = FALSE) + 
  geom_text(data = sigVectors_scaled, 
            aes(x = PC1 * 1.11, y = PC2 * 1.20, label = Anion),
            color = "black", 
            inherit.aes = FALSE) 

ordplotColorAnion

ggsave(ordplotColorAnion, filename = "ordplotColorAnion.png", width = 6, height = 5, bg = "white")
```

# Pairs Plots - Pigments and Anions {.tabset}

## Photosynthetic

```{r, warning = FALSE, message = FALSE, fig.width=9, fig.height=9}
corplotAPChl <- dfAnion %>% 
  ggpairs(  
  columns = c("ChlA", "ChlB", "Fo", "Cl", "NO2NO3", "PO43", "SO42"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE, method = "spearman"))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAPChl

ggsave(corplotAPChl, filename = "corplotAPChl.png", width = 7, height = 7, bg = "white")
```

## Photoprotective

```{r, warning = FALSE, message = FALSE, fig.width=9, fig.height=9}
corplotAPPhotoprotective <- dfAnion %>% 
  ggpairs(  
  columns = c("Astaxanthin", "Lutein", "BetaC", "Fo", "Cl", "NO2NO3", "PO43", "SO42"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE, method = "spearman"))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAPPhotoprotective

ggsave(corplotAPPhotoprotective, filename = "corplotAPPhotoprotective.png", width = 7, height = 7, bg = "white")
```

## Degradation

```{r, warning = FALSE, message = FALSE, fig.width=9, fig.height=9}
corplotAPDegradation <- dfAnion %>% 
  ggpairs(  
  columns = c("Pheophorbide", "Pheophytin", "Fo", "Cl", "NO2NO3", "PO43", "SO42"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE, method = "spearman"))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAPDegradation

ggsave(corplotAPDegradation, filename = "corplotAPDegradation.png", width = 7, height = 7, bg = "white")
```

# Heat Maps - Pigments & Anions {.tabset}

## Pigments
```{r, fig.width=9, fig.height=10}
# Define pigment and anion variables
pigment_vars <- c("ChlA", "ChlB", "Astaxanthin", "Lutein", "BetaC", "Pheophorbide", "Pheophytin")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Correlation function
get_pigment_anion_corr <- function(data, group_name) {

  cor_mat <- rcorr(as.matrix(data), type = "spearman")

  r_vals <- cor_mat$r
  p_vals <- cor_mat$P

  expand.grid(
    Pigment = pigment_vars,
    Anion   = anion_vars,
    stringsAsFactors = FALSE
  ) %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Anion, Pigment],
      p_value    = p_vals[Anion, Pigment],
      significant = p_value < 0.05,
      label = paste0(
        round(Spearman_r, 2),
        "\n(p=", round(p_value, 3), ")"
      ),
      Group = group_name
    ) %>%
    ungroup()
}

# Red & Green (separate) 
cor_color <- dfAnion %>%
  filter(Color %in% c("red", "green")) %>%
  select(Color, Algae, all_of(pigment_vars), all_of(anion_vars)) %>%
  na.omit() %>%
  group_split(Color) %>%
  map_dfr(~get_pigment_anion_corr(
    data = .x %>% select(-Color, -Algae),
    group_name = unique(.x$Color)
  ))

# Algae (red + green pooled)
cor_algae <- dfAnion %>%
  filter(Algae == "algae") %>%
  select(all_of(pigment_vars), all_of(anion_vars)) %>%
  na.omit() %>%
  get_pigment_anion_corr("algae")

# All samples (red + green + clean) 
cor_all <- dfAnion %>%
  select(all_of(pigment_vars), all_of(anion_vars)) %>%
  na.omit() %>%
  get_pigment_anion_corr("all")

# Combine 
cor_results_all <- bind_rows(cor_color, cor_algae, cor_all)

# Factor ordering
cor_results_all$Pigment <- factor(cor_results_all$Pigment, levels = pigment_vars)
cor_results_all$Anion   <- factor(cor_results_all$Anion, levels = anion_vars)
cor_results_all$Group   <- factor(
  cor_results_all$Group,
  levels = c("red", "green", "algae", "all")
)

# Plot 
heatmapPigAnionAP <- ggplot(
  cor_results_all,
  aes(x = Pigment, y = Anion, fill = Spearman_r)
) +
  geom_tile() +
  geom_tile(
    data = filter(cor_results_all, significant),
    color = "black",
    linewidth = 0.75
  ) +
  geom_text(aes(label = label), size = 3) +
  facet_wrap(~Group, ncol = 1) +
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0,
    name = "Spearman r"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

heatmapPigAnionAP

# Save 
ggsave(
  heatmapPigAnionAP,
  filename = "heatmapPigAnionAP.png",
  width = 7,
  height = 10,
  bg = "white"
)
```

## Pigment Ratios
```{r, fig.width=9, fig.height=10}
# Define pigment and anion variables
pigment_vars <- c("ChlA_Astax_Ratio", "ChlA_ChlB_Ratio", "BetaC_Astax_Ratio")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Correlation function
get_pigment_anion_corr <- function(data, group_name) {

  cor_mat <- rcorr(as.matrix(data), type = "spearman")

  r_vals <- cor_mat$r
  p_vals <- cor_mat$P

  expand.grid(
    Pigment = pigment_vars,
    Anion   = anion_vars,
    stringsAsFactors = FALSE
  ) %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Anion, Pigment],
      p_value    = p_vals[Anion, Pigment],
      significant = p_value < 0.05,
      label = paste0(
        round(Spearman_r, 2),
        "\n(p=", round(p_value, 3), ")"
      ),
      Group = group_name
    ) %>%
    ungroup()
}

# Red & Green (separate) 
cor_color <- dfAnion_Ratio %>%
  filter(Color %in% c("red", "green")) %>%
  select(Color, Algae, all_of(pigment_vars), all_of(anion_vars)) %>%
  na.omit() %>%
  group_split(Color) %>%
  map_dfr(~get_pigment_anion_corr(
    data = .x %>% select(-Color, -Algae),
    group_name = unique(.x$Color)
  ))

# Algae (red + green pooled)
cor_algae <- dfAnion_Ratio %>%
  filter(Algae == "algae") %>%
  select(all_of(pigment_vars), all_of(anion_vars)) %>%
  na.omit() %>%
  get_pigment_anion_corr("algae")

# All samples (red + green + clean) 
cor_all <- dfAnion_Ratio %>%
  select(all_of(pigment_vars), all_of(anion_vars)) %>%
  na.omit() %>%
  get_pigment_anion_corr("all")

# Combine 
cor_results_all <- bind_rows(cor_color, cor_algae, cor_all)

# Factor ordering
cor_results_all$Pigment <- factor(cor_results_all$Pigment, levels = pigment_vars)
cor_results_all$Anion   <- factor(cor_results_all$Anion, levels = anion_vars)
cor_results_all$Group   <- factor(
  cor_results_all$Group,
  levels = c("red", "green", "algae", "all")
)

# Plot 
heatmapPigRatioAnion <- ggplot(
  cor_results_all,
  aes(x = Pigment, y = Anion, fill = Spearman_r)
) +
  geom_tile() +
  geom_tile(
    data = filter(cor_results_all, significant),
    color = "black",
    linewidth = 0.75
  ) +
  geom_text(aes(label = label), size = 3) +
  facet_wrap(~Group, ncol = 1) +
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0,
    name = "Spearman r"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

heatmapPigRatioAnion

# Save 
ggsave(
  heatmapPigRatioAnion,
  filename = "heatmapPigRatioAnion.png",
  width = 7,
  height = 10,
  bg = "white"
)
```

# Cell Density, Volume & Cellular Pigment Concentration

## Pairs Plots

```{r, warning = FALSE, message = FALSE, fig.width=9, fig.height=9}
corplotAPCell <- dfPigCell %>% 
  ggpairs(  
  columns = c("CellDensity", "CellVolume", "Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE, method = "spearman"))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAPCell

ggsave(corplotAPCell, filename = "corplotAPCell.png", width = 7, height = 7, bg = "white")
```

## Heat Maps 

```{r}
# Define pigment and cell variables
pigment_vars <- c("ChlA", "ChlB", "Astaxanthin", "Lutein", 
                  "BetaC", "Pheophorbide", "Pheophytin")
cell_vars <- c("CellDensity", "CellVolume")

# ---- Build datasets for each tier ----

# All samples combined (red + green + clean)
df_all <- dfPigCell %>%
  filter(Location == "Antarctic Peninsula") %>%
  mutate(Tier = "All samples") %>%
  select(Tier, all_of(pigment_vars), all_of(cell_vars)) %>%
  na.omit()

# Green only
df_green <- dfPigCell %>%
  filter(Location == "Antarctic Peninsula", Color == "green") %>%
  mutate(Tier = "Green") %>%
  select(Tier, all_of(pigment_vars), all_of(cell_vars)) %>%
  na.omit()

# Red only
df_red <- dfPigCell %>%
  filter(Location == "Antarctic Peninsula", Color == "red") %>%
  mutate(Tier = "Red") %>%
  select(Tier, all_of(pigment_vars), all_of(cell_vars)) %>%
  na.omit()

# Combine all tiers into one dataframe
df_combined <- bind_rows(df_all, df_green, df_red)

# ---- Function to compute pigment–cell Spearman correlations ----
get_pigment_Cell_corr <- function(data, tier_name) {

  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")

  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  pairs <- expand.grid(
    Pigment = pigment_vars,
    Cell    = cell_vars,
    stringsAsFactors = FALSE
  )

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Cell, Pigment],
      p_value    = p_vals[Cell, Pigment],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      Tier = tier_name
    ) %>%
    ungroup()

  return(results)
}

# ---- Compute correlations for each tier ----
cor_results <- df_combined %>%
  group_split(Tier) %>%
  map_dfr(~get_pigment_Cell_corr(.x, unique(.x$Tier)))

# Preserve axis and facet order
cor_results$Pigment <- factor(cor_results$Pigment, levels = pigment_vars)
cor_results$Cell    <- factor(cor_results$Cell, levels = cell_vars)
cor_results$Tier    <- factor(
  cor_results$Tier,
  levels = c("All samples", "Green", "Red")
)

# ---- Plot heatmap ----
heatmapPigCellAP <- ggplot(
  cor_results,
  aes(x = Pigment, y = Cell, fill = Spearman_r)
) +
  geom_tile() +
  geom_tile(
    data = filter(cor_results, significant),
    color = "black",
    linewidth = 0.75
  ) +
  geom_text(aes(label = label), size = 3) +
  facet_wrap(~Tier, ncol = 1) +
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0,
    name = "Spearman r"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

heatmapPigCellAP

# Save
ggsave(
  heatmapPigCellAP,
  filename = "heatmapPigCellAP.png",
  width = 7,
  height = 9,
  bg = "white"
)
```


# Publication Figures

## Algal Properties by Region {.tabset}

### Cell Density
```{r}
boxplotCellDensityPub <- dfPigCell %>% 
  mutate(
    Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = CellDensity, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Cell Density" ~ "(cells·mL"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 3e+04, 2e+05, 5e+05, 1e+06, 2e+06, 4e+06)) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotCellDensityPub
```

### Cell Volume
```{r}
boxplotCellVolumePub <- dfPigCell %>% 
  filter(CellVolume > 0) %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = CellVolume, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Cell Volume" ~ "(μm"^{3}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 1000, 2500, 5000, 10000, 20000, 30000),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotCellVolumePub
```

### Astaxanthin
```{r}
boxplotAstaxPub <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = Astaxanthin, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Astaxanthin" ~ "(μg·L"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 10, 50, 100, 250, 500, 1000),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotAstaxPub
```

### ChlA
```{r}
boxplotChlAPub <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = ChlA, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Chlorophyll a" ~ "(μg·L"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 250, 1500, 5000, 10000, 20000, 30000),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotChlAPub
```

### ChlB
```{r}
boxplotChlBPub <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = ChlB, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Chlorophyll b" ~ "(μg·L"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 150, 1000, 2500, 5000, 10000),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotChlBPub
```

### Lutein
```{r}
boxplotLuteinPub <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = Lutein, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Lutein" ~ "(μg·L"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 150, 1000, 2500, 5000, 10000, 15000),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotLuteinPub
```

### BetaC
```{r}
boxplotBetaCPub <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = BetaC, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("β-carotene" ~ "(μg·L"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 150, 1000, 2500, 5000, 10000, 15000),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotBetaCPub
```

### Pheophytin
```{r}
boxplotPheophytinPub <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = Pheophytin, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Pheophytin" ~ "(μg·L"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 150, 1000, 2500, 5000, 7500, 12500),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotPheophytinPub
```

### Pheophorbide
```{r}
boxplotPheophorbidePub <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = Pheophorbide, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Pheophorbide" ~ "(μg·L"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 10, 50, 150, 300, 500, 1000, 1500),
               labels = label_comma()) +
  
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotPheophorbidePub
```

### Shared Legend
```{r}
legendPlot <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = Pheophorbide, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Pheophorbide" ~ "(μg·L"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 10, 50, 150, 300, 500, 1000, 1500),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    )) +
  theme(
    legend.position = "bottom",
    legend.justification = "center",  # center the legend
    legend.box.just = "center")

legendPlot
```

### Combined Plot
```{r}
# Extract legend
sharedLegend <- cowplot::get_legend(legendPlot)

combinedPlot_AlgalProp <- cowplot::plot_grid(
  boxplotChlAPub, boxplotAstaxPub, boxplotLuteinPub,
  boxplotBetaCPub, boxplotPheophytinPub,
  boxplotPheophorbidePub,
  ncol = 3, nrow = 2, 
  rel_heights = c(1,1), 
  rel_widths = c(1,1),
  align = "hv",
  axis = "tblr",
  labels = NULL
)

combinedPlot_AlgalProp


ggsave(combinedPlot_AlgalProp, filename = "combinedPlot_AlgalProp.png", width = 7, height = 3.75, bg = "white")
```

### Cell Density formatting
```{r}
combinedPlot_CD <- cowplot::plot_grid(
  boxplotCellDensityPub,
  ncol = 3, nrow = 2, 
  rel_heights = c(1,1), 
  rel_widths = c(1,1),
  align = "hv",
  axis = "tblr",
  labels = NULL
)

combinedPlot_CD


ggsave(combinedPlot_CD, filename = "combinedPlot_CD.png", width = 7, height = 3.75, bg = "white")
```





## Algal Properties by Color {.tabset}

### Cell Density
```{r}
boxplotCellDensityPub_Color <- dfPigCell %>% 
  mutate(Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = CellDensity, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Cell Density" ~ "(cells·mL"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 3e+04, 2e+05, 5e+05, 1e+06, 2e+06, 4e+06)) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotCellDensityPub_Color
```

### Cell Volume
```{r}
boxplotCellVolumePub_Color <- dfPigCell %>% 
  filter(CellVolume > 0) %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = CellVolume, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Cell Volume" ~ "(μm"^{3}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 1000, 2500, 5000, 10000, 20000, 30000),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotCellVolumePub_Color
```

### Astaxanthin
```{r}
boxplotAstaxPub_Color <- dfPig %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = Astaxanthin, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Astaxanthin" ~ "(μg·L"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 10, 50, 100, 250, 500, 1000),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotAstaxPub_Color
```

### ChlA
```{r}
boxplotChlAPub_Color <- dfPig %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = ChlA, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Chlorophyll a" ~ "(μg·L"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 250, 1500, 5000, 10000, 20000, 30000),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotChlAPub_Color
```

### ChlB
```{r}
boxplotChlBPub_Color <- dfPig %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = ChlB, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Chlorophyll b" ~ "(μg·L"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 150, 1000, 2500, 5000, 10000),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotChlBPub_Color
```

### Lutein
```{r}
boxplotLuteinPub_Color <- dfPig %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = Lutein, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Lutein" ~ "(μg·L"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 150, 1000, 2500, 5000, 10000, 15000),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotLuteinPub_Color
```

### BetaC
```{r}
boxplotBetaCPub_Color <- dfPig %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = BetaC, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("β-carotene" ~ "(μg·L"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 150, 1000, 2500, 5000, 10000, 15000),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotBetaCPub_Color
```

### Pheophytin
```{r}
boxplotPheophytinPub_Color <- dfPig %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = Pheophytin, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Pheophytin" ~ "(μg·L"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 150, 1000, 2500, 5000, 7500, 12500),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotPheophytinPub_Color
```

### Pheophorbide
```{r}
boxplotPheophorbidePub_Color <- dfPig %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = Pheophorbide, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Pheophorbide" ~ "(μg·L"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 10, 50, 150, 300, 500, 1000, 1500),
               labels = label_comma()) +
  
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotPheophorbidePub_Color
```

### Combined Plot
```{r}
# Extract legend
sharedLegend <- cowplot::get_legend(legendPlot)

combinedPlot_AlgalProp_Color <- cowplot::plot_grid(
  boxplotChlAPub_Color, boxplotAstaxPub_Color, boxplotLuteinPub_Color,
  boxplotBetaCPub_Color, boxplotPheophytinPub_Color,
  boxplotPheophorbidePub_Color,
  ncol = 3, nrow = 2, 
  rel_heights = c(1,1), 
  rel_widths = c(1,1),
  align = "hv",
  axis = "tblr",
  labels = NULL
)

combinedPlot_AlgalProp_Color


ggsave(combinedPlot_AlgalProp_Color, filename = "combinedPlot_AlgalProp_Color.png", width = 7, height = 3.75, bg = "white")
```

### Cell Density formatting
```{r}
combinedPlot_CD_Color <- cowplot::plot_grid(
  boxplotCellDensityPub_Color,
  ncol = 3, nrow = 2, 
  rel_heights = c(1,1), 
  rel_widths = c(1,1),
  align = "hv",
  axis = "tblr",
  labels = NULL
)

combinedPlot_CD_Color


ggsave(combinedPlot_CD_Color, filename = "combinedPlot_CD_Color.png", width = 7, height = 3.75, bg = "white")
```

# ChlA:Astax Ratio Investigations

```{r}
heatmapPigRatioAnion
```

```{r}
dunnPlot_Ratio
```


It is interesting that from the heatmap above, correlations between PO43 and both ChlA:Astax and BetaC:Astax are strongly negative in green samples. I would suspect the opposite. Below, I do a little data visualization to see if I can better understand what is going on.

From the below plot, it looks like there is an inverse, 1/x relationship between ChlA:Astax and PO43 for green samples. However, there are also a lot of extreme outliers at low PO43 concentrations.

```{r}
dfAnion_Ratio %>% 
  filter(Color == "green") %>% 
  ggplot() + 
  geom_point(aes(x = PO43, y = ChlA_Astax_Ratio))
```

Spearman’s correlation between PO₄³⁻ and ChlA:Astax in green samples is strongly negative (ρ ≈ −0.57, p < 0.001). When extreme ratio values (>200) are removed, the correlation actually becomes slightly stronger (ρ ≈ −0.62), indicating that the relationship is not driven by a few high-ratio outliers.

The ratio distribution is highly right-skewed (median ≈ 23.7, max ≈ 1510), but the negative monotonic trend persists even when extreme values are excluded.

To understand this further, I checked correlations between PO₄³⁻ and the individual pigments in green samples:
	•	PO₄³⁻ vs Chl a: ρ ≈ 0.27 (weak positive)
	•	PO₄³⁻ vs Astaxanthin: ρ ≈ 0.76 (strong positive)

This explains the negative ratio relationship. Both pigments increase with phosphate, but astaxanthin increases much more strongly than chlorophyll a. Because the ratio is ChlA / Astaxanthin, a stronger increase in the denominator causes the ratio to decrease with increasing phosphate.

So this is not a statistical artifact — it reflects a proportional shift in pigment allocation with increasing phosphate in green snow samples.

```{r}
dfAnion_Ratio %>%
  filter(Color == "green") %>%
  summarise(
    rho = cor(PO43, ChlA_Astax_Ratio, method = "spearman"),
    p   = cor.test(PO43, ChlA_Astax_Ratio, method = "spearman")$p.value
  )

dfAnion_Ratio %>%
  filter(Color == "green", ChlA_Astax_Ratio < 200) %>%
  summarise(
    rho = cor(PO43, ChlA_Astax_Ratio, method = "spearman"),
    p   = cor.test(PO43, ChlA_Astax_Ratio, method = "spearman")$p.value
  )
```

```{r}
summary(dfAnion_Ratio$ChlA_Astax_Ratio[dfAnion_Ratio$Color == "green"])
```

```{r}
dfAnion %>%
  filter(Color == "green") %>%
  summarise(
    rho_chla  = cor(PO43, ChlA, method = "spearman"),
    rho_astax = cor(PO43, Astaxanthin, method = "spearman")
  )
```
