---
title: "Data Vizualization - Pigment Concentration (Antarctic Peninsula)"
author: "Ellie Ryan"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: flatly
    highlight: zenburn
    toc: true
    toc_float: true
    toc_collapsed: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Created:** 8/13/2025

**Last Modified:** 8/20/2025

```{r, message=FALSE, warning=FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(stringr)
library(kableExtra)
library(scales)
library(GGally)
library(patchwork)
library(Hmisc)
library(reshape2)
library(car)
library(FSA)
library(vegan)
```

```{r}

APPal <- c(
  "North" = "#364C56", 
  "South" = "#ABBF26"
)
locationPal <- c(
  "Lemon Creek Glacier" = "#e0b10d",
  "Camp Kiser" = "#4c7575",
  "Antarctic Peninsula" = "#a34108"
) 

locAbbrevPal <- c(
  "LC" = "#e0b10d",
  "CK" = "#4c7575",
  "AP" = "#a34108"
)

locationPal_APcolor <- c(
  "Lemon Creek Glacier" = "#e0b10d",
  "Camp Kiser" = "#4c7575",
  "AP - red" = "#933410", 
  "AP - green" = "#326e35", 
  "AP - clean" = "#1d1f42"
)

colorPal <- c(
  "clean" = "#769a9c",
  "red" = "#933410",
  "green" = "#326e35"
)

algaePal <- c(
  "clean" = "#a5bdc0",
  "algae" = "#735a69"
)

pigPal <- c(
  "Astaxanthin" = "#b43f1f",
  "Lutein" = "#f2bc1a",
  "ChlB" = "#439468",
  "ChlA" = "#37a133",
  "BetaC" = "#bf7915",
  "Pheophorbide" = "#534f8d", 
  "Pheophytin" = "#95879f"
)

plot_theme <- theme(
  axis.title = element_text(size = 12, face = "bold"), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  plot.margin = margin(20, 20, 20, 20)
)

plot_theme2 <- theme(
  axis.title = element_text(size = 12, face = "bold"), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  axis.text.x = element_text(angle = 45, hjust = 1), 
  plot.margin = margin(20, 20, 20, 20)
)

pairsplot_theme <- theme(
  text = element_text(size = 10),  # Adjust variable name text size
  axis.text = element_text(size = 8),  # Adjust axis text size
  plot.title = element_text(size = 12, face = "bold"),  # Adjust plot title size
  legend.text = element_text(size = 12))  # Adjust legend text size
```

# Overview

The purpose of this `.Rmd` is to visualize pigment concentration (ppb) by AP region and color, compare pigment concentration between sampling visits for site with repeat visits, and obtain basic summary statistics associated with each pigment. Before proceeding, I will clean the dataset so that it features only the variables I am interested in for this analysis. In this dataset, 241 snow samples have been collected between three locations: Lemon Creek Glacier, Camp Kiser, and the Antarctic Peninsula. These observations are broken down by `Color`, or the apparent snow color at time of sample collection. To keep things simple, there are only three levels to the `Color` category: red, green and clean. The breakdown of samples collected at each site is as follows:

  -   Antarctic Peninsula:
    -   113 samples
        -   45 red
        -   49 green
        -   19 clean
        
In this analysis, samples will also be divided by region in the Antarctic Peninsula. The breakdown is as follows: 

  -   North:
        -   Pendulum Cove, Deception Island
        -   Orne Harbour
        -   Danco Island
        -   Neko Harbour
        -   Damoy Point
        -   Pléneau Island
        -   Petermann Island
        
  -   South:
        -   Stonington Island
        -   Horseshoe Island

        
*NOTE: Fewer samples were used for analysis, since there are NA values for pigment concentrations of some samples.*
        
From pigment outlier analysis, the following outliers were identified for removal:

-   2025-02-19-SI-algae3
-   2025-02-24-DA-algae2
-   2025-02-24-DA-algae6
-   2025-03-07-NH-algae3
-   2025-02-24-DA-algae1
-   2025-01-12-PI-algae10


See the Preliminary Results and Figures doc for justification for removal of each. 

## Data Wrangling

```{r}
df <- read.csv("CompiledDataPI.csv")

df <- df %>% 
  filter(Location == "Antarctic Peninsula")
```

Adding an `Algae` column to classify samples as algae or clean snow samples and a `LocColor` column to classify AP algae samples as red or green, and an `APLoc` column to separate samples by region. 

```{r}
df <- df %>% 
  mutate(Site = dplyr::recode(Site, "Pendulum Cove, Deception Island" = "Pendulum Cove"))

# Format date column 
df <- df %>%
  mutate(
    Date = as.Date(Date, format = "%m/%d/%Y"))

# Add LocColor
df <- df %>% mutate(LocColor = case_when(
    Location == "Antarctic Peninsula" ~ paste("AP -", Color),
    TRUE ~ Location
  )) %>% 
  relocate(LocColor, .after = Site) %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  relocate(Algae, .after = Color)

# Add LocAbbrev
df <- df %>% mutate(LocAbbrev = case_when(
  Location == "Antarctic Peninsula" ~ paste("AP"),
  Location == "Lemon Creek Glacier" ~ paste("LC"),
  Location == "Camp Kiser" ~ paste("CK")
)) %>%  
  relocate(LocAbbrev, .after = LocColor)

# Add APLoc
df <- df %>% mutate(APLoc = case_when(
  Site %in% c("Pendulum Cove",
              "Orne Harbour",
              "Danco Island",
              "Neko Harbour",
              "Damoy Point",
              "Pléneau Island",
              "Petermann Island") ~ paste("North"),
  Site %in% c("Stonington Island",
              "Horseshoe Island")~ paste("South")
)) %>% 
  relocate(APLoc, .after = Site)
```

```{r}
str(df)
```

```{r}
outliers <- c(
  "2025-02-19-SI-algae3",
  "2025-02-24-DA-algae2",
  "2025-02-24-DA-algae6",
  "2025-03-07-NH-algae3",
  "2023-08-24-LC-algae33",
  "2025-01-12-PI-algae10")

df <- df %>% 
  filter(!SampleID %in% outliers)
```

### Pigments
```{r}
dfPig <- df %>% 
  select(c("SampleID", "Date", "Location", "Latitude", "Site", "APLoc", "LocColor", "LocAbbrev", "Color", "Algae", Astaxanthin:Pheophytin)) %>% 
  na.omit(dfPig)
```

### Pigments & Anions 
```{r}
dfAnion <- df %>% 
  select(c("SampleID", "Date", "Location", "Latitude", "Site", "APLoc", "LocColor", "LocAbbrev", "Color", "Algae", Astaxanthin:Pheophytin, Fo:SO42)) %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  mutate(NO2NO3 = NO2 + NO3) %>% 
  relocate(Algae, .after = Color) %>% 
  relocate(NO2NO3, .after = Cl) %>% 
  select(!NO2) %>% 
  select(!NO3) %>% 
  na.omit(dfAnion)

outliersAnion <- c("2025-02-24-DA-algae1")

dfAnion <- dfAnion %>% 
  filter(!SampleID %in% outliersAnion)

```

### Pigments (Pivoted)
```{r}
dfPig_long <- dfPig %>% 
  pivot_longer(cols = Astaxanthin:Pheophytin,
               names_to = "Pigment",
               values_to = "Concentration")
```

### Anions (Pivoted)
```{r}
dfAnion_long <- dfAnion %>% 
  pivot_longer(cols = c(Fo, Cl, SO42, PO43, NO2NO3),
               names_to = "Anion",
               values_to = "Concentration")
```


# Summary Tables {.tabset}

## Primary Pigments
```{r}
# Pigment Summary (ChlA and ChlB)
ChlSummary <- dfPig %>% 
  group_by(APLoc, Color) %>% 
  summarise(
    n = n(),
    ChlA_mean = comma(round(mean(ChlA, na.rm = TRUE), 2)),
    ChlA_median = comma(round(median(ChlA, na.rm = TRUE), 2)),
    ChlA_max = comma(round(max(ChlA, na.rm = TRUE), 2)),
    ChlA_min = comma(round(min(ChlA, na.rm = TRUE), 2)),
    ChlA_sd = comma(round(sd(ChlA, na.rm = TRUE), 2)),
    ChlB_mean = comma(round(mean(ChlB, na.rm = TRUE), 2)),
    ChlB_median = comma(round(median(ChlB, na.rm = TRUE), 2)),
    ChlB_max = comma(round(max(ChlB, na.rm = TRUE), 2)),
    ChlB_min = comma(round(min(ChlB, na.rm = TRUE), 2)),
    ChlB_sd = comma(round(sd(ChlB, na.rm = TRUE), 2)),
    .groups = "drop"
  )

# Add "all" for each site
ChlSummaryAll <- dfPig %>% 
  group_by(APLoc) %>% 
  summarise(
    Color = "all",
    n = n(),
    ChlA_mean = comma(round(mean(ChlA, na.rm = TRUE), 2)),
    ChlA_median = comma(round(median(ChlA, na.rm = TRUE), 2)),
    ChlA_max = comma(round(max(ChlA, na.rm = TRUE), 2)),
    ChlA_min = comma(round(min(ChlA, na.rm = TRUE), 2)),
    ChlA_sd = comma(round(sd(ChlA, na.rm = TRUE), 2)),
    ChlB_mean = comma(round(mean(ChlB, na.rm = TRUE), 2)),
    ChlB_median = comma(round(median(ChlB, na.rm = TRUE), 2)),
    ChlB_max = comma(round(max(ChlB, na.rm = TRUE), 2)),
    ChlB_min = comma(round(min(ChlB, na.rm = TRUE), 2)),
    ChlB_sd = comma(round(sd(ChlB, na.rm = TRUE), 2)),
    .groups = "drop"
  )

# Combine and order rows
ChlSummary <- bind_rows(ChlSummaryAll, ChlSummary) %>% 
  arrange(APLoc, match(Color, c("green", "red", "clean", "all")))

# Format table with kableExtra
ChlSummary %>% 
  kable("html", caption = "Summary Statistics for ChlA and ChlB by Snow Color and AP Region (ppb)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Carotenoids
```{r}
# Pigment Summary (Astaxanthin, Beta Carotene and Lutein)
CarotenoidSummary <- dfPig %>% 
  group_by(APLoc, Color) %>% 
  summarise(
    n = n(),
    Astaxanthin_mean = comma(round(mean(Astaxanthin, na.rm = TRUE), 2)),
    Astaxanthin_median = comma(round(median(Astaxanthin, na.rm = TRUE), 2)),
    Astaxanthin_max = comma(round(max(Astaxanthin, na.rm = TRUE), 2)),
    Astaxanthin_min = comma(round(min(Astaxanthin, na.rm = TRUE), 2)),
    Astaxanthin_sd = comma(round(sd(Astaxanthin, na.rm = TRUE), 2)),
    Lutein_mean = comma(round(mean(Lutein, na.rm = TRUE), 2)),
    Lutein_median = comma(round(median(Lutein, na.rm = TRUE), 2)),
    Lutein_max = comma(round(max(Lutein, na.rm = TRUE), 2)),
    Lutein_min = comma(round(min(Lutein, na.rm = TRUE), 2)),
    Lutein_sd = comma(round(sd(Lutein, na.rm = TRUE), 2)),
    BetaC_mean = comma(round(mean(BetaC, na.rm = TRUE), 2)),
    BetaC_median = comma(round(median(BetaC, na.rm = TRUE), 2)),
    BetaC_max = comma(round(max(BetaC, na.rm = TRUE), 2)),
    BetaC_min = comma(round(min(BetaC, na.rm = TRUE), 2)),
    BetaC_sd = comma(round(sd(BetaC, na.rm = TRUE), 2)),
    .groups = "drop"
  )

# Add "all" for each site
CarotenoidSummaryAll <- dfPig %>% 
  group_by(APLoc) %>% 
  summarise(
    Color = "all",
    n = n(),
    Astaxanthin_mean = comma(round(mean(Astaxanthin, na.rm = TRUE), 2)),
    Astaxanthin_median = comma(round(median(Astaxanthin, na.rm = TRUE), 2)),
    Astaxanthin_max = comma(round(max(Astaxanthin, na.rm = TRUE), 2)),
    Astaxanthin_min = comma(round(min(Astaxanthin, na.rm = TRUE), 2)),
    Astaxanthin_sd = comma(round(sd(Astaxanthin, na.rm = TRUE), 2)),
    Lutein_mean = comma(round(mean(Lutein, na.rm = TRUE), 2)),
    Lutein_median = comma(round(median(Lutein, na.rm = TRUE), 2)),
    Lutein_max = comma(round(max(Lutein, na.rm = TRUE), 2)),
    Lutein_min = comma(round(min(Lutein, na.rm = TRUE), 2)),
    Lutein_sd = comma(round(sd(Lutein, na.rm = TRUE), 2)),
    BetaC_mean = comma(round(mean(BetaC, na.rm = TRUE), 2)),
    BetaC_median = comma(round(median(BetaC, na.rm = TRUE), 2)),
    BetaC_max = comma(round(max(BetaC, na.rm = TRUE), 2)),
    BetaC_min = comma(round(min(BetaC, na.rm = TRUE), 2)),
    BetaC_sd = comma(round(sd(BetaC, na.rm = TRUE), 2)),
    .groups = "drop"
  )


# Combine and order rows
CarotenoidSummary <- bind_rows(CarotenoidSummaryAll, CarotenoidSummary) %>% 
  arrange(APLoc, match(Color, c("green", "red", "clean", "all")))

# Format table with kableExtra
CarotenoidSummary %>% 
  kable("html", caption = "Summary Statistics for Astaxanthin and Lutein by Snow Color and AP Region (ppb)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```


## Degradation Products 
```{r}
# Pigment Summary (Pheophytin and Pheophorbide)
DegSummary <- dfPig %>% 
  group_by(APLoc, Color) %>% 
  summarise(
    n = n(),
    Pheophytin_mean = comma(round(mean(Pheophytin, na.rm = TRUE), 2)),
    Pheophytin_median = comma(round(median(Pheophytin, na.rm = TRUE), 2)),
    Pheophytin_max = comma(round(max(Pheophytin, na.rm = TRUE), 2)),
    Pheophytin_min = comma(round(min(Pheophytin, na.rm = TRUE), 2)),
    Pheophytin_sd = comma(round(sd(Pheophytin, na.rm = TRUE), 2)),
    Pheophorbide_mean = comma(round(mean(Pheophorbide, na.rm = TRUE), 2)),
    Pheophorbide_median = comma(round(median(Pheophorbide, na.rm = TRUE), 2)),
    Pheophorbide_max = comma(round(max(Pheophorbide, na.rm = TRUE), 2)),
    Pheophorbide_min = comma(round(min(Pheophorbide, na.rm = TRUE), 2)),
    Pheophorbide_sd = comma(round(sd(Pheophorbide, na.rm = TRUE), 2)),
    .groups = "drop"
  )

# Add "all" for each site
DegSummaryAll <- dfPig %>% 
  group_by(APLoc) %>% 
  summarise(
    Color = "all",
    n = n(),
    Pheophytin_mean = comma(round(mean(Pheophytin, na.rm = TRUE), 2)),
    Pheophytin_median = comma(round(median(Pheophytin, na.rm = TRUE), 2)),
    Pheophytin_max = comma(round(max(Pheophytin, na.rm = TRUE), 2)),
    Pheophytin_min = comma(round(min(Pheophytin, na.rm = TRUE), 2)),
    Pheophytin_sd = comma(round(sd(Pheophytin, na.rm = TRUE), 2)),
    Pheophorbide_mean = comma(round(mean(Pheophorbide, na.rm = TRUE), 2)),
    Pheophorbide_median = comma(round(median(Pheophorbide, na.rm = TRUE), 2)),
    Pheophorbide_max = comma(round(max(Pheophorbide, na.rm = TRUE), 2)),
    Pheophorbide_min = comma(round(min(Pheophorbide, na.rm = TRUE), 2)),
    Pheophorbide_sd = comma(round(sd(Pheophorbide, na.rm = TRUE), 2)),
    .groups = "drop"
  )

# Combine and order rows
DegSummary <- bind_rows(DegSummaryAll, DegSummary) %>% 
  arrange(APLoc, match(Color, c("green", "red", "clean", "all")))

# Format table with kableExtra
DegSummary %>% 
  kable("html", caption = "Summary Statistics for Pheophytin and Pheophorbide by Snow Color and AP Region (ppb)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Repeat Visits {.tabset}

### Petermann Island
```{r}
PISummary <- dfPig %>% 
  filter(Site == "Petermann Island") %>% 
  group_by(Date, Color) %>% 
  summarise(
    n = n(),
    ChlA_median = comma(round(median(ChlA, na.rm = TRUE), 2)),
    ChlB_median = comma(round(median(ChlB, na.rm = TRUE), 2)),
    Astax_median = comma(round(median(Astaxanthin, na.rm = TRUE), 2)),
    Lutein_median = comma(round(median(Lutein, na.rm = TRUE), 2)),
    BetaC_median = comma(round(median(BetaC, na.rm = TRUE), 2)),
    Pheophytin_median = comma(round(median(Pheophytin, na.rm = TRUE), 2)),
    Pheophorbide_median = comma(round(median(Pheophorbide, na.rm = TRUE), 2)),
    .groups = "drop"
  )

PISummaryAll <- dfPig %>% 
  filter(Site == "Petermann Island") %>% 
  group_by(Date) %>% 
  summarise(
    Color = "all",
    n = n(),
    ChlA_median = comma(round(median(ChlA, na.rm = TRUE), 2)),
    ChlB_median = comma(round(median(ChlB, na.rm = TRUE), 2)),
    Astax_median = comma(round(median(Astaxanthin, na.rm = TRUE), 2)),
    Lutein_median = comma(round(median(Lutein, na.rm = TRUE), 2)),
    BetaC_median = comma(round(median(BetaC, na.rm = TRUE), 2)),
    Pheophytin_median = comma(round(median(Pheophytin, na.rm = TRUE), 2)),
    Pheophorbide_median = comma(round(median(Pheophorbide, na.rm = TRUE), 2)),
    .groups = "drop"
  )

# Combine and order rows
PISummary <- bind_rows(PISummaryAll, PISummary) %>% 
  arrange(Date, match(Color, c("green", "red", "clean", "all")))

# Format table with kableExtra
PISummary %>% 
  kable("html", caption = "Summary Statistics for Petermann Island by Snow Color and Sampling Date (ppm)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

### Damoy Point
```{r}
PISummary <- dfPig %>% 
  filter(Site == "Damoy Point") %>% 
  group_by(Date, Color) %>% 
  summarise(
    n = n(),
    ChlA_median = comma(round(median(ChlA, na.rm = TRUE), 2)),
    ChlB_median = comma(round(median(ChlB, na.rm = TRUE), 2)),
    Astax_median = comma(round(median(Astaxanthin, na.rm = TRUE), 2)),
    Lutein_median = comma(round(median(Lutein, na.rm = TRUE), 2)),
    BetaC_median = comma(round(median(BetaC, na.rm = TRUE), 2)),
    Pheophytin_median = comma(round(median(Pheophytin, na.rm = TRUE), 2)),
    Pheophorbide_median = comma(round(median(Pheophorbide, na.rm = TRUE), 2)),
    .groups = "drop"
  )

PISummaryAll <- dfPig %>% 
  filter(Site == "Damoy Point") %>% 
  group_by(Date) %>% 
  summarise(
    Color = "all",
    n = n(),
    ChlA_median = comma(round(median(ChlA, na.rm = TRUE), 2)),
    ChlB_median = comma(round(median(ChlB, na.rm = TRUE), 2)),
    Astax_median = comma(round(median(Astaxanthin, na.rm = TRUE), 2)),
    Lutein_median = comma(round(median(Lutein, na.rm = TRUE), 2)),
    BetaC_median = comma(round(median(BetaC, na.rm = TRUE), 2)),
    Pheophytin_median = comma(round(median(Pheophytin, na.rm = TRUE), 2)),
    Pheophorbide_median = comma(round(median(Pheophorbide, na.rm = TRUE), 2)),
    .groups = "drop"
  )

# Combine and order rows
PISummary <- bind_rows(PISummaryAll, PISummary) %>% 
  arrange(Date, match(Color, c("green", "red", "clean", "all")))

# Format table with kableExtra
PISummary %>% 
  kable("html", caption = "Summary Statistics for Damoy Point by Snow Color and Sampling Date (ppm)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

### Orne Harbour
```{r}
PISummary <- dfPig %>% 
  filter(Site == "Orne Harbour") %>% 
  group_by(Date, Color) %>% 
  summarise(
    n = n(),
    ChlA_median = comma(round(median(ChlA, na.rm = TRUE), 2)),
    ChlB_median = comma(round(median(ChlB, na.rm = TRUE), 2)),
    Astax_median = comma(round(median(Astaxanthin, na.rm = TRUE), 2)),
    Lutein_median = comma(round(median(Lutein, na.rm = TRUE), 2)),
    BetaC_median = comma(round(median(BetaC, na.rm = TRUE), 2)),
    Pheophytin_median = comma(round(median(Pheophytin, na.rm = TRUE), 2)),
    Pheophorbide_median = comma(round(median(Pheophorbide, na.rm = TRUE), 2)),
    .groups = "drop"
  )

PISummaryAll <- dfPig %>% 
  filter(Site == "Orne Harbour") %>% 
  group_by(Date) %>% 
  summarise(
    Color = "all",
    n = n(),
    ChlA_median = comma(round(median(ChlA, na.rm = TRUE), 2)),
    ChlB_median = comma(round(median(ChlB, na.rm = TRUE), 2)),
    Astax_median = comma(round(median(Astaxanthin, na.rm = TRUE), 2)),
    Lutein_median = comma(round(median(Lutein, na.rm = TRUE), 2)),
    BetaC_median = comma(round(median(BetaC, na.rm = TRUE), 2)),
    Pheophytin_median = comma(round(median(Pheophytin, na.rm = TRUE), 2)),
    Pheophorbide_median = comma(round(median(Pheophorbide, na.rm = TRUE), 2)),
    .groups = "drop"
  )

# Combine and order rows
PISummary <- bind_rows(PISummaryAll, PISummary) %>% 
  arrange(Date, match(Color, c("green", "red", "clean", "all")))

# Format table with kableExtra
PISummary %>% 
  kable("html", caption = "Summary Statistics for Orne Harbour by Snow Color and Sampling Date (ppm)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

### Pléneau Island
```{r}
PISummary <- dfPig %>% 
  filter(Site == "Pléneau Island") %>% 
  group_by(Date, Color) %>% 
  summarise(
    n = n(),
    ChlA_median = comma(round(median(ChlA, na.rm = TRUE), 2)),
    ChlB_median = comma(round(median(ChlB, na.rm = TRUE), 2)),
    Astax_median = comma(round(median(Astaxanthin, na.rm = TRUE), 2)),
    Lutein_median = comma(round(median(Lutein, na.rm = TRUE), 2)),
    BetaC_median = comma(round(median(BetaC, na.rm = TRUE), 2)),
    Pheophytin_median = comma(round(median(Pheophytin, na.rm = TRUE), 2)),
    Pheophorbide_median = comma(round(median(Pheophorbide, na.rm = TRUE), 2)),
    .groups = "drop"
  )

PISummaryAll <- dfPig %>% 
  filter(Site == "Pléneau Island") %>% 
  group_by(Date) %>% 
  summarise(
    Color = "all",
    n = n(),
    ChlA_median = comma(round(median(ChlA, na.rm = TRUE), 2)),
    ChlB_median = comma(round(median(ChlB, na.rm = TRUE), 2)),
    Astax_median = comma(round(median(Astaxanthin, na.rm = TRUE), 2)),
    Lutein_median = comma(round(median(Lutein, na.rm = TRUE), 2)),
    BetaC_median = comma(round(median(BetaC, na.rm = TRUE), 2)),
    Pheophytin_median = comma(round(median(Pheophytin, na.rm = TRUE), 2)),
    Pheophorbide_median = comma(round(median(Pheophorbide, na.rm = TRUE), 2)),
    .groups = "drop"
  )

# Combine and order rows
PISummary <- bind_rows(PISummaryAll, PISummary) %>% 
  arrange(Date, match(Color, c("green", "red", "clean", "all")))

# Format table with kableExtra
PISummary %>% 
  kable("html", caption = "Summary Statistics for Pléneau Island by Snow Color and Sampling Date (ppm)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

# ANOVA/Kruskal Wallis Test
Before proceeding with an ANOVA test to assess differences in pigment concentration between locations, I am going to check to see if these variables follow the assumptions of normality and equal variance required to run an ANOVA: 
```{r}
# Step 1: Group by pigment and nest data
nested <- dfPig_long %>%
  group_by(Pigment) %>%
  nest()

# Step 2: Fit ANOVA model, extract residuals, and run tests
anovaAssumptionsResults <- nested %>%
  mutate(
    # Fit ANOVA model
    aov_model = map(data, ~ aov(Concentration ~ APLoc, data = .x)),
    
    # Shapiro-Wilk test for normality of residuals
    shapiro_p = map_dbl(aov_model, ~ shapiro.test(residuals(.x))$p.value),
    
    # Levene's Test for homogeneity of variance
    levene_p = map_dbl(data, ~leveneTest(Concentration ~ factor(APLoc), data = .x)$`Pr(>F)`[1])
  )
```

From the below results, it appears that none of the pigments are both normally distributed and have equal variances. Because of this, I am going to proceed with a Kruskal-Wallis test in lieu of an ANOVA. 

```{r}
anovaAssumptionsResults %>%
  select(Pigment, shapiro_p, levene_p) %>%
  mutate(
    normality_ok = shapiro_p > 0.05,
    homogeneity_ok = levene_p > 0.05
  )
```

## Kruskal-Wallis Test {.tabset}
Below, I use the Kruskal-Wallis test to run a non-parametric comparison of each pigment across each location. Then, I follow up with a Dunn's post hoc test for the pigments that show significant differences between the locations. I also look at differences in pigment concentration between different snow colors within each site. 

### All Sites (Algae Samples)
```{r}
kruskalResults <- dfPig_long %>%
  group_by(Pigment) %>%
  filter(Algae == "algae") %>% 
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ APLoc, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )


# Step 2: Run Dunn's test
kruskalResults <- kruskalResults %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(APLoc), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

#### Visualize Results
```{r}
dunn_pairs <- kruskalResults %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_Algae <- ggplot(dunn_pairs, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunn_pairs, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_Algae

ggsave(dunnPlot_Algae, filename = "dunnPlot_Algae.png", width = 4, height = 6, bg = "white")
```

### All Sites (Red Samples)
```{r}
kruskalResultsR <- dfPig_long %>%
  group_by(Pigment) %>%
  filter(Color == "red") %>% 
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ APLoc, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )


# Step 2: Run Dunn's test
kruskalResultsR <- kruskalResultsR %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(APLoc), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

#### Visualize Results
```{r}
dunn_pairsR <- kruskalResultsR %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_R <- ggplot(dunn_pairsR, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunn_pairsR, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_R

ggsave(dunnPlot_R, filename = "dunnPlot_R.png", width = 4, height = 6, bg = "white")
```

### All Sites (Green Samples)
```{r}
kruskalResultsG <- dfPig_long %>%
  group_by(Pigment) %>%
  filter(Color == "green") %>% 
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ APLoc, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )


# Step 2: Run Dunn's test
kruskalResultsG <- kruskalResultsG %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(APLoc), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

#### Visualize Results
```{r}
dunn_pairsG <- kruskalResultsG %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_G <- ggplot(dunn_pairsG, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunn_pairsG, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_G

ggsave(dunnPlot_G, filename = "dunnPlot_G.png", width = 4, height = 6, bg = "white")
```

### All Sites (Clean Samples)
```{r}
kruskalResultsC <- dfPig_long %>%
  group_by(Pigment) %>%
  filter(Color == "clean") %>% 
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ APLoc, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )


# Step 2: Run Dunn's test
kruskalResultsC <- kruskalResultsC %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(APLoc), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

#### Visualize Results
```{r}
dunn_pairsC <- kruskalResultsC %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_C <- ggplot(dunn_pairsC, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunn_pairsC, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_C

ggsave(dunnPlot_C, filename = "dunnPlot_C.png", width = 4, height = 6, bg = "white")
```

### Petermann Island {.tabset}

#### All Samples
```{r}
# Run kruskal wallis test 
kruskalResultsPI <- dfPig_long %>%
  filter(Site == "Petermann Island") %>% 
  group_by(Pigment) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ Date, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsPI <- kruskalResultsPI %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(Date), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

##### Visualize Results

```{r}
dunnPairsPI <- kruskalResultsPI %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_PI <- ggplot(dunnPairsPI, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsPI, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_PI

ggsave(dunnPlot_PI, filename = "dunnPlot_PI.png", width = 3, height = 6, bg = "white")
```

#### Red Samples

```{r}
# Run kruskal wallis test 
kruskalResultsPI_Red <- dfPig_long %>%
  filter(Site == "Petermann Island") %>% 
  filter(Color == "red") %>% 
  group_by(Pigment) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ Date, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsPI_Red <- kruskalResultsPI_Red %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(Date), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

##### Visualize Results

```{r}
dunnPairsPI_Red <- kruskalResultsPI_Red %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_PI_Red <- ggplot(dunnPairsPI_Red, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsPI_Red, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_PI_Red

ggsave(dunnPlot_PI_Red, filename = "dunnPlot_PI_Red.png", width = 3, height = 6, bg = "white")
```

#### Green Samples

```{r}
# Run kruskal wallis test 
kruskalResultsPI_Green <- dfPig_long %>%
  filter(Site == "Petermann Island") %>% 
  filter(Color == "green") %>% 
  group_by(Pigment) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ Date, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsPI_Green <- kruskalResultsPI_Green %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(Date), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

##### Visualize Results

```{r}
dunnPairsPI_Green <- kruskalResultsPI_Green %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_PI_Green <- ggplot(dunnPairsPI_Green, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsPI_Green, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_PI_Green

ggsave(dunnPlot_PI_Green, filename = "dunnPlot_PI_Green.png", width = 3, height = 6, bg = "white")
```

### Damoy Point {.tabset}

#### All Samples

```{r}
# Run kruskal wallis test 
kruskalResultsDP <- dfPig_long %>%
  filter(Site == "Damoy Point") %>% 
  group_by(Pigment) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ Date, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsDP <- kruskalResultsDP %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(Date), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

##### Visualize Results

```{r}
dunnPairsDP <- kruskalResultsDP %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_DP <- ggplot(dunnPairsDP, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsDP, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_DP

ggsave(dunnPlot_DP, filename = "dunnPlot_DP.png", width = 5, height = 6, bg = "white")
```

#### Red Samples

```{r}
# Run kruskal wallis test 
kruskalResultsDP_Red <- dfPig_long %>%
  filter(Site == "Damoy Point") %>% 
  filter(Color == "red") %>% 
  group_by(Pigment) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ Date, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsDP_Red <- kruskalResultsDP_Red %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(Date), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

##### Visualize Results

```{r}
dunnPairsDP_Red <- kruskalResultsDP_Red %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_DP_Red <- ggplot(dunnPairsDP_Red, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsDP_Red, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_DP_Red

ggsave(dunnPlot_DP_Red, filename = "dunnPlot_DP_Red.png", width = 5, height = 6, bg = "white")
```

#### Green Samples

```{r}
# Run kruskal wallis test 
kruskalResultsDP_Green <- dfPig_long %>%
  filter(Site == "Damoy Point") %>% 
  filter(Color == "green") %>% 
  group_by(Pigment) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ Date, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsDP_Green <- kruskalResultsDP_Green %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(Date), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

##### Visualize Results

```{r}
dunnPairsDP_Green <- kruskalResultsDP_Green %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_DP_Green <- ggplot(dunnPairsDP_Green, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsDP_Green, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_DP_Green

ggsave(dunnPlot_DP_Green, filename = "dunnPlot_DP_Green.png", width = 5, height = 6, bg = "white")
```

### Orne Harbour {.tabset}

#### All Samples

```{r}
# Run kruskal wallis test 
kruskalResultsOH <- dfPig_long %>%
  filter(Site == "Orne Harbour") %>% 
  group_by(Pigment) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ Date, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsOH <- kruskalResultsOH %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(Date), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

##### Visualize Results

```{r}
dunnPairsOH <- kruskalResultsOH %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_OH <- ggplot(dunnPairsOH, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsOH, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_OH

ggsave(dunnPlot_OH, filename = "dunnPlot_OH.png", width = 3, height = 6, bg = "white")
```

#### Red Samples

```{r}
# Run kruskal wallis test 
kruskalResultsOH_Red <- dfPig_long %>%
  filter(Site == "Orne Harbour") %>% 
  filter(Color == "red") %>% 
  group_by(Pigment) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ Date, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsOH_Red <- kruskalResultsOH_Red %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(Date), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

##### Visualize Results

```{r}
dunnPairsOH_Red <- kruskalResultsOH_Red %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_OH_Red <- ggplot(dunnPairsOH_Red, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsOH_Red, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_OH_Red

ggsave(dunnPlot_OH_Red, filename = "dunnPlot_OH_Red.png", width = 3, height = 6, bg = "white")
```

#### Green Samples

```{r}
# Run kruskal wallis test 
kruskalResultsOH_Green <- dfPig_long %>%
  filter(Site == "Orne Harbour") %>% 
  filter(Color == "green") %>% 
  group_by(Pigment) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ Date, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsOH_Green <- kruskalResultsOH_Green %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(Date), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

##### Visualize Results

```{r}
dunnPairsOH_Green <- kruskalResultsOH_Green %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_OH_Green <- ggplot(dunnPairsOH_Green, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsOH_Green, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_OH_Green

ggsave(dunnPlot_OH_Green, filename = "dunnPlot_OH_Green.png", width = 3, height = 6, bg = "white")
```

### Pléneau Island {.tabset}

#### All Samples

```{r}
# Run kruskal wallis test 
kruskalResultsPLE <- dfPig_long %>%
  filter(Site == "Pléneau Island") %>% 
  group_by(Pigment) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ Date, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsPLE <- kruskalResultsPLE %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(Date), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

##### Visualize Results

```{r}
dunnPairsPLE <- kruskalResultsPLE %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_PLE <- ggplot(dunnPairsPLE, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsPLE, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_PLE

ggsave(dunnPlot_PLE, filename = "dunnPlot_PLE.png", width = 3, height = 6, bg = "white")
```

#### Red Samples

```{r}
# Run kruskal wallis test 
kruskalResultsPLE_Red <- dfPig_long %>%
  filter(Site == "Pléneau Island") %>% 
  filter(Color == "red") %>% 
  group_by(Pigment) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ Date, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsPLE_Red <- kruskalResultsPLE_Red %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(Date), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

##### Visualize Results

```{r}
dunnPairsPLE_Red <- kruskalResultsPLE_Red %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_PLE_Red <- ggplot(dunnPairsPLE_Red, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsPLE_Red, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_PLE_Red

ggsave(dunnPlot_PLE_Red, filename = "dunnPlot_PLE_Red.png", width = 3, height = 6, bg = "white")
```

#### Green Samples

```{r}
# Run kruskal wallis test 
kruskalResultsPLE_Green <- dfPig_long %>%
  filter(Site == "Pléneau Island") %>% 
  filter(Color == "green") %>% 
  group_by(Pigment) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ Date, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsPLE_Green <- kruskalResultsPLE_Green %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(Date), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

##### Visualize Results

```{r}
dunnPairsPLE_Green <- kruskalResultsPLE_Green %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_PLE_Green <- ggplot(dunnPairsPLE_Green, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsPLE_Green, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_PLE_Green

ggsave(dunnPlot_PLE_Green, filename = "dunnPlot_PLE_Green.png", width = 3, height = 6, bg = "white")
```

# Box Plots {.tabset}

## Algae Pigments (All Samples)
```{r}
boxplotAll <- dfPig_long %>%
  ggplot(aes(x = Pigment, y = Concentration, fill = APLoc, color = APLoc)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = APPal) +
  scale_color_manual(values = APPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotAll

ggsave(boxplotAll, filename = "boxplotAll.png", width = 7, height = 5, bg = "white")
```

## All Pigments (Algae Samples)
```{r}
boxplotAlgae <- dfPig_long %>%
  filter(Algae == "algae") %>% 
  ggplot(aes(x = Pigment, y = Concentration, fill = APLoc, color = APLoc)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = APPal) +
  scale_color_manual(values = APPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotAlgae

ggsave(boxplotAlgae, filename = "boxplotAlgae.png", width = 7, height = 5, bg = "white")
```

## All Pigments (Red Samples)
```{r}
boxplotAlgaeR <- dfPig_long %>%
  filter(Color == "red") %>% 
  ggplot(aes(x = Pigment, y = Concentration, fill = APLoc, color = APLoc)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = APPal) +
  scale_color_manual(values = APPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotAlgaeR

ggsave(boxplotAlgaeR, filename = "boxplotAlgaeR.png", width = 7, height = 5, bg = "white")
```

## All Pigments (Green Samples)
```{r}
boxplotAlgaeG <- dfPig_long %>%
  filter(Color == "green") %>% 
  ggplot(aes(x = Pigment, y = Concentration, fill = APLoc, color = APLoc)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = APPal) +
  scale_color_manual(values = APPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotAlgaeG

ggsave(boxplotAlgaeG, filename = "boxplotAlgaeG.png", width = 7, height = 5, bg = "white")
```

## Astaxanthin {.tabset}
```{r}
boxplotAstax <- dfPig %>%
  ggplot(aes(x = factor(APLoc, levels = c("North", "South")), 
             y = Astaxanthin, 
             fill = Color, 
             color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotAstax

ggsave(boxplotAstax, filename = "boxplotAstax.png", width = 7, height = 5, bg = "white")
```

```{r}
boxplotAstax2 <- dfPig %>%
  ggplot(aes(x = fct_reorder(Site, Latitude, .desc = TRUE), 
               y = Astaxanthin, 
               fill = Color, 
               color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotAstax2

ggsave(boxplotAstax2, filename = "boxplotAstax2.png", width = 7, height = 5, bg = "white")
```

## Beta Carotene
```{r}
boxplotBetaC <- dfPig %>%
  ggplot(aes(x = factor(APLoc, levels = c("North", "South")), 
             y = BetaC, 
             fill = Color, 
             color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotBetaC

ggsave(boxplotBetaC, filename = "boxplotBetaC.png", width = 7, height = 5, bg = "white")
```

```{r}
boxplotBetaC2 <- dfPig %>%
  ggplot(aes(x = fct_reorder(Site, Latitude, .desc = TRUE), 
               y = BetaC, 
               fill = Color, 
               color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotBetaC2

ggsave(boxplotBetaC2, filename = "boxplotBetaC2.png", width = 7, height = 5, bg = "white")
```

## ChlA
```{r}
boxplotChlA <- dfPig %>%
  ggplot(aes(x = factor(APLoc, levels = c("North", "South")), 
             y = ChlA, 
             fill = Color, 
             color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotChlA

ggsave(boxplotChlA, filename = "boxplotChlA.png", width = 7, height = 5, bg = "white")
```

```{r}
boxplotChlA2 <- dfPig %>%
  ggplot(aes(x = fct_reorder(Site, Latitude, .desc = TRUE), 
               y = ChlA, 
               fill = Color, 
               color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotChlA2

ggsave(boxplotChlA2, filename = "boxplotChlA2.png", width = 7, height = 5, bg = "white")
```

## ChlB
```{r}
boxplotChlB <- dfPig %>%
  ggplot(aes(x = factor(APLoc, levels = c("North", "South")), 
             y = ChlB, 
             fill = Color, 
             color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotChlB

ggsave(boxplotChlB, filename = "boxplotChlB.png", width = 7, height = 5, bg = "white")
```

```{r}
boxplotChlB2 <- dfPig %>%
  ggplot(aes(x = fct_reorder(Site, Latitude, .desc = TRUE), 
               y = ChlB, 
               fill = Color, 
               color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotChlB2

ggsave(boxplotChlB2, filename = "boxplotChlB2.png", width = 7, height = 5, bg = "white")
```

## Lutein
```{r}
boxplotLutein <- dfPig %>%
  ggplot(aes(x = factor(APLoc, levels = c("North", "South")), 
             y = Lutein, 
             fill = Color, 
             color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotLutein

ggsave(boxplotLutein, filename = "boxplotLutein.png", width = 7, height = 5, bg = "white")
```

```{r}
boxplotLutein2 <- dfPig %>%
  ggplot(aes(x = fct_reorder(Site, Latitude, .desc = TRUE), 
               y = Lutein, 
               fill = Color, 
               color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotLutein2

ggsave(boxplotLutein2, filename = "boxplotLutein2.png", width = 7, height = 5, bg = "white")
```

## Pheophorbide 
```{r}
boxplotPheophorbide <- dfPig %>%
  ggplot(aes(x = factor(APLoc, levels = c("North", "South")), 
             y = Pheophorbide, 
             fill = Color, 
             color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotPheophorbide

ggsave(boxplotPheophorbide, filename = "boxplotPheophorbide.png", width = 7, height = 5, bg = "white")
```

```{r}
boxplotPheophorbide2 <- dfPig %>%
  ggplot(aes(x = fct_reorder(Site, Latitude, .desc = TRUE), 
               y = Pheophorbide, 
               fill = Color, 
               color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotPheophorbide2

ggsave(boxplotPheophorbide2, filename = "boxplotPheophorbide2.png", width = 7, height = 5, bg = "white")
```

## Pheophytin
```{r}
boxplotPheophytin <- dfPig %>%
  ggplot(aes(x = factor(APLoc, levels = c("North", "South")), 
             y = Pheophytin, 
             fill = Color, 
             color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotPheophytin

ggsave(boxplotPheophytin, filename = "boxplotPheophytin.png", width = 7, height = 5, bg = "white")
```

```{r}
boxplotPheophytin2 <- dfPig %>%
  ggplot(aes(x = fct_reorder(Site, Latitude, .desc = TRUE), 
               y = Pheophytin, 
               fill = Color, 
               color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotPheophytin2

ggsave(boxplotPheophytin2, filename = "boxplotPheophytin2.png", width = 7, height = 5, bg = "white")
```

## Repeat Visits 

### Petermann Island {.tabset}

#### All 
```{r}
boxplotPI <- dfPig_long %>%
  filter(Site == "Petermann Island") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotPI

ggsave(boxplotPI, filename = "boxplotPI.png", width = 8, height = 5, bg = "white")
```

#### Algae 
```{r}
boxplotPI_algae <- dfPig_long %>%
  filter(Site == "Petermann Island") %>% 
  filter(Algae == "algae") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotPI_algae

ggsave(boxplotPI_algae, filename = "boxplotPI_algae.png", width = 8, height = 5, bg = "white")
```

#### Red
```{r}
boxplotPI_R <- dfPig_long %>%
  filter(Site == "Petermann Island") %>% 
  filter(Color == "red") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotPI_R

ggsave(boxplotPI_R, filename = "boxplotPI_R.png", width = 8, height = 5, bg = "white")
```

#### Green
```{r}
boxplotPI_G <- dfPig_long %>%
  filter(Site == "Petermann Island") %>% 
  filter(Color == "green") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotPI_G

ggsave(boxplotPI_G, filename = "boxplotPI_G.png", width = 8, height = 5, bg = "white")
```

#### ChlA
```{r}
boxplotPIChlA <- dfPig %>%
  filter(Site == "Petermann Island") %>% 
  ggplot(aes(x = factor(Date), 
             y = ChlA, 
             fill = Color, 
             color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotPIChlA

ggsave(boxplotPIChlA, filename = "boxplotPIChlA.png", width = 7, height = 5, bg = "white")
```

#### Astaxanthin
```{r}
boxplotPIAstax <- dfPig %>%
  filter(Site == "Petermann Island") %>% 
  ggplot(aes(x = factor(Date), 
             y = Astaxanthin, 
             fill = Color, 
             color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotPIAstax

ggsave(boxplotPIAstax, filename = "boxplotPIAstax.png", width = 7, height = 5, bg = "white")
```

### Damoy Point {.tabset}

#### All 
```{r}
boxplotDP <- dfPig_long %>%
  filter(Site == "Damoy Point") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotDP

ggsave(boxplotDP, filename = "boxplotDP.png", width = 8, height = 5, bg = "white")
```

#### Algae 
```{r}
boxplotDP_algae <- dfPig_long %>%
  filter(Site == "Damoy Point") %>% 
  filter(Algae == "algae") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotDP_algae

ggsave(boxplotDP_algae, filename = "boxplotDP_algae.png", width = 8, height = 5, bg = "white")
```

#### Red
```{r}
boxplotDP_R <- dfPig_long %>%
  filter(Site == "Damoy Point") %>% 
  filter(Color == "red") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotDP_R

ggsave(boxplotDP_R, filename = "boxplotDP_R.png", width = 8, height = 5, bg = "white")
```

#### Green
```{r}
boxplotDP_G <- dfPig_long %>%
  filter(Site == "Damoy Point") %>% 
  filter(Color == "green") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotDP_G

ggsave(boxplotDP_G, filename = "boxplotDP_G.png", width = 8, height = 5, bg = "white")
```

#### ChlA
```{r}
boxplotDPChlA <- dfPig %>%
  filter(Site == "Damoy Point") %>% 
  ggplot(aes(x = factor(Date), 
             y = ChlA, 
             fill = Color, 
             color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotDPChlA

ggsave(boxplotDPChlA, filename = "boxplotDPChlA.png", width = 7, height = 5, bg = "white")
```

#### Astaxanthin
```{r}
boxplotDPAstax <- dfPig %>%
  filter(Site == "Damoy Point") %>% 
  ggplot(aes(x = factor(Date), 
             y = Astaxanthin, 
             fill = Color, 
             color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotDPAstax

ggsave(boxplotDPAstax, filename = "boxplotDPAstax.png", width = 7, height = 5, bg = "white")
```

### Orne Harbour {.tabset}

#### All 
```{r}
boxplotOH <- dfPig_long %>%
  filter(Site == "Orne Harbour") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotOH

ggsave(boxplotOH, filename = "boxplotOH.png", width = 8, height = 5, bg = "white")
```

#### Algae 
```{r}
boxplotOH_algae <- dfPig_long %>%
  filter(Site == "Orne Harbour") %>% 
  filter(Algae == "algae") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotOH_algae

ggsave(boxplotOH_algae, filename = "boxplotOH_algae.png", width = 8, height = 5, bg = "white")
```

#### Red
```{r}
boxplotOH_R <- dfPig_long %>%
  filter(Site == "Orne Harbour") %>% 
  filter(Color == "red") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotOH_R

ggsave(boxplotOH_R, filename = "boxplotOH_R.png", width = 8, height = 5, bg = "white")
```

#### Green
```{r}
boxplotOH_G <- dfPig_long %>%
  filter(Site == "Orne Harbour") %>% 
  filter(Color == "green") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotOH_G

ggsave(boxplotOH_G, filename = "boxplotOH_G.png", width = 8, height = 5, bg = "white")
```

#### ChlA
```{r}
boxplotOHChlA <- dfPig %>%
  filter(Site == "Orne Harbour") %>% 
  ggplot(aes(x = factor(Date), 
             y = ChlA, 
             fill = Color, 
             color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotOHChlA

ggsave(boxplotOHChlA, filename = "boxplotOHChlA.png", width = 7, height = 5, bg = "white")
```

#### Astaxanthin
```{r}
boxplotOHAstax <- dfPig %>%
  filter(Site == "Orne Harbour") %>% 
  ggplot(aes(x = factor(Date), 
             y = Astaxanthin, 
             fill = Color, 
             color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotOHAstax

ggsave(boxplotOHAstax, filename = "boxplotOHAstax.png", width = 7, height = 5, bg = "white")
```

### Pléneau Island {.tabset}

#### All 
```{r}
boxplotPLE <- dfPig_long %>%
  filter(Site == "Pléneau Island") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotPLE

ggsave(boxplotPLE, filename = "boxplotPLE.png", width = 8, height = 5, bg = "white")
```

#### Algae 
```{r}
boxplotPLE_algae <- dfPig_long %>%
  filter(Site == "Pléneau Island") %>% 
  filter(Algae == "algae") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotPLE_algae

ggsave(boxplotPLE_algae, filename = "boxplotPLE_algae.png", width = 8, height = 5, bg = "white")
```

#### Red
```{r}
boxplotPLE_R <- dfPig_long %>%
  filter(Site == "Pléneau Island") %>% 
  filter(Color == "red") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotPLE_R

ggsave(boxplotPLE_R, filename = "boxplotPLE_R.png", width = 8, height = 5, bg = "white")
```

#### Green
```{r}
boxplotPLE_G <- dfPig_long %>%
  filter(Site == "Pléneau Island") %>% 
  filter(Color == "green") %>% 
  ggplot(aes(x = Pigment, 
             y = Concentration, 
             fill = as.factor(Date),
             color = as.factor(Date))) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotPLE_G

ggsave(boxplotPLE_G, filename = "boxplotPLE_G.png", width = 8, height = 5, bg = "white")
```

#### ChlA
```{r}
boxplotPLEChlA <- dfPig %>%
  filter(Site == "Pléneau Island") %>% 
  ggplot(aes(x = factor(Date), 
             y = ChlA, 
             fill = Color, 
             color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotPLEChlA

ggsave(boxplotPLEChlA, filename = "boxplotPLEChlA.png", width = 7, height = 5, bg = "white")
```

#### Astaxanthin
```{r}
boxplotPLEAstax <- dfPig %>%
  filter(Site == "Pléneau Island") %>% 
  ggplot(aes(x = factor(Date), 
             y = Astaxanthin, 
             fill = Color, 
             color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotPLEAstax

ggsave(boxplotPLEAstax, filename = "boxplotPLEAstax.png", width = 7, height = 5, bg = "white")
```

# Pairs Plots - Pigments {.tabset}

## North
```{r, warning = FALSE, message = FALSE}
corplotNorth <- dfPig %>% 
  filter(APLoc == "North") %>% 
  ggpairs(
  columns = c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotNorth

ggsave(corplotNorth, filename = "corplotNorth.png", width = 7, height = 7, bg = "white")
```

## South
```{r, warning = FALSE, message = FALSE}
corplotSouth <- dfPig %>% 
  filter(APLoc == "South") %>% 
  ggpairs(
  columns = c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotSouth

ggsave(corplotSouth, filename = "corplotSouth.png", width = 7, height = 7, bg = "white")
```

# Heat Maps - Pigments {.tabset}

## All Sites (All Samples)
```{r}
cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
pigment_vars <- c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin")

# List of locations
locations <- unique(dfPig$APLoc)

# Function to calculate correlation matrices for each location
get_corr_data <- function(location) {
  data_sub <- dfPig %>%
    filter(APLoc == location) %>%
    select(all_of(pigment_vars)) %>%
    na.omit()

  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p

  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)

  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Location <- location

  return(merged)
}

# Apply across locations
heatmap_data <- do.call(rbind, lapply(locations, get_corr_data))

# Factor levels for correct ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = pigment_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = pigment_vars)

# Create label
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Remove upper triangle + diagonal if desired (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot
heatmapPigAll <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Location) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank()) 

heatmapPigAll

ggsave(heatmapPigAll, filename = "heatmapPigAll.png", width = 8, height = 5, bg = "white")

```

## North and South Regions (Red Samples)
```{r}
cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
pigment_vars <- c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin")

# List of locations
locations <- unique(dfPig$APLoc)

get_corr_data <- function(location) {
  data_sub <- dfPig %>%
    filter(APLoc %in% c("North", "South")) %>%
    filter(APLoc == location) %>%
    filter(Color == "red") %>%
    select(all_of(pigment_vars)) %>%
    mutate(across(everything(), ~ as.numeric(as.character(.)))) %>%  # ensure numeric
    na.omit()

  # Skip if less than 2 rows after filtering
  if (nrow(data_sub) < 2) return(NULL)

  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p

  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)

  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Location <- location

  return(merged)
}

# Only keep locations with at least 2 "algae" rows
locations_to_use <- dfPig %>%
  filter(APLoc %in% c("North", "South"), Algae == "algae") %>%
  group_by(APLoc) %>%
  filter(n() >= 2) %>%
  pull(APLoc) %>%
  unique()

heatmap_data <- do.call(rbind, lapply(locations_to_use, get_corr_data))

# Factor levels for correct ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = pigment_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = pigment_vars)

# Create label
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Remove upper triangle + diagonal if desired (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot
heatmapPigRed <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Location) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank()) 

heatmapPigRed

ggsave(heatmapPigRed, filename = "heatmapPigRed.png", width = 8, height = 5, bg = "white")

```

## North and South Regions (Green Samples)
```{r}
cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
pigment_vars <- c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin")

# List of locations
locations <- unique(dfPig$APLoc)

get_corr_data <- function(location) {
  data_sub <- dfPig %>%
    filter(APLoc %in% c("North", "South")) %>%
    filter(APLoc == location) %>%
    filter(Color == "green") %>%
    select(all_of(pigment_vars)) %>%
    mutate(across(everything(), ~ as.numeric(as.character(.)))) %>%  # ensure numeric
    na.omit()

  # Skip if less than 2 rows after filtering
  if (nrow(data_sub) < 2) return(NULL)

  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p

  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)

  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Location <- location

  return(merged)
}

# Only keep locations with at least 2 "algae" rows
locations_to_use <- dfPig %>%
  filter(APLoc %in% c("North", "South"), Algae == "algae") %>%
  group_by(APLoc) %>%
  filter(n() >= 2) %>%
  pull(APLoc) %>%
  unique()

heatmap_data <- do.call(rbind, lapply(locations_to_use, get_corr_data))

# Factor levels for correct ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = pigment_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = pigment_vars)

# Create label
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Remove upper triangle + diagonal if desired (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot
heatmapPigGreen <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Location) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank()) 

heatmapPigGreen

ggsave(heatmapPigGreen, filename = "heatmapPigGreen.png", width = 8, height = 5, bg = "white")

```

# Scatter Plots - Pigments & Anions

## North Region {.tabset}

### ChlA
```{r}
scatterMatrixNorthChlA <- dfAnion_long %>% 
  filter(APLoc == "North") %>% 
  ggplot(aes(x = Concentration, y = ChlA, color = Color)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", linewidth = 0.8) +
  scale_color_manual(values = colorPal) +
  scale_x_sqrt() +
  facet_grid(. ~ Anion, scales = "free_x") +
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text.x = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "ChlA ",
       color = NULL)

scatterMatrixNorthChlA
```

### ChlB
```{r}
scatterMatrixNorthChlB <- dfAnion_long %>% 
  filter(APLoc == "North") %>%
  ggplot(aes(x = Concentration, y = ChlB, color = Color)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", linewidth = 0.8) +
  scale_color_manual(values = colorPal) +
  scale_x_sqrt() +
  facet_grid(. ~ Anion, scales = "free_x") +
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text.x = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "ChlB ",
       color = NULL)

scatterMatrixNorthChlB
```

### Astaxanthin
```{r}
scatterMatrixNorthAstaxanthin <- dfAnion_long %>% 
  filter(APLoc == "North") %>%
  ggplot(aes(x = Concentration, y = Astaxanthin, color = Color)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", linewidth = 0.8) +
  scale_color_manual(values = colorPal) +
  scale_x_sqrt() +
  facet_grid(. ~ Anion, scales = "free_x") +
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text.x = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "Astaxanthin ",
       color = NULL)

scatterMatrixNorthAstaxanthin
```

### Lutein
```{r}
scatterMatrixNorthLutein <- dfAnion_long %>% 
  filter(APLoc == "North") %>% 
  ggplot(aes(x = Concentration, y = Lutein, color = Color)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", linewidth = 0.8) +
  scale_color_manual(values = colorPal) +
  scale_x_sqrt() +
  facet_grid(. ~ Anion, scales = "free_x") +
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text.x = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "Lutein ",
       color = NULL)

scatterMatrixNorthLutein
```

### BetaC
```{r}
scatterMatrixNorthBetaC <- dfAnion_long %>% 
  filter(APLoc == "North") %>%
  ggplot(aes(x = Concentration, y = BetaC, color = Color)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", linewidth = 0.8) +
  scale_color_manual(values = colorPal) +
  scale_x_sqrt() +
  facet_grid(. ~ Anion, scales = "free_x") +
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text.x = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "BetaC ",
       color = NULL)

scatterMatrixNorthBetaC
```

### Pheophytin
```{r}
scatterMatrixNorthPheophytin <- dfAnion_long %>% 
  filter(APLoc == "North") %>%
  ggplot(aes(x = Concentration, y = Pheophytin, color = Color)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", linewidth = 0.8) +
  scale_color_manual(values = colorPal) +
  scale_x_sqrt() +
  facet_grid(. ~ Anion, scales = "free_x") +
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text.x = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "Pheophytin ",
       color = NULL)

scatterMatrixNorthPheophytin
```

### Pheophorbide
```{r}
scatterMatrixNorthPheophorbide <- dfAnion_long %>% 
  filter(APLoc == "North") %>% 
  ggplot(aes(x = Concentration, y = Pheophorbide, color = Color)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", linewidth = 0.8) +
  scale_color_manual(values = colorPal) +
  scale_x_sqrt() +
  facet_grid(. ~ Anion, scales = "free_x") +
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text.x = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "Pheophorbide ",
       color = NULL)

scatterMatrixNorthPheophorbide
```

### Combined
```{r}
scatterMatrixNorthChlA <- scatterMatrixNorthChlA + theme(axis.title.x = element_blank(), axis.text.x = element_blank())
scatterMatrixNorthChlB <- scatterMatrixNorthChlB + theme(axis.title.x = element_blank(), axis.text.x = element_blank())
scatterMatrixNorthAstaxanthin <- scatterMatrixNorthAstaxanthin + theme(axis.title.x = element_blank(), axis.text.x = element_blank())
scatterMatrixNorthLutein <- scatterMatrixNorthLutein + theme(axis.title.x = element_blank(), axis.text.x = element_blank())
scatterMatrixNorthPheophorbide <- scatterMatrixNorthPheophorbide + theme(axis.title.x = element_blank(), axis.text.x = element_blank())

scatterMatrixNorth <- (
  scatterMatrixNorthChlA /
  scatterMatrixNorthChlB /
  scatterMatrixNorthAstaxanthin /
  scatterMatrixNorthLutein /
  scatterMatrixNorthBetaC 
) +
  plot_layout(guides = "collect", heights = c(1,1,1,1,1)) &
  theme(legend.position = "bottom",
        plot.margin = margin(2, 2, 2, 2)) 

scatterMatrixNorth

ggsave(scatterMatrixNorth, filename = "scatterMatrixNorth.png", width = 8, height = 9.5, bg = "white")
```

```{r}
scatterMatrixNorthDeg <- (
  scatterMatrixNorthPheophorbide /
  scatterMatrixNorthPheophytin 
) +
  plot_layout(guides = "collect", heights = c(1,1)) &
  theme(legend.position = "bottom",
        plot.margin = margin(2,2,2,2)) 

scatterMatrixNorthDeg

ggsave(scatterMatrixNorthDeg, filename = "scatterMatrixNorthDeg.png", width = 8, height = 5, bg = "white")
```

## South Region {.tabset}

### ChlA
```{r}
scatterMatrixSouthChlA <- dfAnion_long %>% 
  filter(APLoc == "South") %>% 
  ggplot(aes(x = Concentration, y = ChlA, color = Color)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", linewidth = 0.8) +
  scale_color_manual(values = colorPal) +
  scale_x_sqrt() +
  facet_grid(. ~ Anion, scales = "free_x") +
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text.x = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "ChlA ",
       color = NULL)

scatterMatrixSouthChlA
```

### ChlB
```{r}
scatterMatrixSouthChlB <- dfAnion_long %>% 
  filter(APLoc == "South") %>%
  ggplot(aes(x = Concentration, y = ChlB, color = Color)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", linewidth = 0.8) +
  scale_color_manual(values = colorPal) +
  scale_x_sqrt() +
  facet_grid(. ~ Anion, scales = "free_x") +
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text.x = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "ChlB ",
       color = NULL)

scatterMatrixSouthChlB
```

### Astaxanthin
```{r}
scatterMatrixSouthAstaxanthin <- dfAnion_long %>% 
  filter(APLoc == "South") %>%
  ggplot(aes(x = Concentration, y = Astaxanthin, color = Color)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", linewidth = 0.8) +
  scale_color_manual(values = colorPal) +
  scale_x_sqrt() +
  facet_grid(. ~ Anion, scales = "free_x") +
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text.x = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "Astaxanthin ",
       color = NULL)

scatterMatrixSouthAstaxanthin
```

### Lutein
```{r}
scatterMatrixSouthLutein <- dfAnion_long %>% 
  filter(APLoc == "South") %>% 
  ggplot(aes(x = Concentration, y = Lutein, color = Color)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", linewidth = 0.8) +
  scale_color_manual(values = colorPal) +
  scale_x_sqrt() +
  facet_grid(. ~ Anion, scales = "free_x") +
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text.x = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "Lutein ",
       color = NULL)

scatterMatrixSouthLutein
```

### BetaC
```{r}
scatterMatrixSouthBetaC <- dfAnion_long %>% 
  filter(APLoc == "South") %>%
  ggplot(aes(x = Concentration, y = BetaC, color = Color)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", linewidth = 0.8) +
  scale_color_manual(values = colorPal) +
  scale_x_sqrt() +
  facet_grid(. ~ Anion, scales = "free_x") +
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text.x = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "BetaC ",
       color = NULL)

scatterMatrixSouthBetaC
```

### Pheophytin
```{r}
scatterMatrixSouthPheophytin <- dfAnion_long %>% 
  filter(APLoc == "South") %>%
  ggplot(aes(x = Concentration, y = Pheophytin, color = Color)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", linewidth = 0.8) +
  scale_color_manual(values = colorPal) +
  scale_x_sqrt() +
  facet_grid(. ~ Anion, scales = "free_x") +
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text.x = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "Pheophytin ",
       color = NULL)

scatterMatrixSouthPheophytin
```

### Pheophorbide
```{r}
scatterMatrixSouthPheophorbide <- dfAnion_long %>% 
  filter(APLoc == "South") %>% 
  ggplot(aes(x = Concentration, y = Pheophorbide, color = Color)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", linewidth = 0.8) +
  scale_color_manual(values = colorPal) +
  scale_x_sqrt() +
  facet_grid(. ~ Anion, scales = "free_x") +
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text.x = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "Pheophorbide ",
       color = NULL)

scatterMatrixSouthPheophorbide
```

### Combined
```{r}
scatterMatrixSouthChlA <- scatterMatrixSouthChlA + theme(axis.title.x = element_blank(), axis.text.x = element_blank())
scatterMatrixSouthChlB <- scatterMatrixSouthChlB + theme(axis.title.x = element_blank(), axis.text.x = element_blank())
scatterMatrixSouthAstaxanthin <- scatterMatrixSouthAstaxanthin + theme(axis.title.x = element_blank(), axis.text.x = element_blank())
scatterMatrixSouthLutein <- scatterMatrixSouthLutein + theme(axis.title.x = element_blank(), axis.text.x = element_blank())
scatterMatrixSouthPheophorbide <- scatterMatrixSouthPheophorbide + theme(axis.title.x = element_blank(), axis.text.x = element_blank())

scatterMatrixSouth <- (
  scatterMatrixSouthChlA /
  scatterMatrixSouthChlB /
  scatterMatrixSouthAstaxanthin /
  scatterMatrixSouthLutein /
  scatterMatrixSouthBetaC 
) +
  plot_layout(guides = "collect", heights = c(1,1,1,1,1)) &
  theme(legend.position = "bottom",
        plot.margin = margin(2, 2, 2, 2)) 

scatterMatrixSouth

ggsave(scatterMatrixSouth, filename = "scatterMatrixSouth.png", width = 8, height = 9.5, bg = "white")
```

```{r}
scatterMatrixSouthDeg <- (
  scatterMatrixSouthPheophorbide /
  scatterMatrixSouthPheophytin 
) +
  plot_layout(guides = "collect", heights = c(1,1)) &
  theme(legend.position = "bottom",
        plot.margin = margin(2,2,2,2)) 

scatterMatrixSouthDeg

ggsave(scatterMatrixSouthDeg, filename = "scatterMatrixSouthDeg.png", width = 8, height = 5, bg = "white")
```

# Heat Maps - Pigments & Anions {.tabset}

## North and South Regions (All Samples)
```{r}
# Define pigment and anion variables
pigment_vars <- c("ChlA", "ChlB", "Astaxanthin", "Lutein", "BetaC", "Pheophorbide", "Pheophytin")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filter to just the necessary columns & regions
df_filtered <- dfAnion %>%
  select(APLoc, all_of(pigment_vars), all_of(anion_vars)) %>%
  filter(APLoc %in% c("North", "South")) %>%
  na.omit()

# Function to compute pigment–anion Spearman correlations per region
get_pigment_anion_corr <- function(data, aploc_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")

  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Create all pigment–anion pair combinations
  pairs <- expand.grid(Pigment = pigment_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Anion, Pigment],
      p_value = p_vals[Anion, Pigment],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      APLoc = aploc_name
    ) %>%
    ungroup()

  return(results)
}

# Compute correlations for each APLoc region
cor_results <- df_filtered %>%
  group_split(APLoc) %>%
  map_dfr(~get_pigment_anion_corr(.x, unique(.x$APLoc)))

# Convert to factors to preserve axis order
cor_results$Pigment <- factor(cor_results$Pigment, levels = pigment_vars)
cor_results$Anion <- factor(cor_results$Anion, levels = anion_vars)

# Plot
heatmapPigAnion <- ggplot(cor_results, aes(x = Pigment, y = Anion, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  facet_wrap(~APLoc, ncol = 1) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

heatmapPigAnion

ggsave(heatmapPigAnion, filename = "heatmapPigAnion.png", width = 7, height = 9, bg = "white")
```

## North and South Regions (Red Samples)
```{r}
# Define pigment and anion variables
pigment_vars <- c("ChlA", "ChlB", "Astaxanthin", "Lutein", "BetaC", "Pheophorbide", "Pheophytin")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filter to just the necessary columns & regions
df_filtered <- dfAnion %>%
  filter(Color == "red") %>% 
  select(APLoc, all_of(pigment_vars), all_of(anion_vars)) %>%
  filter(APLoc %in% c("North", "South")) %>%
  na.omit()

# Function to compute pigment–anion Spearman correlations per region
get_pigment_anion_corr <- function(data, aploc_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")

  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Create all pigment–anion pair combinations
  pairs <- expand.grid(Pigment = pigment_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Anion, Pigment],
      p_value = p_vals[Anion, Pigment],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      APLoc = aploc_name
    ) %>%
    ungroup()

  return(results)
}

# Compute correlations for each APLoc region
cor_results <- df_filtered %>%
  group_split(APLoc) %>%
  map_dfr(~get_pigment_anion_corr(.x, unique(.x$APLoc)))

# Convert to factors to preserve axis order
cor_results$Pigment <- factor(cor_results$Pigment, levels = pigment_vars)
cor_results$Anion <- factor(cor_results$Anion, levels = anion_vars)

# Plot
heatmapPigAnion_Red <- ggplot(cor_results, aes(x = Pigment, y = Anion, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  facet_wrap(~APLoc, ncol = 1) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

heatmapPigAnion_Red

ggsave(heatmapPigAnion_Red, filename = "heatmapPigAnion_Red.png", width = 7, height = 9, bg = "white")
```

## North and South Regions (Green Samples)
```{r}
# Define pigment and anion variables
pigment_vars <- c("ChlA", "ChlB", "Astaxanthin", "Lutein", "BetaC", "Pheophorbide", "Pheophytin")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filter to just the necessary columns & regions
df_filtered <- dfAnion %>%
  filter(Color == "green") %>% 
  select(APLoc, all_of(pigment_vars), all_of(anion_vars)) %>%
  filter(APLoc %in% c("North", "South")) %>%
  na.omit()

# Function to compute pigment–anion Spearman correlations per region
get_pigment_anion_corr <- function(data, aploc_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")

  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Create all pigment–anion pair combinations
  pairs <- expand.grid(Pigment = pigment_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Anion, Pigment],
      p_value = p_vals[Anion, Pigment],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      APLoc = aploc_name
    ) %>%
    ungroup()

  return(results)
}

# Compute correlations for each APLoc region
cor_results <- df_filtered %>%
  group_split(APLoc) %>%
  map_dfr(~get_pigment_anion_corr(.x, unique(.x$APLoc)))

# Convert to factors to preserve axis order
cor_results$Pigment <- factor(cor_results$Pigment, levels = pigment_vars)
cor_results$Anion <- factor(cor_results$Anion, levels = anion_vars)

# Plot
heatmapPigAnion_Green <- ggplot(cor_results, aes(x = Pigment, y = Anion, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  facet_wrap(~APLoc, ncol = 1) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

heatmapPigAnion_Green

ggsave(heatmapPigAnion_Green, filename = "heatmapPigAnion_Green.png", width = 7, height = 9, bg = "white")
```