---
title: "Data Vizualization - Pigment Concentration Normalized by Cell Density"
author: "Ellie Ryan"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: flatly
    highlight: zenburn
    toc: true
    toc_float: true
    toc_collapsed: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Created:** 7/16/2025

**Last Modified:** 1/23/2025

```{r, message=FALSE, warning=FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(stringr)
library(kableExtra)
library(scales)
library(GGally)
library(patchwork)
library(Hmisc)
library(reshape2)
library(car)
library(FSA)
library(vegan)
```

```{r}
regionPal <- c(
  "North" = "#284C46FF",
  "South" = "#b04a25"
) 

colorPal <- c(
  "clean" = "#95879f",
  "red" = "#b04a25",
  "green" = "#647d6a"
)

algaePal <- c(
  "clean" = "#a5bdc0",
  "algae" = "#735a69"
)

pigPal <- c(
  "Astaxanthin" = "#b04a25",
  "Lutein" = "#d9983b",
  "ChlA" = "#284C46FF",
  "ChlB" = "#647d6a",
  "BetaC" = "#4d1d05",
  "Pheophorbide" = "#414c66", 
  "Pheophytin" = "#95879f"
)

plot_theme <- theme(
  axis.title = element_text(size = 12, face = "bold"), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  plot.margin = margin(20, 20, 20, 20)
)

plot_themePub <- theme(
  text = element_text(family = "Helvetica"),
  axis.title.x = element_text(size = 10, margin = margin(t = 2)),  # t = top margin
  axis.title.y = element_text(size = 10, margin = margin(r = 4)),  # r = right margin
  axis.text = element_text(size = 10),
  axis.line = element_blank(),
  axis.ticks.x = element_blank(),
  axis.ticks.y = element_line(color = "black", linewidth = 0.25),  # Add tick marks
  axis.ticks.length = unit(-0.15, "cm"),  # Length of tick marks
  panel.grid = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
  legend.position = "top",
  legend.justification = c(0, 1),
  legend.text = element_text(size = 10),
  legend.title = element_text(size = 10),
  plot.title = element_text(size = 14, face = "bold"),
  plot.subtitle = element_text(size = 13, face = "italic"),
  plot.margin = margin(3, 5, 3, 3)
)

plot_theme2 <- theme(
  axis.title = element_text(size = 12, face = "bold"), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  axis.text.x = element_text(angle = 45, hjust = 1), 
  plot.margin = margin(20, 20, 20, 20)
)

pairsplot_theme <- theme(
  text = element_text(size = 10),  # Adjust variable name text size
  axis.text = element_text(size = 8),  # Adjust axis text size
  plot.title = element_text(size = 12, face = "bold"),  # Adjust plot title size
  legend.text = element_text(size = 12))  # Adjust legend text size
```


# Overview

The purpose of this `.Rmd` is to visualize normalized pigment concentration (ng/cell) by site and color, and obtain basic summary statistics associated with each pigment in order to inform my next steps in analysis. I will additionally perform an ordination to reduce the dimensions of my pigment data for simpler visualizations. Before proceeding, I will clean the dataset so that it features only the variables I am interested in for this analysis, and normalize pigment concentration by cell density. In this dataset, 106 samples are used from the Antarctic Peninsula, following the removal of 5 outliers. These observations are broken down by `Color`, or the apparent snow color at time of sample collection. To keep things simple, there are only three levels to the `Color` category: red, green and clean. The breakdown of samples collected in the WAP is as follows:

-   Antarctic Peninsula:
    -   107 samples
        -   43 red
        -   44 green
        -   18 clean
        
*NOTE: Fewer samples were used for analysis, since there are NA values for pigment concentrations of some samples.*
        
From anion outlier analysis, the following outliers were identified for removal:

-   2025-02-19-SI-algae3
-   2025-02-24-DA-algae2
-   2025-02-24-DA-algae6
-   2025-03-07-NH-algae3
-   2025-01-12-PI-algae10
-   2025-01-17-PI-algae12 
-   2025-01-17-PI-algae22

See the Preliminary Results and Figures doc for justification for removal of each. 

## Data Wrangling

```{r}
df <- read.csv("CompiledDataPI.csv")
str(df)
```

Adding an `Algae` column to classify samples as algae or clean snow samples and a `Region` column to separate North and South samples.

```{r}
# Format date column 
df <- df %>%
  mutate(
    Date = as.Date(Date, format = "%m/%d/%y"))

# Add Algae
df <- df %>%  
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  relocate(Algae, .after = Color)

# Add Region
df <- df %>% 
  mutate(
  Region = if_else(
    Site %in% c("Horseshoe Island", "Stonington Island"),
    "South",
    "North")) %>% 
  relocate(Region, .after = Site)

# Add LocAbbrev
df <- df %>% mutate(LocAbbrev = case_when(
  Location == "Antarctic Peninsula" ~ paste("WAP")
)) %>%  
  relocate(LocAbbrev, .after = Site)
```

And normalize pigment concentration by dividing by cell density and multiplying by 1000 to get units of ng/cell. This results in NaN and Inf values for samples with a cell density of 0 or NA. I'll remove these before proceeding. 
```{r}
dfNorm <- df %>% 
  mutate(ChlA = (ChlA/CellDensity)*1000,
         ChlB = (ChlB/CellDensity)*1000,
         Astaxanthin = (Astaxanthin/CellDensity)*1000,
         Lutein = (Lutein/CellDensity)*1000,
         BetaC = (BetaC/CellDensity)*1000,
         Pheophytin = (Pheophytin/CellDensity)*1000,
         Pheophorbide = (Pheophorbide/CellDensity)*1000) 
```

```{r}
str(dfNorm)
```

```{r}
outliers <- c(
  "2025-02-19-SI-algae3",
  "2025-02-24-DA-algae2",
  "2025-02-24-DA-algae6",
  "2025-03-07-NH-algae3",
  "2025-01-12-PI-algae10",
  "2025-01-17-PI-algae12",
  "2025-01-17-PI-algae22"
  )

dfNorm <- dfNorm %>% 
  filter(!SampleID %in% outliers)
```

### Pigments
```{r}
dfPig <- dfNorm %>% 
  select(c("SampleID", "Date", "Location", "Site", "Region", "LocAbbrev", "Color", "Algae", Astaxanthin:Pheophytin)) %>% 
  na.omit(dfPig)
```

### Pigments & Cells
```{r}
dfPigCell <- dfNorm %>% 
  select(c("SampleID", "Date", "Location", "Site", "Region", "LocAbbrev", "Color", "Algae", "CellDensity", "CellLength", Astaxanthin:Pheophytin)) %>% 
  na.omit(dfPig)

dfPigCell <- dfPigCell %>% 
  mutate(
    # Calulate volume (µm^3)
    CellVolume = (pi/6)*(CellLength^3)) %>% 
  relocate(CellVolume, .after = CellLength)
```

### Pigments & Anions 
```{r}
dfAnion <- dfNorm %>% 
  select(c("SampleID", "Date", "Location", "Site", "Region", "LocAbbrev", "Color", "Algae", Astaxanthin:Pheophytin, Fo:SO42)) %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  mutate(NO2NO3 = NO2 + NO3) %>% 
  relocate(Algae, .after = Color) %>% 
  relocate(NO2NO3, .after = Cl) %>% 
  select(!NO2) %>% 
  select(!NO3) %>% 
  na.omit(dfAnion)
```

### Pigments (Pivoted)
```{r}
dfPig_long <- dfPig %>% 
  pivot_longer(cols = Astaxanthin:Pheophytin,
               names_to = "Pigment",
               values_to = "Concentration")
```

### Anions (Pivoted)
```{r}
dfAnion_long <- dfAnion %>% 
  pivot_longer(cols = c(Fo, Cl, SO42, PO43, NO2NO3),
               names_to = "Anion",
               values_to = "Concentration")
```

# ANOVA/Kruskal Wallis Test
Before proceeding with an ANOVA test to assess differences in pigment concentration between snow colors, I am going to check to see if these variables follow the assumptions of normality and equal variance required to run an ANOVA: 
```{r}
# Step 1: Group by pigment and nest data
nested <- dfPig_long %>%
  group_by(Pigment) %>%
  nest()

# Step 2: Fit ANOVA model, extract residuals, and run tests
anovaAssumptionsResults <- nested %>%
  mutate(
    # Fit ANOVA model
    aov_model = map(data, ~ aov(Concentration ~ Color, data = .x)),
    
    # Shapiro-Wilk test for normality of residuals
    shapiro_p = map_dbl(aov_model, ~ shapiro.test(residuals(.x))$p.value),
    
    # Levene's Test for homogeneity of variance
    levene_p = map_dbl(data, ~leveneTest(Concentration ~ factor(Color), data = .x)$`Pr(>F)`[1])
  )
```

From the below results, it appears that the majority of the pigments are neither normally distributed or have equal variances. Because of this, I am going to proceed with a Kruskal-Wallis test in lieu of an ANOVA. 

```{r}
anovaAssumptionsResults %>%
  select(Pigment, shapiro_p, levene_p) %>%
  mutate(
    normality_ok = shapiro_p > 0.05,
    homogeneity_ok = levene_p > 0.05
  )
```

## Kruskal-Wallis Test
Below, I use the Kruskal-Wallis test to run a non-parametric comparison of each pigment across each snow color. Then, I follow up with a Dunn's post hoc test for the pigments that show significant differences between snow colors. 

```{r}
# Run kruskal wallis test 
kruskalResultsAP <- dfPig_long %>%
  filter(Location == "Antarctic Peninsula") %>% 
  group_by(Pigment) %>%
  nest() %>%
  mutate(
    kruskal = map(data, ~ kruskal.test(Concentration ~ Color, data = .x)),
    kw_p = map_dbl(kruskal, ~ .x$p.value)
  )

# Run Dunn's test 
kruskalResultsAP <- kruskalResultsAP %>%
  mutate(
    dunn = map(data, ~ dunnTest(Concentration ~ factor(Color), data = .x, method = "bonferroni")),
    dunn_table = map(dunn, ~ .x$res)
  )
```

#### Visualize Results
```{r}
dunnPairsAP <- kruskalResultsAP %>%
  select(Pigment, dunn_table) %>%
  unnest(dunn_table) %>%
  mutate(
    significant = P.adj < 0.05,
    label = paste0("p=", round(P.adj, 3))
  )
```

```{r}
dunnPlot_AP <- ggplot(dunnPairsAP, aes(x = Comparison, y = Pigment, fill = P.adj)) +
  geom_tile() +
  geom_tile(data = filter(dunnPairsAP, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

dunnPlot_AP

ggsave(dunnPlot_AP, filename = "dunnPlot_AP.png", width = 7, height = 6, bg = "white")
```

# Summary Tables {.tabset}

## Primary Pigments
```{r}
# Helper function: significant figures + commas + preserve .0 for integers
fmt_sigfig <- function(x, sig = 4) {
  # Round to significant figures
  y <- signif(x, sig)
  
  # Convert to string with commas
  out <- prettyNum(y, big.mark = ",", scientific = FALSE)
  
  # Add ".0" if number is an integer after rounding
  out[grepl("^\\d+$", out)] <- paste0(out[grepl("^\\d+$", out)], ".0")
  
  return(out)
}

# Pigment Summary (ChlA and ChlB)
ChlSummary <- dfPig %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    ChlA_mean   = fmt_sigfig(mean(ChlA, na.rm = TRUE)),
    ChlA_median = fmt_sigfig(median(ChlA, na.rm = TRUE)),
    ChlA_max    = fmt_sigfig(max(ChlA, na.rm = TRUE)),
    ChlA_min    = fmt_sigfig(min(ChlA, na.rm = TRUE)),
    ChlA_sd     = fmt_sigfig(sd(ChlA, na.rm = TRUE)),
    ChlB_mean   = fmt_sigfig(mean(ChlB, na.rm = TRUE)),
    ChlB_median = fmt_sigfig(median(ChlB, na.rm = TRUE)),
    ChlB_max    = fmt_sigfig(max(ChlB, na.rm = TRUE)),
    ChlB_min    = fmt_sigfig(min(ChlB, na.rm = TRUE)),
    ChlB_sd     = fmt_sigfig(sd(ChlB, na.rm = TRUE)),
    .groups = "drop"
  )

# Add "all" for each site
ChlSummaryAll <- dfPig %>%
  summarise(
    Color = "all",
    n = n(),
    ChlA_mean   = fmt_sigfig(mean(ChlA, na.rm = TRUE)),
    ChlA_median = fmt_sigfig(median(ChlA, na.rm = TRUE)),
    ChlA_max    = fmt_sigfig(max(ChlA, na.rm = TRUE)),
    ChlA_min    = fmt_sigfig(min(ChlA, na.rm = TRUE)),
    ChlA_sd     = fmt_sigfig(sd(ChlA, na.rm = TRUE)),
    ChlB_mean   = fmt_sigfig(mean(ChlB, na.rm = TRUE)),
    ChlB_median = fmt_sigfig(median(ChlB, na.rm = TRUE)),
    ChlB_max    = fmt_sigfig(max(ChlB, na.rm = TRUE)),
    ChlB_min    = fmt_sigfig(min(ChlB, na.rm = TRUE)),
    ChlB_sd     = fmt_sigfig(sd(ChlB, na.rm = TRUE)),
    .groups = "drop"
  )

# Filter Antarctic red + green
ChlSummaryAPRG <- dfPig %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Color = "red+green",
    n = n(),
    ChlA_mean   = fmt_sigfig(mean(ChlA, na.rm = TRUE)),
    ChlA_median = fmt_sigfig(median(ChlA, na.rm = TRUE)),
    ChlA_max    = fmt_sigfig(max(ChlA, na.rm = TRUE)),
    ChlA_min    = fmt_sigfig(min(ChlA, na.rm = TRUE)),
    ChlA_sd     = fmt_sigfig(sd(ChlA, na.rm = TRUE)),
    ChlB_mean   = fmt_sigfig(mean(ChlB, na.rm = TRUE)),
    ChlB_median = fmt_sigfig(median(ChlB, na.rm = TRUE)),
    ChlB_max    = fmt_sigfig(max(ChlB, na.rm = TRUE)),
    ChlB_min    = fmt_sigfig(min(ChlB, na.rm = TRUE)),
    ChlB_sd     = fmt_sigfig(sd(ChlB, na.rm = TRUE))
  )

# Combine and order rows
ChlSummary <- bind_rows(ChlSummaryAll, ChlSummary, ChlSummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
ChlSummary %>% 
  kable("html", caption = "Summary Statistics for ChlA and ChlB by Snow Color (ng/cell)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Carotenoids
```{r}
# Pigment Summary (Astaxanthin, Beta Carotene and Lutein)
CarotenoidSummary <- dfPig %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    Astaxanthin_mean = fmt_sigfig(mean(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_median = fmt_sigfig(median(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_max = fmt_sigfig(max(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_min = fmt_sigfig(min(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_sd = fmt_sigfig(sd(Astaxanthin, na.rm = TRUE)),
    Lutein_mean = fmt_sigfig(mean(Lutein, na.rm = TRUE)),
    Lutein_median = fmt_sigfig(median(Lutein, na.rm = TRUE)),
    Lutein_max = fmt_sigfig(max(Lutein, na.rm = TRUE)),
    Lutein_min = fmt_sigfig(min(Lutein, na.rm = TRUE)),
    Lutein_sd = fmt_sigfig(sd(Lutein, na.rm = TRUE)),
    BetaC_mean = fmt_sigfig(mean(BetaC, na.rm = TRUE)),
    BetaC_median = fmt_sigfig(median(BetaC, na.rm = TRUE)),
    BetaC_max = fmt_sigfig(max(BetaC, na.rm = TRUE)),
    BetaC_min = fmt_sigfig(min(BetaC, na.rm = TRUE)),
    BetaC_sd = fmt_sigfig(sd(BetaC, na.rm = TRUE)),
    .groups = "drop"
  )

# Add "all" for each site
CarotenoidSummaryAll <- dfPig %>% 
  summarise(
    Color = "all",
    n = n(),
    Astaxanthin_mean = fmt_sigfig(mean(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_median = fmt_sigfig(median(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_max = fmt_sigfig(max(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_min = fmt_sigfig(min(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_sd = fmt_sigfig(sd(Astaxanthin, na.rm = TRUE)),
    Lutein_mean = fmt_sigfig(mean(Lutein, na.rm = TRUE)),
    Lutein_median = fmt_sigfig(median(Lutein, na.rm = TRUE)),
    Lutein_max = fmt_sigfig(max(Lutein, na.rm = TRUE)),
    Lutein_min = fmt_sigfig(min(Lutein, na.rm = TRUE)),
    Lutein_sd = fmt_sigfig(sd(Lutein, na.rm = TRUE)),
    BetaC_mean = fmt_sigfig(mean(BetaC, na.rm = TRUE)),
    BetaC_median = fmt_sigfig(median(BetaC, na.rm = TRUE)),
    BetaC_max = fmt_sigfig(max(BetaC, na.rm = TRUE)),
    BetaC_min = fmt_sigfig(min(BetaC, na.rm = TRUE)),
    BetaC_sd = fmt_sigfig(sd(BetaC, na.rm = TRUE)),
    .groups = "drop"
  )

# Filter Antarctic red + green
CarotenoidSummaryAPRG <- dfPig %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Color = "red+green",
    n = n(),
    Astaxanthin_mean = fmt_sigfig(mean(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_median = fmt_sigfig(median(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_max = fmt_sigfig(max(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_min = fmt_sigfig(min(Astaxanthin, na.rm = TRUE)),
    Astaxanthin_sd = fmt_sigfig(sd(Astaxanthin, na.rm = TRUE)),
    Lutein_mean = fmt_sigfig(mean(Lutein, na.rm = TRUE)),
    Lutein_median = fmt_sigfig(median(Lutein, na.rm = TRUE)),
    Lutein_max = fmt_sigfig(max(Lutein, na.rm = TRUE)),
    Lutein_min = fmt_sigfig(min(Lutein, na.rm = TRUE)),
    Lutein_sd = fmt_sigfig(sd(Lutein, na.rm = TRUE)),
    BetaC_mean = fmt_sigfig(mean(BetaC, na.rm = TRUE)),
    BetaC_median = fmt_sigfig(median(BetaC, na.rm = TRUE)),
    BetaC_max = fmt_sigfig(max(BetaC, na.rm = TRUE)),
    BetaC_min = fmt_sigfig(min(BetaC, na.rm = TRUE)),
    BetaC_sd = fmt_sigfig(sd(BetaC, na.rm = TRUE))
  )

# Combine and order rows
CarotenoidSummary <- bind_rows(CarotenoidSummaryAll, CarotenoidSummary, CarotenoidSummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
CarotenoidSummary %>% 
  kable("html", caption = "Summary Statistics for Astaxanthin and Lutein by Snow Color (ng/cell)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```


## Degradation Products 
```{r}
# Pigment Summary (Pheophytin and Pheophorbide)
DegSummary <- dfPig %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    Pheophytin_mean = fmt_sigfig(mean(Pheophytin, na.rm = TRUE)),
    Pheophytin_median = fmt_sigfig(median(Pheophytin, na.rm = TRUE)),
    Pheophytin_max = fmt_sigfig(max(Pheophytin, na.rm = TRUE)),
    Pheophytin_min = fmt_sigfig(min(Pheophytin, na.rm = TRUE)),
    Pheophytin_sd = fmt_sigfig(sd(Pheophytin, na.rm = TRUE)),
    Pheophorbide_mean = fmt_sigfig(mean(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_median = fmt_sigfig(median(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_max = fmt_sigfig(max(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_min = fmt_sigfig(min(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_sd = fmt_sigfig(sd(Pheophorbide, na.rm = TRUE)),
    .groups = "drop"
  )

# Add "all" for each site
DegSummaryAll <- dfPig %>%
  summarise(
    Color = "all",
    n = n(),
    Pheophytin_mean = fmt_sigfig(mean(Pheophytin, na.rm = TRUE)),
    Pheophytin_median = fmt_sigfig(median(Pheophytin, na.rm = TRUE)),
    Pheophytin_max = fmt_sigfig(max(Pheophytin, na.rm = TRUE)),
    Pheophytin_min = fmt_sigfig(min(Pheophytin, na.rm = TRUE)),
    Pheophytin_sd = fmt_sigfig(sd(Pheophytin, na.rm = TRUE)),
    Pheophorbide_mean = fmt_sigfig(mean(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_median = fmt_sigfig(median(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_max = fmt_sigfig(max(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_min = fmt_sigfig(min(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_sd = fmt_sigfig(sd(Pheophorbide, na.rm = TRUE)),
    .groups = "drop"
  )

# Filter Antarctic red + green
DegSummaryAPRG <- dfPig %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Location = "Antarctic Peninsula",
    Color = "red+green",
    n = n(),
    Pheophytin_mean = fmt_sigfig(mean(Pheophytin, na.rm = TRUE)),
    Pheophytin_median = fmt_sigfig(median(Pheophytin, na.rm = TRUE)),
    Pheophytin_max = fmt_sigfig(max(Pheophytin, na.rm = TRUE)),
    Pheophytin_min = fmt_sigfig(min(Pheophytin, na.rm = TRUE)),
    Pheophytin_sd = fmt_sigfig(sd(Pheophytin, na.rm = TRUE)),
    Pheophorbide_mean = fmt_sigfig(mean(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_median = fmt_sigfig(median(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_max = fmt_sigfig(max(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_min = fmt_sigfig(min(Pheophorbide, na.rm = TRUE)),
    Pheophorbide_sd = fmt_sigfig(sd(Pheophorbide, na.rm = TRUE))
  )

# Combine and order rows
DegSummary <- bind_rows(DegSummaryAll, DegSummary, DegSummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
DegSummary %>% 
  kable("html", caption = "Summary Statistics for Pheophytin and Pheophorbide by Snow Color (ppb)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

# Box Plots {.tabset}

## All Pigments

```{r}
boxplotAll <- dfPig_long %>%
  ggplot(aes(x = Pigment, y = Concentration, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotAll

ggsave(boxplotAll, filename = "boxplotAll.png", width = 7, height = 5, bg = "white")
```

## Astaxanthin

```{r}
boxplotAstax <- dfPig %>%
  ggplot(aes(x = Region, y = Astaxanthin, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotAstax

ggsave(boxplotAstax, filename = "boxplotAstax.png", width = 7, height = 5, bg = "white")
```

## Beta Carotene

```{r}
boxplotBetaC <- dfPig %>%
  ggplot(aes(x = Region, y = BetaC, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotBetaC

ggsave(boxplotBetaC, filename = "boxplotBetaC.png", width = 7, height = 5, bg = "white")
```

## ChlA

```{r}
boxplotChlA <- dfPig %>%
  ggplot(aes(x = Region, y = ChlA, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotChlA

ggsave(boxplotChlA, filename = "boxplotChlA.png", width = 7, height = 5, bg = "white")
```

## ChlB

```{r}
boxplotChlB <- dfPig %>%
  ggplot(aes(x = Region, y = ChlB, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotChlB

ggsave(boxplotChlB, filename = "boxplotChlB.png", width = 7, height = 5, bg = "white")
```

## Lutein

```{r}
boxplotLutein <- dfPig %>%
  ggplot(aes(x = Region, y = Lutein, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotLutein

ggsave(boxplotLutein, filename = "boxplotLutein.png", width = 7, height = 5, bg = "white")
```

## Pheophorbide

```{r}
boxplotPheophorbide <- dfPig %>%
  ggplot(aes(x = Region, y = Pheophorbide, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotPheophorbide

ggsave(boxplotPheophorbide, filename = "boxplotPheophorbide.png", width = 7, height = 5, bg = "white")
```

## Pheophytin

```{r}
boxplotPheophytin <- dfPig %>%
  ggplot(aes(x = Region, y = Pheophytin, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.9,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Concentration (ppb)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotPheophytin

ggsave(boxplotPheophytin, filename = "boxplotPheophytin.png", width = 7, height = 5, bg = "white")
```

# Pairs Plots - Pigments

```{r, warning = FALSE, message = FALSE}
corplotAll <- ggpairs(
  dfPig,  
  columns = c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAll

ggsave(corplotAll, filename = "corplotAll.png", width = 7, height = 7, bg = "white")
```

# Heat Maps - Pigments {.tabset}

## All Sites (All Samples)
```{r}
cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
pigment_vars <- c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin")

# List of locations
locations <- unique(dfPig$LocAbbrev)

# Function to calculate correlation matrices for each location
get_corr_data <- function(location) {
  data_sub <- dfPig %>%
    filter(LocAbbrev == location) %>%
    select(all_of(pigment_vars)) %>%
    na.omit()

  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p

  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)

  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Location <- location

  return(merged)
}

# Apply across locations
heatmap_data <- do.call(rbind, lapply(locations, get_corr_data))

# Factor levels for correct ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = pigment_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = pigment_vars)

# Create label
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Remove upper triangle + diagonal if desired (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot
heatmapPigAll <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Location) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank()) 

heatmapPigAll

ggsave(heatmapPigAll, filename = "heatmapPigAll.png", width = 11.5, height = 7, bg = "white")

```

## All Sites (Algae Samples)
```{r}
cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
pigment_vars <- c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin")

# List of locations
locations <- unique(dfPig$LocAbbrev)

# Function to calculate correlation matrices for each location
get_corr_data <- function(location) {
  data_sub <- dfPig %>%
    filter(Algae == "algae") %>% 
    filter(LocAbbrev == location) %>%
    select(all_of(pigment_vars)) %>%
    na.omit()

  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p

  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)

  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Location <- location

  return(merged)
}

# Apply across locations
heatmap_data <- do.call(rbind, lapply(locations, get_corr_data))

# Factor levels for correct ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = pigment_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = pigment_vars)

# Create label
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Remove upper triangle + diagonal if desired (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot
heatmapPigAlgae <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Location) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank()) 

heatmapPigAlgae

ggsave(heatmapPigAlgae, filename = "heatmapPigAlgae.png", width = 11.5, height = 7, bg = "white")

```

## All Sites (Clean Samples)
```{r}
cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
pigment_vars <- c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin")

# List of locations
locations <- unique(dfPig$LocAbbrev)

# Function to calculate correlation matrices for each location
get_corr_data <- function(location) {
  data_sub <- dfPig %>%
    filter(Algae == "clean") %>% 
    filter(LocAbbrev == location) %>%
    select(all_of(pigment_vars)) %>%
    na.omit()

  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p

  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)

  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Location <- location

  return(merged)
}

# Apply across locations
heatmap_data <- do.call(rbind, lapply(locations, get_corr_data))

# Factor levels for correct ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = pigment_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = pigment_vars)

# Create label
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Remove upper triangle + diagonal if desired (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot
heatmapPigClean <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Location) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank()) 

heatmapPigClean

ggsave(heatmapPigClean, filename = "heatmapPigClean.png", width = 11.5, height = 7, bg = "white")

```

## Antarctic Peninsula (Red vs. Green)
```{r}

cor.mtest <- function(mat, method = "spearman") {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = method)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  return(list(p = p.mat))
}

# Define pigment variables
pigment_vars <- c("Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin")

# Filter data once for Antarctic Peninsula red and green samples
df_filtered <- dfPig %>%
  filter(Location == "Antarctic Peninsula",
         Color %in% c("red", "green"))

# Unique colors to iterate over
colors <- unique(df_filtered$Color)

# Function to calculate correlation matrices for each color group
get_corr_data <- function(color_group) {
  data_sub <- df_filtered %>%
    filter(Color == color_group) %>%
    select(all_of(pigment_vars)) %>%
    na.omit()
  
  cor_matrix <- cor(data_sub, method = "spearman")
  p_matrix <- cor.mtest(data_sub, method = "spearman")$p
  
  cor_long <- reshape2::melt(cor_matrix)
  p_long <- reshape2::melt(p_matrix)
  
  merged <- merge(cor_long, p_long, by = c("Var1", "Var2"))
  colnames(merged) <- c("Variable1", "Variable2", "Spearman_r", "p_value")
  merged$Color <- color_group
  
  return(merged)
}

# Get correlation data for each color
heatmap_data_list <- lapply(colors, get_corr_data)
heatmap_data <- do.call(rbind, heatmap_data_list)

# Set factor levels for ordering
heatmap_data$Variable1 <- factor(heatmap_data$Variable1, levels = pigment_vars)
heatmap_data$Variable2 <- factor(heatmap_data$Variable2, levels = pigment_vars)

# Create labels for plot
heatmap_data$label <- paste0(round(heatmap_data$Spearman_r, 2), "\n(p=", round(heatmap_data$p_value, 3), ")")
heatmap_data$significant <- heatmap_data$p_value < 0.05

# Keep only lower triangle (optional)
heatmap_data <- heatmap_data %>%
  filter(as.numeric(Variable1) > as.numeric(Variable2))

# Plot faceted by Color
heatmapPigAPRG <- ggplot(heatmap_data, aes(x = Variable1, y = Variable2, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = subset(heatmap_data, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  facet_wrap(~Color) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank())
  
heatmapPigAPRG
ggsave(heatmapPigAPRG, filename = "heatmapPigAPRG.png", width = 9, height = 5, bg = "white")
```

# Principal Components Analysis - Pigments
From the above plots, it seems as though pigment data are relatively linear. I am going to proceed with a PCA ordination over a nonparametric ordination, like nMDS. 

## Standardize Data & Run PCA
Below, I use prcomp to standardize pigment data and run a PCA. Even though the pigment concentrations are on the same scale, some pigments are present in much higher concentrations than others. Because of this, I will standardize pigments so they are all on the same scale.

PC1 explains 55.11% of the variance in pigment concentration. PC2 explains 21.60% of the variance in pigment concentration. Collectively, these two components explain 76.71% of the variance, and I will likely only include these two PCs in my visualizations. 

```{r}
Pigments <- dfAnion[9:15]

# Standardize
PigPCA <- prcomp(Pigments, scale = TRUE)

summary(PigPCA)
```

## Biplots

```{r}
biplot(PigPCA)
```

## PC Interpretation and Exploration 

```{r}
PigPCA$rotation[,1:3]
```

## Ordination Plots {.tabset}
```{r}
## Add PC1 &2 to dataset
dfAnion$PC1 <- PigPCA$x[,1]
dfAnion$PC2 <- PigPCA$x[,2]
dfAnion$PC3 <- PigPCA$x[,3]
```

```{r}
ordplotPigColor <- dfAnion %>% 
  ggplot(aes(x = PC1, y = PC2, color = Color, fill = Color)) + 
  geom_point(alpha = 0.7, size = 2.5) + 
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(x = "PC1 (55.11%)", 
       y = "PC2 (21.60%)",
       fill = NULL,
       color = NULL) + 
  theme_minimal() + 
  plot_theme

ordplotPigColor

ggsave(ordplotPigColor, filename = "ordplotPigColor.png", width = 7, height = 5, bg = "white")
```

## Fit Anions to Ordination
In this section, I am going to fit the nutrient variables onto the PCA ordination.

None of the anions fit onto the ordination 
```{r}
anions <- dfAnion[16:20]

anionsScaled <- scale(anions)

pigAnionEnvfit <- envfit(PigPCA, anionsScaled)
pigAnionEnvfit
```

# Pairs Plots - Pigments and Anions

## Photosynthetic

```{r, warning = FALSE, message = FALSE}
corplotAPChl <- dfAnion %>% 
  ggpairs(  
  columns = c("ChlA", "ChlB", "Fo", "Cl", "NO2NO3", "PO43", "SO42"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE, method = "spearman"))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAPChl

ggsave(corplotAPChl, filename = "corplotAPChl.png", width = 7, height = 7, bg = "white")
```

## Photoprotective

```{r, warning = FALSE, message = FALSE}
corplotAPPhotoprotective <- dfAnion %>% 
  ggpairs(  
  columns = c("Astaxanthin", "Lutein", "BetaC", "Fo", "Cl", "NO2NO3", "PO43", "SO42"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE, method = "spearman"))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAPPhotoprotective

ggsave(corplotAPPhotoprotective, filename = "corplotAPPhotoprotective.png", width = 7, height = 7, bg = "white")
```

## Degradation

```{r, warning = FALSE, message = FALSE}
corplotAPDegradation <- dfAnion %>% 
  ggpairs(  
  columns = c("Pheophorbide", "Pheophytin", "Fo", "Cl", "NO2NO3", "PO43", "SO42"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE, method = "spearman"))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAPDegradation

ggsave(corplotAPDegradation, filename = "corplotAPDegradation.png", width = 7, height = 7, bg = "white")
```

# Heat Maps - Pigments & Anions {.tabset}

```{r}
# Define pigment and anion variables
pigment_vars <- c("ChlA", "ChlB", "Astaxanthin", "Lutein", "BetaC", "Pheophorbide", "Pheophytin")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Correlation function
get_pigment_anion_corr <- function(data, group_name) {

  cor_mat <- rcorr(as.matrix(data), type = "spearman")

  r_vals <- cor_mat$r
  p_vals <- cor_mat$P

  expand.grid(
    Pigment = pigment_vars,
    Anion   = anion_vars,
    stringsAsFactors = FALSE
  ) %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Anion, Pigment],
      p_value    = p_vals[Anion, Pigment],
      significant = p_value < 0.05,
      label = paste0(
        round(Spearman_r, 2),
        "\n(p=", round(p_value, 3), ")"
      ),
      Group = group_name
    ) %>%
    ungroup()
}

# Red & Green (separate) 
cor_color <- dfAnion %>%
  filter(Color %in% c("red", "green")) %>%
  select(Color, Algae, all_of(pigment_vars), all_of(anion_vars)) %>%
  na.omit() %>%
  group_split(Color) %>%
  map_dfr(~get_pigment_anion_corr(
    data = .x %>% select(-Color, -Algae),
    group_name = unique(.x$Color)
  ))

# Algae (red + green pooled)
cor_algae <- dfAnion %>%
  filter(Algae == "algae") %>%
  select(all_of(pigment_vars), all_of(anion_vars)) %>%
  na.omit() %>%
  get_pigment_anion_corr("algae")

# All samples (red + green + clean) 
cor_all <- dfAnion %>%
  select(all_of(pigment_vars), all_of(anion_vars)) %>%
  na.omit() %>%
  get_pigment_anion_corr("all")

# Combine 
cor_results_all <- bind_rows(cor_color, cor_algae, cor_all)

# Factor ordering
cor_results_all$Pigment <- factor(cor_results_all$Pigment, levels = pigment_vars)
cor_results_all$Anion   <- factor(cor_results_all$Anion, levels = anion_vars)
cor_results_all$Group   <- factor(
  cor_results_all$Group,
  levels = c("red", "green", "algae", "all")
)

# Plot 
heatmapPigAnionAP <- ggplot(
  cor_results_all,
  aes(x = Pigment, y = Anion, fill = Spearman_r)
) +
  geom_tile() +
  geom_tile(
    data = filter(cor_results_all, significant),
    color = "black",
    linewidth = 0.75
  ) +
  geom_text(aes(label = label), size = 3) +
  facet_wrap(~Group, ncol = 1) +
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0,
    name = "Spearman r"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

heatmapPigAnionAP

# Save 
ggsave(
  heatmapPigAnionAP,
  filename = "heatmapPigAnionAP.png",
  width = 7,
  height = 10,
  bg = "white"
)
```

# Cell Density, Volume & Cellular Pigment Concentration

## Pairs Plots 

```{r, warning = FALSE, message = FALSE}
corplotAPCell <- dfPigCell %>%  
  ggpairs(  
  columns = c("CellDensity", "CellVolume", "Astaxanthin", "Lutein", "ChlA", "ChlB", "BetaC", "Pheophorbide", "Pheophytin"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE, method = "spearman"))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAPCell

ggsave(corplotAPCell, filename = "corplotAPCell.png", width = 7, height = 7, bg = "white")
```

## Heat Maps {.tabset}

```{r}
# Define pigment and cell variables
pigment_vars <- c("ChlA", "ChlB", "Astaxanthin", "Lutein", 
                  "BetaC", "Pheophorbide", "Pheophytin")
cell_vars <- c("CellDensity", "CellVolume")

# ---- Build datasets for each tier ----

# All samples combined (red + green + clean)
df_all <- dfPigCell %>%
  filter(Location == "Antarctic Peninsula") %>%
  mutate(Tier = "All samples") %>%
  select(Tier, all_of(pigment_vars), all_of(cell_vars)) %>%
  na.omit()

# Green only
df_green <- dfPigCell %>%
  filter(Location == "Antarctic Peninsula", Color == "green") %>%
  mutate(Tier = "Green") %>%
  select(Tier, all_of(pigment_vars), all_of(cell_vars)) %>%
  na.omit()

# Red only
df_red <- dfPigCell %>%
  filter(Location == "Antarctic Peninsula", Color == "red") %>%
  mutate(Tier = "Red") %>%
  select(Tier, all_of(pigment_vars), all_of(cell_vars)) %>%
  na.omit()

# Combine all tiers into one dataframe
df_combined <- bind_rows(df_all, df_green, df_red)

# ---- Function to compute pigment–cell Spearman correlations ----
get_pigment_Cell_corr <- function(data, tier_name) {

  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")

  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  pairs <- expand.grid(
    Pigment = pigment_vars,
    Cell    = cell_vars,
    stringsAsFactors = FALSE
  )

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Cell, Pigment],
      p_value    = p_vals[Cell, Pigment],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      Tier = tier_name
    ) %>%
    ungroup()

  return(results)
}

# ---- Compute correlations for each tier ----
cor_results <- df_combined %>%
  group_split(Tier) %>%
  map_dfr(~get_pigment_Cell_corr(.x, unique(.x$Tier)))

# Preserve axis and facet order
cor_results$Pigment <- factor(cor_results$Pigment, levels = pigment_vars)
cor_results$Cell    <- factor(cor_results$Cell, levels = cell_vars)
cor_results$Tier    <- factor(
  cor_results$Tier,
  levels = c("All samples", "Green", "Red")
)

# ---- Plot heatmap ----
heatmapPigCellAP <- ggplot(
  cor_results,
  aes(x = Pigment, y = Cell, fill = Spearman_r)
) +
  geom_tile() +
  geom_tile(
    data = filter(cor_results, significant),
    color = "black",
    linewidth = 0.75
  ) +
  geom_text(aes(label = label), size = 3) +
  facet_wrap(~Tier, ncol = 1) +
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0,
    name = "Spearman r"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

heatmapPigCellAP

# Save
ggsave(
  heatmapPigCellAP,
  filename = "heatmapPigCellAP.png",
  width = 7,
  height = 9,
  bg = "white"
)
```


# Publication Figures

## Algal Properties by Region {.tabset}

### Cell Density
```{r}
boxplotCellDensityPub <- dfPigCell %>% 
  mutate(
    Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = CellDensity, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Cell Density" ~ "(cells·mL"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 3e+04, 2e+05, 5e+05, 1e+06, 2e+06, 4e+06)) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotCellDensityPub
```

### Cell Volume
```{r}
boxplotCellVolumePub <- dfPigCell %>% 
  filter(CellVolume > 0) %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = CellVolume, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Cell Volume" ~ "(μm"^{3}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 1000, 2500, 5000, 10000, 20000, 30000),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotCellVolumePub
```

### Astaxanthin
```{r}
boxplotAstaxPub <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = Astaxanthin, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Astaxanthin" ~ "(ng·cell"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 0.5, 2.5, 5, 10, 20, 30, 40, 50),
               labels = function(x) ifelse(x < 5, x, as.integer(x))) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotAstaxPub
```

### ChlA
```{r}
boxplotChlAPub <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = ChlA, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Chlorophyll a" ~ "(ng·cell"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 2.5, 10, 25, 50, 100, 150, 200),
               labels = function(x) ifelse(x < 5, x, as.integer(x))) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotChlAPub
```

### ChlB
```{r}
boxplotChlBPub <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = ChlB, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Chlorophyll b" ~ "(ng·cell"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 2.5, 10, 25, 50, 100, 150, 200),
               labels = function(x) ifelse(x < 5, x, as.integer(x))) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotChlBPub
```

### Lutein
```{r}
boxplotLuteinPub <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = Lutein, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Lutein" ~ "(ng·cell"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 2.5, 10, 25, 50, 100, 150, 200),
               labels = function(x) ifelse(x < 5, x, as.integer(x))) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotLuteinPub
```

### BetaC
```{r}
boxplotBetaCPub <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = BetaC, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("β-carotene" ~ "(ng·cell"^{-1}*")"),
    fill = NULL,
    color = NULL) +
 scale_y_sqrt(breaks = c(0, 2.5, 10, 25, 50, 100, 150, 200),
               labels = function(x) ifelse(x < 5, x, as.integer(x))) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotBetaCPub
```

### Pheophytin
```{r}
boxplotPheophytinPub <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = Pheophytin, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Pheophytin" ~ "(ng·cell"^{-1}*")"),
    fill = NULL,
    color = NULL) +
scale_y_sqrt(breaks = c(0, 2.5, 10, 25, 50, 100, 150, 200),
               labels = function(x) ifelse(x < 5, x, as.integer(x))) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotPheophytinPub
```

### Pheophorbide
```{r}
boxplotPheophorbidePub <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = Pheophorbide, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Pheophorbide" ~ "(ng·cell"^{-1}*")"),
    fill = NULL,
    color = NULL) +
scale_y_sqrt(breaks = c(0, 0.25, 1, 2.5, 5, 10, 25),
               labels = function(x) ifelse(x < 5, x, as.integer(x))) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotPheophorbidePub
```

### Shared Legend
```{r}
legendPlot <- dfPig %>% 
  mutate(Site = factor(Region, levels = c("North", "South")),
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Site, y = Pheophorbide, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Pheophorbide" ~ "(ng·cell"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 10, 50, 150, 300, 500, 1000, 1500),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    )) +
  theme(
    legend.position = "bottom",
    legend.justification = "center",  # center the legend
    legend.box.just = "center")

legendPlot
```

### Combined Plot
```{r}
# Extract legend
sharedLegend <- cowplot::get_legend(legendPlot)

combinedPlot_AlgalProp <- cowplot::plot_grid(
  boxplotChlAPub, boxplotAstaxPub, boxplotLuteinPub,
  boxplotBetaCPub, boxplotPheophytinPub,
  boxplotPheophorbidePub,
  ncol = 3, nrow = 2, 
  rel_heights = c(1,1), 
  rel_widths = c(1,1),
  align = "hv",
  axis = "tblr",
  labels = NULL
)

combinedPlot_AlgalProp


ggsave(combinedPlot_AlgalProp, filename = "combinedPlot_AlgalProp.png", width = 7, height = 3.75, bg = "white")
```

### Cell Density formatting
```{r}
combinedPlot_CD <- cowplot::plot_grid(
  boxplotCellDensityPub,
  ncol = 3, nrow = 2, 
  rel_heights = c(1,1), 
  rel_widths = c(1,1),
  align = "hv",
  axis = "tblr",
  labels = NULL
)

combinedPlot_CD


ggsave(combinedPlot_CD, filename = "combinedPlot_CD.png", width = 7, height = 3.75, bg = "white")
```

## Algal Properties by Color {.tabset}

### Cell Density
```{r}
boxplotCellDensityPub_Color <- dfPigCell %>% 
  mutate(Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = CellDensity, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Cell Density" ~ "(cells·mL"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 3e+04, 2e+05, 5e+05, 1e+06, 2e+06, 4e+06)) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotCellDensityPub_Color
```

### Cell Volume
```{r}
boxplotCellVolumePub_Color <- dfPigCell %>% 
  filter(CellVolume > 0) %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = CellVolume, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Cell Volume" ~ "(μm"^{3}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 1000, 2500, 5000, 10000, 20000, 30000),
               labels = label_comma()) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotCellVolumePub_Color
```

### Astaxanthin
```{r}
boxplotAstaxPub_Color <- dfPig %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = Astaxanthin, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Astaxanthin" ~ "(ng·cell"^{-1}*")"),
    fill = NULL,
    color = NULL) +
   scale_y_sqrt(breaks = c(0, 0.5, 2.5, 5, 10, 20, 30, 40, 50),
               labels = function(x) ifelse(x < 5, x, as.integer(x))) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotAstaxPub_Color
```

### ChlA
```{r}
boxplotChlAPub_Color <- dfPig %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = ChlA, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Chlorophyll a" ~ "(ng·cell"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 2.5, 10, 25, 50, 100, 150, 200),
               labels = function(x) ifelse(x < 5, x, as.integer(x))) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotChlAPub_Color
```

### ChlB
```{r}
boxplotChlBPub_Color <- dfPig %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = ChlB, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Chlorophyll b" ~ "(ng·cell"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 2.5, 10, 25, 50, 100, 150, 200),
               labels = function(x) ifelse(x < 5, x, as.integer(x))) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotChlBPub_Color
```

### Lutein
```{r}
boxplotLuteinPub_Color <- dfPig %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = Lutein, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Lutein" ~ "(ng·cell"^{-1}*")"),
    fill = NULL,
    color = NULL) +
  scale_y_sqrt(breaks = c(0, 2.5, 10, 25, 50, 100, 150, 200),
               labels = function(x) ifelse(x < 5, x, as.integer(x))) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotLuteinPub_Color
```

### BetaC
```{r}
boxplotBetaCPub_Color <- dfPig %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = BetaC, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("β-carotene" ~ "(ng·cell"^{-1}*")"),
    fill = NULL,
    color = NULL) +
 scale_y_sqrt(breaks = c(0, 2.5, 10, 25, 50, 100, 150, 200),
               labels = function(x) ifelse(x < 5, x, as.integer(x))) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotBetaCPub_Color
```

### Pheophytin
```{r}
boxplotPheophytinPub_Color <- dfPig %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = Pheophytin, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Pheophytin" ~ "(ng·cell"^{-1}*")"),
    fill = NULL,
    color = NULL) +
scale_y_sqrt(breaks = c(0, 2.5, 10, 25, 50, 100, 150, 200),
               labels = function(x) ifelse(x < 5, x, as.integer(x))) +
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotPheophytinPub_Color
```

### Pheophorbide
```{r}
boxplotPheophorbidePub_Color <- dfPig %>% 
  mutate(
    Color = factor(Color, levels = c("green", "red", "clean"))) %>% 
  ggplot(aes(x = Color, y = Pheophorbide, fill = Color, color = Color)) + 
  geom_boxplot(alpha = 0.9, width = 0.8) + 
  scale_fill_manual(values = colorPal) + 
  scale_color_manual(values = colorPal) + 
  labs(
    x = NULL, 
    y = expression("Pheophorbide" ~ "(ng·cell"^{-1}*")"),
    fill = NULL,
    color = NULL) +
scale_y_sqrt(breaks = c(0, 0.25, 1, 2.5, 5, 10, 25),
               labels = function(x) ifelse(x < 5, x, as.integer(x))) +
  
  theme_minimal() +
  plot_themePub +
  theme(axis.text.x = element_text(margin = margin(t = 5)),
        legend.position = "none") +
  guides(
    fill = guide_legend(
      override.aes = list(linetype = 0, shape = 22),
      keywidth = unit(1, "lines"),   # width of the legend key
      keyheight = unit(1.5, "lines")   # height of the legend key
    ))
  
boxplotPheophorbidePub_Color
```

### Combined Plot
```{r}
# Extract legend
sharedLegend <- cowplot::get_legend(legendPlot)

combinedPlot_AlgalProp_Color <- cowplot::plot_grid(
  boxplotChlAPub_Color, boxplotAstaxPub_Color, boxplotLuteinPub_Color,
  boxplotBetaCPub_Color, boxplotPheophytinPub_Color,
  boxplotPheophorbidePub_Color,
  ncol = 3, nrow = 2, 
  rel_heights = c(1,1), 
  rel_widths = c(1,1),
  align = "hv",
  axis = "tblr",
  labels = NULL
)

combinedPlot_AlgalProp_Color


ggsave(combinedPlot_AlgalProp_Color, filename = "combinedPlot_AlgalProp_Color.png", width = 7, height = 3.75, bg = "white")
```

### Cell Density formatting
```{r}
combinedPlot_CD_Color <- cowplot::plot_grid(
  boxplotCellDensityPub_Color,
  ncol = 3, nrow = 2, 
  rel_heights = c(1,1), 
  rel_widths = c(1,1),
  align = "hv",
  axis = "tblr",
  labels = NULL
)

combinedPlot_CD_Color


ggsave(combinedPlot_CD_Color, filename = "combinedPlot_CD_Color.png", width = 7, height = 3.75, bg = "white")
```

