---
title: "Data Visualization - Cell Density (AP)"
author: "Ellie Ryan"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: flatly
    highlight: zenburn
    toc: true
    toc_float: true
    toc_collapsed: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Created:** 6/30/2025

**Last Modified:** 7/8/2025

```{r, message=FALSE, warning=FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(stringr)
library(kableExtra)
library(scales)
library(GGally)
library(patchwork)
library(Hmisc)
library(reshape2)
library(forcats)
library(ggtext)
```

```{r}

colorPal <- c(
  "clean" = "#769a9c",
  "red" = "#933410",
  "green" = "#326e35"
)

algaePal<- c(
  "clean" = "#a5bdc0",
  "algae" = "#735a69"
)

plot_theme <- theme(
  axis.title = element_text(size = 12, face = "bold"), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  plot.margin = margin(20, 20, 20, 20)
)

plot_theme2 <- theme(
  axis.title = element_text(size = 12, face = "bold"), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  axis.text.x = element_text(angle = 45, hjust = 1), 
  plot.margin = margin(20, 20, 20, 20)
)

pairsplot_theme <- theme(
  text = element_text(size = 10),  # Adjust variable name text size
  axis.text = element_text(size = 8),  # Adjust axis text size
  plot.title = element_text(size = 12, face = "bold"),  # Adjust plot title size
  legend.text = element_text(size = 12))  # Adjust legend text size
```

# Overview
The purpose of this `.Rmd` is to visualize cell density by site and color within the Antarctic Peninsula, and obtain basic summary statistics associated with cell density in order to inform my next steps in analysis. Before proceeding, This dataset exclusively features samples collected from the Antarctic Peninsula. These observations are broken down by `Color`, or the apparent snow color at time of sample collection. To keep things simple, there are only three levels to the `Color` category: red, green and clean. The breakdown of samples collected are as follows:

-   Antarctic Peninsula:
    -   84 samples
        -   33 red
        -   33 green
        -   18 clean
        
From outlier analysis conducted using figures generated in the first run-through of this script, the following outliers were identified for removal:

-   2025-02-19-SI-algae3
-   2025-02-24-DA-algae2
-   2025-02-24-DA-algae6
-   2025-03-07-NH-algae3

See the Thesis Figures (Preliminary) doc for justification for removal of each.

Based on the results from the cell density analysis for all sites, it seems as though cell density and biovolume concentration are closely related. I will proceed by calculating figures for both, but am going to only focus on biovolume in leiu of biovolume and cell length for this analysis and from this point forward. 

## Data Wrangling

```{r}
df <- read.csv("CompiledData.csv")
str(df)
```

```{r}
# Format date column 
df <- df %>%
  mutate(
    Date = as.Date(Date, format = "%m/%d/%y")) %>% 
  filter(Location == "Antarctic Peninsula")
```

```{r}
str(df)
```

```{r}
outliers <- c(
  "2025-02-19-SI-algae3",
  "2025-02-24-DA-algae2",
  "2025-02-24-DA-algae6",
  "2025-03-07-NH-algae3")

df <- df %>% 
  filter(!SampleID %in% outliers)
```

### Cell Density/Biovolume Comparison

Below, I calculate biovolume and total cellular biovolume and store these values in `dfBio` in order to compare these metrics with cell density and use biovolume in place of cell length. 

**Biovolume:** This represents the volume of each cell, assuming a spherical shape for snow algae (µm^3^ or mm^3^). This method was first categorized by Hillebrand et al., 1999, *Biovolume Calculation for Pelagic and Benthic Microalgae*. The equation is as follows:

-   V~bio~ = 4/3πr^3^

**Total Biovolume:** This represents the total volume of cells per unit volume of sample, or biovolume density (µm^3^/ml or mm^3^/ml). The equation is as follows:

-   TV~bio~ = V~bio~ \* cells/ml

```{r}
dfBio <- df %>% 
  select(c("SampleID", "Date", "Location", "Site", "SiteLat", "Color", "CellDensity", "CellLength", Fo:SO42 )) %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  mutate(NO2NO3 = NO2 + NO3) %>% 
  relocate(Algae, .after = Color)%>% 
  select(-NO2, -NO3)
  
dfBio <- dfBio %>% 
  mutate(
    # Calulate biovolume (µm^3)
    Biovolume = (pi/6)*(CellLength^3),
    # Calculate total biovolume concentration (µm^3/mL)
    TotalBiovolume = Biovolume * CellDensity 
  ) %>% 
  relocate(Biovolume, .after = CellLength) %>% 
  relocate(TotalBiovolume, .after = CellDensity)

# Create a new factor combining Site and Date, ordered by SiteLat and Date
dfBio <- dfBio %>%
  mutate(SiteDate = paste(Site, Date, sep = " - ")) %>%
  arrange(SiteLat, Date) %>%
  mutate(SiteDate = factor(SiteDate, levels = unique(SiteDate)))%>% 
  relocate(SiteDate, .after = Site)
```


I am also going to make another dataframe, `dfIndex`, that includes the index scores for the wildlife, rock, and oceanic indices. I will use this df to see if there are any trends in cell density/biovolume and index scores, as well as to see whether any trends between anion concentration and cell density/biovolume pop out when grouped by index score. 

```{r}
df2 <- read.csv("CompiledDataAp_Indices.csv")

df2 <- df2 %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%y"))

df2 <- df2 %>% 
  filter(!SampleID %in% outliers)

dfIndex <- df2 %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  mutate(NO2NO3 = NO2 + NO3) %>% 
  relocate(Algae, .after = Color) %>% 
  relocate(NO2NO3, .after = SO42) %>% 
  select(-NO2, -NO3)
  
dfIndex <- dfIndex %>% 
  mutate(
    # Calulate biovolume (µm^3)
    Biovolume = (pi/6)*(CellLength^3),
    # Calculate total biovolume concentration (µm^3/mL)
    TotalBiovolume = Biovolume * CellDensity 
  ) %>% 
  relocate(Biovolume, .after = CellLength) %>% 
  relocate(TotalBiovolume, .after = CellDensity)

# Create a new factor combining Site and Date, ordered by SiteLat and Date
dfIndex <- dfIndex %>%
  mutate(SiteDate = paste(Site, Date, sep = " - ")) %>%
  mutate(SiteDate = factor(SiteDate, levels = unique(SiteDate)))%>% 
  relocate(SiteDate, .after = Site)

# Group sites by location along the Peninsula
dfIndex <- dfIndex %>% 
  mutate(SiteLoc = case_when(
    Site %in% c("Horseshoe Island", "Stonington Island") ~ "South",
    Site %in% c("Pléneau Island", "Damoy Point") ~ "South Central",
    Site %in% c("Neko Harbour", "Danco Island", "Orne Harbour") ~ "North Central", 
    Site %in% c("Pendulum Cove, Deception Island") ~ "North",
  )) %>% 
  relocate(SiteLoc, .after = Site)
```

# Summary Statistics {.tabset}
Similar to the cell density analysis for all sites, I am going to create summary tables that depict summary stats for cell density by snow color in the AP samples. I will also compare by site. 

## Cell Density (Color)

Below is a summary stats table for cell density, grouped by `Color`. The `all` level of the `Color` classification includes red, green and clean snow observations.

```{r}
# Cell Density 
dfCSummary <- dfBio %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    mean = comma(round(mean(CellDensity, na.rm = TRUE))),
    median = comma(round(median(CellDensity, na.rm = TRUE))),
    max = comma(max(CellDensity, na.rm = TRUE)),
    min = comma(min(CellDensity, na.rm = TRUE)),
    sd = comma(round(sd(CellDensity, na.rm = TRUE))),
    .groups = "drop")

# Add "All" for each site 
dfCSummaryAll <- dfBio %>% 
  summarise(
    Color = "all",
    n = n(),
    mean = comma(round(mean(CellDensity, na.rm = TRUE))),
    median = comma(round(median(CellDensity, na.rm = TRUE))),
    max = comma(max(CellDensity, na.rm = TRUE)),
    min = comma(min(CellDensity, na.rm = TRUE)),
    sd = comma(round(sd(CellDensity, na.rm = TRUE))),
    .groups = "drop"
  )

# Combine 
dfCSummary <- bind_rows(dfCSummaryAll, dfCSummary) %>% 
  arrange(match(Color, c("green", "red", "clean", "all")))

# Format table using kableExtra
dfCSummary %>% 
  kable("html", caption = "Summary Statistics for Cell Density by Snow Color in the Antarctic Peninsula (cells/mL) ", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Cell Density (Color & Site)

Below is a summary stats table for cell density, grouped by `Color` and sampling site. The `all` level of the `Color` classification includes red, green and clean snow observations.

```{r}
# Cell Density 
dfCSummarySite <- dfBio %>% 
  group_by(Color, Site) %>% 
  summarise(
    n = n(),
    mean = comma(round(mean(CellDensity, na.rm = TRUE))),
    median = comma(round(median(CellDensity, na.rm = TRUE))),
    max = comma(max(CellDensity, na.rm = TRUE)),
    min = comma(min(CellDensity, na.rm = TRUE)),
    sd = comma(round(sd(CellDensity, na.rm = TRUE))),
    .groups = "drop")

# Add "All" for each site 
dfCSummaryAllSite <- dfBio %>% 
  group_by(Site) %>% 
  summarise(
    Color = "all",
    n = n(),
    mean = comma(round(mean(CellDensity, na.rm = TRUE))),
    median = comma(round(median(CellDensity, na.rm = TRUE))),
    max = comma(max(CellDensity, na.rm = TRUE)),
    min = comma(min(CellDensity, na.rm = TRUE)),
    sd = comma(round(sd(CellDensity, na.rm = TRUE))),
    .groups = "drop"
  )

# Combine 
dfCSummarySite <- bind_rows(dfCSummarySite, dfCSummaryAllSite) %>% 
  arrange(Site, match(Color, c("green", "red", "clean", "all")))

# Format table using kableExtra
dfCSummarySite %>% 
  kable("html", caption = "Summary Statistics for Cell Density by Snow Color at Each Sampling Location in the Antarctic Peninsula (cells/mL) ", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```


## Biovolume Concentration (Color)

Below is a summary stats table for biovolume concentration, grouped by `Color`. The `all` level of the `Color` classification includes red, green and clean snow observations.

```{r}
# Total Biovolume 
dfCSummaryTotBio <- dfBio %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    mean = scientific(round(mean(TotalBiovolume, na.rm = TRUE))),
    median = scientific(round(median(TotalBiovolume, na.rm = TRUE))),
    max = scientific(max(TotalBiovolume, na.rm = TRUE)),
    min = scientific(min(TotalBiovolume, na.rm = TRUE)),
    sd = scientific(round(sd(TotalBiovolume, na.rm = TRUE))),
    .groups = "drop")

# Add "All" for each site 
dfCSummaryTotBioAll <- dfBio %>% 
  summarise(
    Color = "all",
    n = n(),
    mean = scientific(round(mean(TotalBiovolume, na.rm = TRUE))),
    median = scientific(round(median(TotalBiovolume, na.rm = TRUE))),
    max = scientific(max(TotalBiovolume, na.rm = TRUE)),
    min = scientific(min(TotalBiovolume, na.rm = TRUE)),
    sd = scientific(round(sd(TotalBiovolume, na.rm = TRUE))),
    .groups = "drop"
  )

# Combine 
dfCSummaryTotBio <- bind_rows(dfCSummaryTotBioAll, dfCSummaryTotBio) %>% 
  arrange(match(Color, c("green", "red", "clean", "all")))

# Format table using kableExtra
dfCSummaryTotBio %>% 
  kable("html", caption = "Summary Statistics for Cell Biovolume Concentration by Snow Color in the Antarctic Peninsula (um^3/mL)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Biovolume Concentration (Color & Site)

Below is a summary stats table for cell density, grouped by `Color` and sampling site. The `all` level of the `Color` classification includes red, green and clean snow observations.

```{r}
# Total Biovolume 
dfCSummaryTotBioSite <- dfBio %>% 
  group_by(Color, Site) %>% 
  summarise(
    n = n(),
    mean = scientific(round(mean(TotalBiovolume, na.rm = TRUE))),
    median = scientific(round(median(TotalBiovolume, na.rm = TRUE))),
    max = scientific(max(TotalBiovolume, na.rm = TRUE)),
    min = scientific(min(TotalBiovolume, na.rm = TRUE)),
    sd = scientific(round(sd(TotalBiovolume, na.rm = TRUE))),
    .groups = "drop")

# Add "All" for each site 
dfCSummaryTotBioAllSite <- dfBio %>% 
  group_by(Site) %>% 
  summarise(
    Color = "all",
    n = n(),
    mean = scientific(round(mean(TotalBiovolume, na.rm = TRUE))),
    median = scientific(round(median(TotalBiovolume, na.rm = TRUE))),
    max = scientific(max(TotalBiovolume, na.rm = TRUE)),
    min = scientific(min(TotalBiovolume, na.rm = TRUE)),
    sd = scientific(round(sd(TotalBiovolume, na.rm = TRUE))),
    .groups = "drop"
  )

# Combine 
dfCSummaryTotBioSite <- bind_rows(dfCSummaryTotBioSite, dfCSummaryTotBioAllSite) %>% 
  arrange(Site, match(Color, c("green", "red", "clean", "all")))

# Format table using kableExtra
dfCSummaryTotBioSite %>% 
  kable("html", caption = "Summary Statistics for Cell Biovolume Concentration at Each Sampling Location in the Antarctic Peninsula (um^3/mL)", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Cell Biovolume (Color)

Below is a summary stats table for cell biovolume, grouped by `Color`. The `all` level of the `Color` classification includes red, green and clean snow observations.

```{r}
# Cell Density 
dfLSummary <- dfBio %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    mean = comma(round(mean(Biovolume, na.rm = TRUE))),
    median = comma(round(median(Biovolume, na.rm = TRUE))),
    max = comma(max(Biovolume, na.rm = TRUE)),
    min = comma(min(Biovolume, na.rm = TRUE)),
    sd = comma(round(sd(Biovolume, na.rm = TRUE))),
    .groups = "drop")

# Add "All" for each site 
dfLSummaryAll <- dfBio %>% 
  summarise(
    Color = "all",
    n = n(),
    mean = comma(round(mean(Biovolume, na.rm = TRUE))),
    median = comma(round(median(Biovolume, na.rm = TRUE))),
    max = comma(max(Biovolume, na.rm = TRUE)),
    min = comma(min(Biovolume, na.rm = TRUE)),
    sd = comma(round(sd(Biovolume, na.rm = TRUE))),
    .groups = "drop"
  )

# Combine 
dfLSummary <- bind_rows(dfLSummaryAll, dfLSummary) %>% 
  arrange(match(Color, c("green", "red", "clean", "all")))

# Format table using kableExtra
dfLSummary %>% 
  kable("html", caption = "Summary Statistics for Cell Biovolume by Snow Color in the Antarctic Peninsula (µm^3) ", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

## Cell Biovolume (Color & Site)

Below is a summary stats table for cell biovolume, grouped by `Color` and sampling site. The `all` level of the `Color` classification includes red, green and clean snow observations.

```{r}
# Cell Biovolume
dfLSummarySite <- dfBio %>% 
  group_by(Color, Site) %>% 
  summarise(
    n = n(),
    mean = comma(round(mean(Biovolume, na.rm = TRUE))),
    median = comma(round(median(Biovolume, na.rm = TRUE))),
    max = comma(max(Biovolume, na.rm = TRUE)),
    min = comma(min(Biovolume, na.rm = TRUE)),
    sd = comma(round(sd(Biovolume, na.rm = TRUE))),
    .groups = "drop")

# Add "All" for each site 
dfLSummaryAllSite <- dfBio %>% 
  group_by(Site) %>% 
  summarise(
    Color = "all",
    n = n(),
    mean = comma(round(mean(Biovolume, na.rm = TRUE))),
    median = comma(round(median(Biovolume, na.rm = TRUE))),
    max = comma(max(Biovolume, na.rm = TRUE)),
    min = comma(min(Biovolume, na.rm = TRUE)),
    sd = comma(round(sd(Biovolume, na.rm = TRUE))),
    .groups = "drop"
  )

# Combine 
dfLSummarySite <- bind_rows(dfLSummarySite, dfLSummaryAllSite) %>% 
  arrange(Site, match(Color, c("green", "red", "clean", "all")))

# Format table using kableExtra
dfLSummarySite %>% 
  kable("html", caption = "Summary Statistics for Cell Biovolume by Snow Color at Each Sampling Location in the Antarctic Peninsula (µm^3) ", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

# Box Plots 

## Cell Density {.tabset}

### Cell Density (Color)

```{r}
boxplotDensity <- dfBio %>%
  ggplot(aes(x = Color, y = CellDensity, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Density (cells/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme + 
  theme(legend.position = "none")

boxplotDensity

ggsave(boxplotDensity, filename = "boxplotDensity.png", width = 6, height = 4, bg = "white")
```

### Cell Density (Site & Color)

```{r}
boxplotDensitySite <- dfBio %>%
  mutate(Site = fct_reorder(Site, SiteLat)) %>% 
  ggplot(aes(x = Site, y = CellDensity, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Density (cells/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotDensitySite

ggsave(boxplotDensitySite, filename = "boxplotDensitySiteLat.png", width = 7, height = 6, bg = "white")
```

### Cell Density (Site & Date) {.tabset}

#### Pléneau Island
```{r}
boxplotDensityDatePLI <- dfBio %>%
  filter(Site == "Pléneau Island") %>% 
  mutate(Date = factor(Date)) %>% 
  ggplot(aes(x = Date, y = CellDensity, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Density (cells/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotDensityDatePLI

ggsave(boxplotDensityDatePLI, filename = "boxplotDensityDatePLI.png", width = 7, height = 6, bg = "white")
```

#### Neko Harbour
```{r}
boxplotDensityDateNH <- dfBio %>%
  filter(Site == "Neko Harbour") %>% 
  mutate(Date = factor(Date)) %>% 
  ggplot(aes(x = Date, y = CellDensity, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Density (cells/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotDensityDateNH

ggsave(boxplotDensityDateNH, filename = "boxplotDensityDateNH.png", width = 7, height = 6, bg = "white")
```

#### Damoy Point
```{r}
boxplotDensityDateDP <- dfBio %>%
  filter(Site == "Damoy Point") %>% 
  mutate(Date = factor(Date)) %>% 
  ggplot(aes(x = Date, y = CellDensity, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Density (cells/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotDensityDateDP

ggsave(boxplotDensityDateDP, filename = "boxplotDensityDateDP.png", width = 7, height = 6, bg = "white")
```

#### Orne Harbour
```{r}
boxplotDensityDateOH <- dfBio %>%
  filter(Site == "Orne Harbour") %>% 
  mutate(Date = factor(Date)) %>% 
  ggplot(aes(x = Date, y = CellDensity, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Density (cells/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotDensityDateOH

ggsave(boxplotDensityDateOH, filename = "boxplotDensityDateOH.png", width = 7, height = 6, bg = "white")
```

## Biovolume Concentration {.tabset}

### Biovolume Concentration (Color)

```{r}
boxplotTotBio <- dfBio %>%
  ggplot(aes(x = Color, y = TotalBiovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Biovolume Concentration (µm^3/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme + 
  theme(legend.position = "none")

boxplotTotBio

ggsave(boxplotTotBio, filename = "boxplotTotBio.png", width = 6, height = 4, bg = "white")
```

### Biovolume Concentration (Site & Color)

```{r}
boxplotTotBioSite <- dfBio %>%
  mutate(Site = fct_reorder(Site, SiteLat)) %>% 
  ggplot(aes(x = Site, y = TotalBiovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Biovolume Concentration (µm^3/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotTotBioSite

ggsave(boxplotTotBioSite, filename = "boxplotTotBioSiteLat.png", width = 7, height = 6, bg = "white")
```

### Biovolume Concentration (Site & Date) {.tabset}

#### Pléneau Island
```{r}
boxplotTotBioDatePLI <- dfBio %>%
  filter(Site == "Pléneau Island") %>% 
  mutate(Date = factor(Date)) %>% 
  ggplot(aes(x = Date, y = TotalBiovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Biovolume Concentration (µm^3/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotTotBioDatePLI

ggsave(boxplotTotBioDatePLI, filename = "boxplotTotBioDatePLI.png", width = 7, height = 6, bg = "white")
```

#### Neko Harbour
```{r}
boxplotTotBioDateNH <- dfBio %>%
  filter(Site == "Neko Harbour") %>% 
  mutate(Date = factor(Date)) %>% 
  ggplot(aes(x = Date, y = TotalBiovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Biovolume Concentration (µm^3/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotTotBioDateNH

ggsave(boxplotTotBioDateNH, filename = "boxplotTotBioDateNH.png", width = 7, height = 6, bg = "white")
```

#### Damoy Point
```{r}
boxplotTotBioDateDP <- dfBio %>%
  filter(Site == "Damoy Point") %>% 
  mutate(Date = factor(Date)) %>% 
  ggplot(aes(x = Date, y = TotalBiovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Biovolume Concentration (µm^3/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotTotBioDateDP

ggsave(boxplotTotBioDateDP, filename = "boxplotTotBioDateDP.png", width = 7, height = 6, bg = "white")
```

#### Orne Harbour
```{r}
boxplotTotBioDateOH <- dfBio %>%
  filter(Site == "Orne Harbour") %>% 
  mutate(Date = factor(Date)) %>% 
  ggplot(aes(x = Date, y = TotalBiovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Biovolume Concentration (µm^3/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme2

boxplotTotBioDateOH

ggsave(boxplotTotBioDateOH, filename = "boxplotTotBioDateOH.png", width = 7, height = 6, bg = "white")
```

## Biovolume {.tabset}

### Biovolume (Color)

```{r}
boxplotBio <- dfBio %>%
  ggplot(aes(x = Color, y = Biovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Biovolume (µm^3)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme + 
  theme(legend.position = "none")

boxplotBio

ggsave(boxplotBio, filename = "boxplotBio.png", width = 6, height = 4, bg = "white")
```

### Biovolume (Site & Color)

```{r}
boxplotBioSite <- dfBio %>%
  mutate(Site = fct_reorder(Site, SiteLat)) %>% 
  ggplot(aes(x = Site, y = Biovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Biovolume (µm^3)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme2

boxplotBioSite

ggsave(boxplotBioSite, filename = "boxplotBioSiteLat.png", width = 7, height = 6, bg = "white")
```


### Biovolume (Site & Date) {.tabset}

#### Pléneau Island
```{r}
boxplotBioDatePLI <- dfBio %>%
  filter(Site == "Pléneau Island") %>% 
  mutate(Date = factor(Date)) %>% 
  ggplot(aes(x = Date, y = Biovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Biovolume (µm^3)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme2

boxplotBioDatePLI

ggsave(boxplotBioDatePLI, filename = "boxplotBioDatePLI.png", width = 7, height = 6, bg = "white")
```

```{r}
boxplotPLI <- (boxplotDensityDatePLI | boxplotTotBioDatePLI | boxplotBioDatePLI) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

boxplotPLI

ggsave(boxplotPLI, filename = "boxplotPLI.png", width = 9, height = 4.5, bg = "white")
```


#### Neko Harbour
```{r}
boxplotBioDateNH <- dfBio %>%
  filter(Site == "Neko Harbour") %>% 
  mutate(Date = factor(Date)) %>% 
  ggplot(aes(x = Date, y = Biovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Biovolume (µm^3)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme2

boxplotBioDateNH

ggsave(boxplotBioDateNH, filename = "boxplotBioDateNH.png", width = 7, height = 6, bg = "white")
```

```{r}
boxplotNH <- (boxplotDensityDateNH | boxplotTotBioDateNH | boxplotBioDateNH) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

boxplotNH

ggsave(boxplotNH, filename = "boxplotNH.png", width = 9, height = 4.5, bg = "white")
```

#### Damoy Point
```{r}
boxplotBioDateDP <- dfBio %>%
  filter(Site == "Damoy Point") %>% 
  mutate(Date = factor(Date)) %>% 
  ggplot(aes(x = Date, y = Biovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Biovolume (µm^3)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme2

boxplotBioDateDP

ggsave(boxplotBioDateDP, filename = "boxplotBioDateDP.png", width = 7, height = 6, bg = "white")
```

```{r}
boxplotDP <- (boxplotDensityDateDP | boxplotTotBioDateDP | boxplotBioDateDP) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

boxplotDP

ggsave(boxplotDP, filename = "boxplotDP.png", width = 9, height = 4.5, bg = "white")
```

#### Orne Harbour
```{r}
boxplotBioDateOH <- dfBio %>%
  filter(Site == "Orne Harbour") %>% 
  mutate(Date = factor(Date)) %>% 
  ggplot(aes(x = Date, y = Biovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Biovolume (µm^3)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme2

boxplotBioDateOH

ggsave(boxplotBioDateOH, filename = "boxplotBioDateOH.png", width = 7, height = 6, bg = "white")
```

```{r}
boxplotOH <- (boxplotDensityDateOH | boxplotTotBioDateOH | boxplotBioDateOH) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

boxplotOH

ggsave(boxplotOH, filename = "boxplotOH.png", width = 9, height = 4.5, bg = "white")
```

## Cell Density & Indices {.tabset}

### Wildlife Scoring Index
```{r}
boxplotDensityWildIndex <- dfIndex %>%
  mutate(WildlifeScoringIndex = factor(WildlifeScoringIndex)) %>% 
  ggplot(aes(x = WildlifeScoringIndex, y = CellDensity, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Density (cells/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotDensityWildIndex

ggsave(boxplotDensityWildIndex, filename = "boxplotDensityWildIndex.png", width = 6, height = 4, bg = "white")
```

### Rocks Index
```{r}
boxplotDensityRockIndex <- dfIndex %>%
  mutate(RocksDustScoringIndex = factor(RocksDustScoringIndex)) %>% 
  ggplot(aes(x = RocksDustScoringIndex, y = CellDensity, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Density (cells/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotDensityRockIndex

ggsave(boxplotDensityRockIndex, filename = "boxplotDensityRockIndex.png", width = 6, height = 4, bg = "white")
```

### Oceanic Influence Index
```{r}
boxplotDensityOceanIndex <- dfIndex %>%
  mutate(OceanicInfluenceScoringIndex = factor(OceanicInfluenceScoringIndex)) %>% 
  ggplot(aes(x = OceanicInfluenceScoringIndex, y = CellDensity, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Density (cells/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotDensityOceanIndex

ggsave(boxplotDensityOceanIndex, filename = "boxplotDensityOceanIndex.png", width = 6, height = 4, bg = "white")
```

## Biovolume Concentration & Indices {.tabset}

### Wildlife Scoring Index
```{r}
boxplotTotBioWildIndex <- dfIndex %>%
  mutate(WildlifeScoringIndex = factor(WildlifeScoringIndex)) %>% 
  ggplot(aes(x = WildlifeScoringIndex, y = TotalBiovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Biovolume Concentration (µm^3/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotTotBioWildIndex

ggsave(boxplotTotBioWildIndex, filename = "boxplotTotBioWildIndex.png", width = 6, height = 4, bg = "white")
```

### Rocks Index
```{r}
boxplotTotBioRockIndex <- dfIndex %>%
  mutate(RocksDustScoringIndex = factor(RocksDustScoringIndex)) %>% 
  ggplot(aes(x = RocksDustScoringIndex, y = TotalBiovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Biovolume Concentration (µm^3/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotTotBioRockIndex

ggsave(boxplotTotBioRockIndex, filename = "boxplotTotBioRockIndex.png", width = 6, height = 4, bg = "white")
```

### Oceanic Influence Index
```{r}
boxplotTotBioOceanIndex <- dfIndex %>%
  mutate(OceanicInfluenceScoringIndex = factor(OceanicInfluenceScoringIndex)) %>% 
  ggplot(aes(x = OceanicInfluenceScoringIndex, y = TotalBiovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Biovolume Concentration (µm^3/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotTotBioOceanIndex

ggsave(boxplotTotBioOceanIndex, filename = "boxplotTotBioOceanIndex.png", width = 6, height = 4, bg = "white")
```


## Biovolume & Indices {.tabset}

### Wildlife Scoring Index
```{r}
boxplotBioWildIndex <- dfIndex %>%
  mutate(WildlifeScoringIndex = factor(WildlifeScoringIndex)) %>% 
  ggplot(aes(x = WildlifeScoringIndex, y = Biovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Biovolume (µm^3)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme

boxplotBioWildIndex

ggsave(boxplotBioWildIndex, filename = "boxplotBioWildIndex.png", width = width = 6, height = 4, bg = "white")
```

### Rocks Index
```{r}
boxplotBioRockIndex <- dfIndex %>%
  mutate(RocksDustScoringIndex = factor(RocksDustScoringIndex)) %>% 
  ggplot(aes(x = RocksDustScoringIndex, y = Biovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Biovolume (µm^3)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme

boxplotBioRockIndex

ggsave(boxplotBioRockIndex, filename = "boxplotBioRockIndex.png", width = width = 6, height = 4, bg = "white")
```

### Oceanic Influence Index
```{r}
boxplotBioOceanIndex <- dfIndex %>%
  mutate(OceanicInfluenceScoringIndex = factor(OceanicInfluenceScoringIndex)) %>% 
  ggplot(aes(x = OceanicInfluenceScoringIndex, y = Biovolume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.85, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Biovolume (µm^3)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme

boxplotBioOceanIndex

ggsave(boxplotBioOceanIndex, filename = "boxplotBioOceanIndex.png",  width = 6, height = 4, bg = "white")
```


## Combined Figures {.tabset}

### Wildlife Scoring Index
```{r}
boxplotWildIndex <- (
  (boxplotDensityWildIndex | boxplotTotBioWildIndex) / 
   boxplotBioWildIndex
) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

boxplotWildIndex

ggsave(boxplotWildIndex, filename = "boxplotWildIndex.png", width = 9, height = 8, bg = "white")
```

### Rocks Index
```{r}
boxplotRockIndex <- (
  (boxplotDensityRockIndex | boxplotTotBioRockIndex) / 
   boxplotBioRockIndex
) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

boxplotRockIndex

ggsave(boxplotRockIndex, filename = "boxplotRockIndex.png", width = 9, height = 8, bg = "white")
```

### Oceanic Influence Index
```{r}
boxplotOceanIndex <- (
  (boxplotDensityOceanIndex | boxplotTotBioOceanIndex) / 
   boxplotBioOceanIndex
) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

boxplotOceanIndex

ggsave(boxplotOceanIndex, filename = "boxplotOceanIndex.png", width = 9, height = 8, bg = "white")
```

# Algal Properties by Latitude {.tabset}
Below are scatterplots of the relationship between latitude and cell density, total biovolume concentration and cell biovolume. 

## Cell Density
```{r}
scatterLatDensity <- dfIndex %>% 
  ggplot(aes(x = Latitude, y = CellDensity, fill = Color, color = Color)) +
  geom_point(alpha = 0.8, 
             shape = 21, 
             size = 3) + 
  geom_smooth(method = "lm") +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Density (cells/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

scatterLatDensity
```

## Biovolume Concentration 
```{r}
scatterLatTotBio <- dfIndex %>% 
  ggplot(aes(x = Latitude, y = TotalBiovolume, fill = Color, color = Color)) +
  geom_point(alpha = 0.8, 
             shape = 21, 
             size = 3) + 
  geom_smooth(method = "lm") +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Biovolume Concentration (µm^3/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

scatterLatTotBio
```

## Biovolume 
```{r}
scatterLatBio <- dfIndex %>% 
  ggplot(aes(x = Latitude, y = Biovolume, fill = Color, color = Color)) +
  geom_point(alpha = 0.8, 
             shape = 21, 
             size = 3) + 
  geom_smooth(method = "lm") +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Biovolume (µm^3)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

scatterLatBio
```

# Anion Scatter Plots {.tabset}
In this section, I will assess the relationship between cell density and each anion, colored by algae presence and faceted by site, to see whether algae/anion relationships differ between locations or if there are any trends that jump out at me. I will do the same for biovolume concentration to see if it differs from cell density. 

## All Sites (Cell Density)

```{r}
dfCAnion_long <- dfBio %>% 
  pivot_longer(cols = c(Fo, Cl, SO42, PO43, NO2NO3),
               names_to = "Anion",
               values_to = "Concentration")

scatterMatrix <- dfCAnion_long %>% 
  ggplot(aes(x = Concentration, y = CellDensity)) +
  geom_point(aes(color = Color), alpha = 0.6) +  
  geom_smooth(method = "lm", color = "tomato", linewidth = 0.5, alpha = 0.5, se = FALSE) +  
  scale_color_manual(values = colorPal) +
  scale_y_sqrt() + 
  scale_x_sqrt() +
  facet_grid(Site ~ Anion, scales = "free_x", labeller = label_wrap_gen(width = 10)) + 
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "Cell Density (cells/mL)",
       color = NULL)

scatterMatrix + 
  theme(strip.text.y = element_text(angle = 0, size = 9),
        axis.text.y = element_text(size = 8))

ggsave(scatterMatrix, filename = "scatterMatrix.png", width = 8, height = 9, bg = "white")
```

## All Sites (Biovolume Concentration)

```{r}
ScatterMatrixTotBio <- dfCAnion_long %>% 
  ggplot(aes(x = Concentration, y = TotalBiovolume)) +
  geom_point(aes(color = Color),alpha = 0.6) + 
  geom_smooth(method = "lm", color = "tomato", linewidth = 0.5, alpha = 0.5, se = FALSE) +  
  scale_color_manual(values = colorPal) +
  scale_y_sqrt() + 
  scale_x_sqrt() +
  facet_grid(Site ~ Anion, scales = "free_x", labeller = label_wrap_gen(width = 10)) + 
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "Cell Biovolume Concentration (µm^3/mL)",
       color = NULL)

ScatterMatrixTotBio + 
  theme(strip.text.y = element_text(angle = 0, size = 9),
        axis.text.y = element_text(size = 8))

ggsave(ScatterMatrixTotBio, filename = "ScatterMatrixTotBio.png", width = 8, height = 9, bg = "white")
```

## All Sites (Biovolume)

```{r}
ScatterMatrixBio <- dfCAnion_long %>% 
  ggplot(aes(x = Concentration, y = Biovolume)) +
  geom_point(aes(color = Color), alpha = 0.6) + 
  geom_smooth(method = "lm", color = "tomato", linewidth = 0.5, alpha = 0.5, se = FALSE) +  
  scale_color_manual(values = colorPal) +
  scale_y_sqrt() + 
  scale_x_sqrt() +
  facet_grid(Site ~ Anion, scales = "free_x", labeller = label_wrap_gen(width = 10)) + 
  theme_minimal() + 
  plot_theme2 + 
  theme(axis.text = element_text(size = 10)) + 
  labs(x = "Anion Concentration (ppm)",
       y = "Cell Biovolume (µm^3)",
       color = NULL)

ScatterMatrixBio + 
  theme(strip.text.y = element_text(angle = 0, size = 9),
        axis.text.y = element_text(size = 8))

ggsave(ScatterMatrixBio, filename = "ScatterMatrixBio.png", width = 8, height = 9, bg = "white")
```

# Heat Maps {.tabset}

## Cell Density vs. Total Biovolume & Anions (by Site)
```{r}
# Variable groups
algae_vars <- c("CellDensity", "TotalBiovolume")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filtered data
df_filtered <- dfBio %>%
  filter(Site != "Pendulum Cove, Deception Island") %>% 
  filter(Site != "Danco Island") %>% 
  select(Site, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Function to extract correlations
get_correlation_data <- function(data, location_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")
  
  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  ## Algae vs Anion correlations
  pairs <- expand.grid(Algae = algae_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Algae, Anion],
      p_value = p_vals[Algae, Anion],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      Site = location_name
    ) %>%
    ungroup()

  ## CellDensity vs TotalBiovolume correlation (the 6th column)
  bio_pair <- tibble(
    Algae = algae_vars,
    Anion = "TotalBiovolume",
    Spearman_r = r_vals["CellDensity", "TotalBiovolume"],
    p_value = p_vals["CellDensity", "TotalBiovolume"],
    significant = p_value < 0.05,
    label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
    Site = location_name
  )

  ## Combine both
  combined_results <- bind_rows(results, bio_pair)

  return(combined_results)
}


# Apply function across all sites
cor_results <- df_filtered %>%
  group_split(Site) %>%
  map_dfr(~get_correlation_data(.x, unique(.x$Site)))


# Create combined y-axis label
cor_results <- cor_results %>%
  mutate(Location_Algae = paste(Site, Algae, sep = " - "))

# Order the rows by site (Location), then Algae variable
location_order <- unique(cor_results$Site)

cor_results <- cor_results %>%
  mutate(Location_Algae = factor(
    Location_Algae,
    levels = unlist(lapply(location_order, function(loc) {
      c(paste(loc, "CellDensity", sep = " - "),
        paste(loc, "TotalBiovolume", sep = " - "))
    }))
  ))

# Order Anion axis (x-axis), adding the 6th column last
cor_results$Anion <- factor(cor_results$Anion, levels = c(anion_vars, "TotalBiovolume"))

# Plot
spearmanPlotBio <- ggplot(cor_results, aes(x = Anion, y = Location_Algae, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), 
            color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red", midpoint = 0, 
    name = "Spearman r"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    axis.text.y = element_text(size = 9)
  )

# Save plot
spearmanPlotBio
ggsave(spearmanPlotBio, filename = "spearmanPlotBio.png", width = 8, height = 9, bg = "white")
```

## Cell Density vs. Total Biovolume & Anions (by SiteLoc)
```{r}
# Variable groups
algae_vars <- c("CellDensity", "TotalBiovolume")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filtered data by SiteLoc 
df_filtered <- dfIndex %>%
  filter(SiteLoc != "North") %>% 
  select(SiteLoc, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Function to extract correlations
get_correlation_data <- function(data, location_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")
  
  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Algae vs Anion correlations
  pairs <- expand.grid(Algae = algae_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Algae, Anion],
      p_value = p_vals[Algae, Anion],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      SiteLoc = location_name
    ) %>%
    ungroup()

  # CellDensity vs TotalBiovolume correlation
  bio_pair <- tibble(
    Algae = algae_vars,
    Anion = "TotalBiovolume",
    Spearman_r = r_vals["CellDensity", "TotalBiovolume"],
    p_value = p_vals["CellDensity", "TotalBiovolume"],
    significant = p_value < 0.05,
    label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
    SiteLoc = location_name
  )

  bind_rows(results, bio_pair)
}

# Apply function across SiteLoc groups
cor_results <- df_filtered %>%
  group_split(SiteLoc) %>%
  map_dfr(~get_correlation_data(.x, unique(.x$SiteLoc)))

# Create combined y-axis label
cor_results <- cor_results %>%
  mutate(Location_Algae = paste(SiteLoc, Algae, sep = " - "))

# Order the rows by SiteLoc, then Algae
location_order <- unique(cor_results$SiteLoc)

cor_results <- cor_results %>%
  mutate(Location_Algae = factor(
    Location_Algae,
    levels = unlist(lapply(location_order, function(loc) {
      c(paste(loc, "CellDensity", sep = " - "),
        paste(loc, "TotalBiovolume", sep = " - "))
    }))
  ))

# Order Anion axis (x-axis), adding TotalBiovolume last
cor_results$Anion <- factor(cor_results$Anion, levels = c(anion_vars))

# Plot
spearmanPlotBioSiteLoc <- ggplot(filter(cor_results, Anion %in% anion_vars), 
                          aes(x = Anion, y = Location_Algae, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, Anion %in% anion_vars & significant), 
            color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red", midpoint = 0, 
    name = "Spearman r"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    axis.text.y = element_text(size = 9)
  )

spearmanPlotBioSiteLoc

ggsave(spearmanPlotBioSiteLoc, filename = "spearmanPlotBioSiteLoc.png", width = 8, height = 7, bg = "white")
```

## Cell Density vs. Total Biovolume & Anions (by Index) {.tabset}

### Wildlife Index
```{r}
# Variable groups
algae_vars <- c("CellDensity", "TotalBiovolume")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filtered data by WildlifeScoringIndex 
df_filtered <- dfIndex %>%
  select(WildlifeScoringIndex, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Function to extract correlations
get_correlation_data <- function(data, location_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")
  
  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Algae vs Anion correlations
  pairs <- expand.grid(Algae = algae_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Algae, Anion],
      p_value = p_vals[Algae, Anion],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      WildlifeScoringIndex = location_name
    ) %>%
    ungroup()

  # CellDensity vs TotalBiovolume correlation
  bio_pair <- tibble(
    Algae = algae_vars,
    Anion = "TotalBiovolume",
    Spearman_r = r_vals["CellDensity", "TotalBiovolume"],
    p_value = p_vals["CellDensity", "TotalBiovolume"],
    significant = p_value < 0.05,
    label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
    WildlifeScoringIndex = location_name
  )

  bind_rows(results, bio_pair)
}

# Apply function across WildlifeScoringIndex groups
cor_results <- df_filtered %>%
  group_split(WildlifeScoringIndex) %>%
  map_dfr(~get_correlation_data(.x, unique(.x$WildlifeScoringIndex)))

# Create combined y-axis label
cor_results <- cor_results %>%
  mutate(Location_Algae = paste(WildlifeScoringIndex, Algae, sep = " - "))

# Order the rows by WildlifeScoringIndex, then Algae
location_order <- unique(cor_results$WildlifeScoringIndex)

cor_results <- cor_results %>%
  mutate(Location_Algae = factor(
    Location_Algae,
    levels = unlist(lapply(location_order, function(loc) {
      c(paste(loc, "CellDensity", sep = " - "),
        paste(loc, "TotalBiovolume", sep = " - "))
    }))
  ))

# Order Anion axis (x-axis), adding TotalBiovolume last
cor_results$Anion <- factor(cor_results$Anion, levels = c(anion_vars))

# Plot
spearmanPlotBioWildlifeScoringIndex <- ggplot(filter(cor_results, Anion %in% anion_vars), 
                          aes(x = Anion, y = Location_Algae, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, Anion %in% anion_vars & significant), 
            color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red", midpoint = 0, 
    name = "Spearman r"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    axis.text.y = element_text(size = 9)
  )

spearmanPlotBioWildlifeScoringIndex

ggsave(spearmanPlotBioWildlifeScoringIndex, filename = "spearmanPlotBioWildlifeScoringIndex.png", width = 8, height = 7, bg = "white")
```

### Rocks Index
```{r}
# Variable groups
algae_vars <- c("CellDensity", "TotalBiovolume")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filtered data by RocksDustScoringIndex 
df_filtered <- dfIndex %>%
  select(RocksDustScoringIndex, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Function to extract correlations
get_correlation_data <- function(data, location_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")
  
  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Algae vs Anion correlations
  pairs <- expand.grid(Algae = algae_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Algae, Anion],
      p_value = p_vals[Algae, Anion],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      RocksDustScoringIndex = location_name
    ) %>%
    ungroup()

  # CellDensity vs TotalBiovolume correlation
  bio_pair <- tibble(
    Algae = algae_vars,
    Anion = "TotalBiovolume",
    Spearman_r = r_vals["CellDensity", "TotalBiovolume"],
    p_value = p_vals["CellDensity", "TotalBiovolume"],
    significant = p_value < 0.05,
    label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
    RocksDustScoringIndex = location_name
  )

  bind_rows(results, bio_pair)
}

# Apply function across RocksDustScoringIndex groups
cor_results <- df_filtered %>%
  group_split(RocksDustScoringIndex) %>%
  map_dfr(~get_correlation_data(.x, unique(.x$RocksDustScoringIndex)))

# Create combined y-axis label
cor_results <- cor_results %>%
  mutate(Location_Algae = paste(RocksDustScoringIndex, Algae, sep = " - "))

# Order the rows by RocksDustScoringIndex, then Algae
location_order <- unique(cor_results$RocksDustScoringIndex)

cor_results <- cor_results %>%
  mutate(Location_Algae = factor(
    Location_Algae,
    levels = unlist(lapply(location_order, function(loc) {
      c(paste(loc, "CellDensity", sep = " - "),
        paste(loc, "TotalBiovolume", sep = " - "))
    }))
  ))

# Order Anion axis (x-axis), adding TotalBiovolume last
cor_results$Anion <- factor(cor_results$Anion, levels = c(anion_vars))

# Plot
spearmanPlotBioRocksDustScoringIndex <- ggplot(filter(cor_results, Anion %in% anion_vars), 
                          aes(x = Anion, y = Location_Algae, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, Anion %in% anion_vars & significant), 
            color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red", midpoint = 0, 
    name = "Spearman r"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    axis.text.y = element_text(size = 9)
  )

spearmanPlotBioRocksDustScoringIndex

ggsave(spearmanPlotBioRocksDustScoringIndex, filename = "spearmanPlotBioRocksDustScoringIndex.png", width = 8, height = 7, bg = "white")
```

### Ocean Index

```{r}
# Variable groups
algae_vars <- c("CellDensity", "TotalBiovolume")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filtered data by OceanicInfluenceScoringIndex 
df_filtered <- dfIndex %>%
  select(OceanicInfluenceScoringIndex, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Function to extract correlations
get_correlation_data <- function(data, location_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")
  
  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Algae vs Anion correlations
  pairs <- expand.grid(Algae = algae_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Algae, Anion],
      p_value = p_vals[Algae, Anion],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      OceanicInfluenceScoringIndex = location_name
    ) %>%
    ungroup()

  # CellDensity vs TotalBiovolume correlation
  bio_pair <- tibble(
    Algae = algae_vars,
    Anion = "TotalBiovolume",
    Spearman_r = r_vals["CellDensity", "TotalBiovolume"],
    p_value = p_vals["CellDensity", "TotalBiovolume"],
    significant = p_value < 0.05,
    label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
    OceanicInfluenceScoringIndex = location_name
  )

  bind_rows(results, bio_pair)
}

# Apply function across OceanicInfluenceScoringIndex groups
cor_results <- df_filtered %>%
  group_split(OceanicInfluenceScoringIndex) %>%
  map_dfr(~get_correlation_data(.x, unique(.x$OceanicInfluenceScoringIndex)))

# Create combined y-axis label
cor_results <- cor_results %>%
  mutate(Location_Algae = paste(OceanicInfluenceScoringIndex, Algae, sep = " - "))

# Order the rows by OceanicInfluenceScoringIndex, then Algae
location_order <- unique(cor_results$OceanicInfluenceScoringIndex)

cor_results <- cor_results %>%
  mutate(Location_Algae = factor(
    Location_Algae,
    levels = unlist(lapply(location_order, function(loc) {
      c(paste(loc, "CellDensity", sep = " - "),
        paste(loc, "TotalBiovolume", sep = " - "))
    }))
  ))

# Order Anion axis (x-axis), adding TotalBiovolume last
cor_results$Anion <- factor(cor_results$Anion, levels = c(anion_vars))

# Plot
spearmanPlotBioOceanicInfluenceScoringIndex <- ggplot(filter(cor_results, Anion %in% anion_vars), 
                          aes(x = Anion, y = Location_Algae, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, Anion %in% anion_vars & significant), 
            color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red", midpoint = 0, 
    name = "Spearman r"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    axis.text.y = element_text(size = 9)
  )

spearmanPlotBioOceanicInfluenceScoringIndex

ggsave(spearmanPlotBioOceanicInfluenceScoringIndex, filename = "spearmanPlotBioOceanicInfluenceScoringIndex.png", width = 8, height = 7, bg = "white")
```

## Cell Density & Anions (by Color)

```{r}
# Variable groups
algae_vars <- c("CellDensity")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filtered data
df_filtered <- dfBio %>%
  filter(Location == "Antarctic Peninsula") %>% 
  filter(Color %in% c("green", "red")) %>% 
  select(Location, Color, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Function to extract correlations
get_correlation_data <- function(data, location_name, algae_name) {
  cor_matrix <- rcorr(as.matrix(data[, c(algae_vars, anion_vars)]), type = "spearman")
  
  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Extract only CellDensity vs anion correlations
  pairs <- expand.grid(AlgaeVar = algae_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[AlgaeVar, Anion],
      p_value = p_vals[AlgaeVar, Anion],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      Location = location_name,
      SnowColor = algae_name
    ) %>%
    ungroup()

  return(results)
}

# Apply across all Location x SnowColor groups
cor_results <- df_filtered %>%
  group_split(Location, Color) %>%
  map_dfr(~get_correlation_data(.x, unique(.x$Location), unique(.x$Color)))

# Convert Anion to factor for plot order
cor_results$Anion <- factor(cor_results$Anion, levels = anion_vars)

# Plot with SnowColor on y-axis nested within Location
spearmanPlotAPRG <- ggplot(cor_results, aes(x = Anion, y = interaction(Location, SnowColor, sep = " - "), fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank(),
        axis.text.y = element_text(size = 8))

spearmanPlotAPRG

# Save
ggsave(spearmanPlotAPRG, filename = "spearmanPlotAPRG.png", width = 7, height = 6, bg = "white")
```

## Cell Density & Anions (Damoy Point by Color)

```{r}
# Variable groups
algae_vars <- c("CellDensity")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filtered data
df_filtered <- dfBio %>%
  filter(Site == "Damoy Point") %>% 
  filter(Color %in% c("green", "red")) %>% 
  select(Location, Color, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Function to extract correlations
get_correlation_data <- function(data, location_name, algae_name) {
  cor_matrix <- rcorr(as.matrix(data[, c(algae_vars, anion_vars)]), type = "spearman")
  
  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Extract only CellDensity vs anion correlations
  pairs <- expand.grid(AlgaeVar = algae_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[AlgaeVar, Anion],
      p_value = p_vals[AlgaeVar, Anion],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      Location = location_name,
      SnowColor = algae_name
    ) %>%
    ungroup()

  return(results)
}

# Apply across all Location x SnowColor groups
cor_results <- df_filtered %>%
  group_split(Location, Color) %>%
  map_dfr(~get_correlation_data(.x, unique(.x$Location), unique(.x$Color)))

# Convert Anion to factor for plot order
cor_results$Anion <- factor(cor_results$Anion, levels = anion_vars)

# Plot with SnowColor on y-axis nested within Location
spearmanPlotAPRG_DP <- ggplot(cor_results, aes(x = Anion, y = interaction(Location, SnowColor, sep = " - "), fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank(),
        axis.text.y = element_text(size = 8))

spearmanPlotAPRG_DP

# Save
ggsave(spearmanPlotAPRG_DP, filename = "spearmanPlotAPRG_DP.png", width = 7, height = 7, bg = "white")
```

## Biovolume Concentration & Anions 
```{r}
# Variable groups
algae_vars <- c("TotalBiovolume")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filtered data
df_filtered <- dfBio %>%
  filter(Location == "Antarctic Peninsula") %>% 
  filter(Color %in% c("green", "red")) %>% 
  select(Location, Color, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Function to extract correlations
get_correlation_data <- function(data, location_name, algae_name) {
  cor_matrix <- rcorr(as.matrix(data[, c(algae_vars, anion_vars)]), type = "spearman")
  
  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Extract only CellDensity vs anion correlations
  pairs <- expand.grid(AlgaeVar = algae_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[AlgaeVar, Anion],
      p_value = p_vals[AlgaeVar, Anion],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      Location = location_name,
      SnowColor = algae_name
    ) %>%
    ungroup()

  return(results)
}

# Apply across all Location x SnowColor groups
cor_results <- df_filtered %>%
  group_split(Location, Color) %>%
  map_dfr(~get_correlation_data(.x, unique(.x$Location), unique(.x$Color)))

# Convert Anion to factor for plot order
cor_results$Anion <- factor(cor_results$Anion, levels = anion_vars)

# Plot with SnowColor on y-axis nested within Location
spearmanPlotAPRGTotBio <- ggplot(cor_results, aes(x = Anion, y = interaction(Location, SnowColor, sep = " - "), fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank(),
        axis.text.y = element_text(size = 8))

spearmanPlotAPRGTotBio

# Save
ggsave(spearmanPlotAPRGTotBio, filename = "spearmanPlotAPRGTotBio.png", width = 7, height = 6, bg = "white")
```

## Biovolume & Anions 
```{r}
# Variable groups
algae_vars <- c("Biovolume")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filtered data
df_filtered <- dfBio %>%
  filter(Site != "Pendulum Cove, Deception Island") %>% 
  filter(Site != "Danco Island") %>% 
  select(Site, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Function to extract correlations
get_correlation_data <- function(data, location_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")
  
  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Extract only CellDensity vs anion correlations
  pairs <- expand.grid(Algae = algae_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Algae, Anion],
      p_value = p_vals[Algae, Anion],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      Site = location_name
    ) %>%
    ungroup()

  return(results)
}

# Apply across all sites
cor_results <- df_filtered %>%
  group_split(Site) %>%
  map_dfr(~get_correlation_data(.x, unique(.x$Site)))

# Convert Anion to factor with desired order
cor_results$Anion <- factor(cor_results$Anion, levels = anion_vars)

# Plot
spearmanPlotL <- ggplot(cor_results, aes(x = Anion, y = Site, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank())
  
spearmanPlotL

ggsave(spearmanPlotL, filename = "spearmanPlotL.png", width = 7, height = 7, bg = "white")
```

## Biovolume & Anions (by Latitude)
```{r}
# Variable groups
algae_vars <- c("Biovolume")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filtered data
df_filtered <- dfIndex %>%
  filter(SiteLoc != "North") %>% 
  select(SiteLoc, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Function to extract correlations
get_correlation_data <- function(data, location_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")
  
  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Extract only Biovolume vs anion correlations
  pairs <- expand.grid(Algae = algae_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Algae, Anion],
      p_value = p_vals[Algae, Anion],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      SiteLoc = location_name
    ) %>%
    ungroup()

  return(results)
}

# Apply across all SiteLoc groups
cor_results <- df_filtered %>%
  group_split(SiteLoc) %>%
  map_dfr(~get_correlation_data(.x, unique(.x$SiteLoc)))

# Convert Anion to factor with desired order
cor_results$Anion <- factor(cor_results$Anion, levels = anion_vars)

# Plot
spearmanPlotLSiteLoc <- ggplot(cor_results, aes(x = Anion, y = SiteLoc, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank())

spearmanPlotLSiteLoc

ggsave(spearmanPlotLSiteLoc, filename = "spearmanPlotLSiteLoc.png", width = 7, height = 7, bg = "white")
```

## Biovolume & Anions (by Index) {.tabset}

### Wildlife Index
```{r}
# Variable groups
algae_vars <- c("Biovolume")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filtered data
df_filtered <- dfIndex %>%
  select(WildlifeScoringIndex, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Function to extract correlations
get_correlation_data <- function(data, location_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")
  
  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Extract only Biovolume vs anion correlations
  pairs <- expand.grid(Algae = algae_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Algae, Anion],
      p_value = p_vals[Algae, Anion],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      WildlifeScoringIndex = location_name
    ) %>%
    ungroup()

  return(results)
}

# Apply across all WildlifeScoringIndex groups
cor_results <- df_filtered %>%
  group_split(WildlifeScoringIndex) %>%
  map_dfr(~get_correlation_data(.x, unique(.x$WildlifeScoringIndex)))

# Convert Anion to factor with desired order
cor_results$Anion <- factor(cor_results$Anion, levels = anion_vars)

# Plot
spearmanPlotLWildlifeScoringIndex <- ggplot(cor_results, aes(x = Anion, y = WildlifeScoringIndex, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank())

spearmanPlotLWildlifeScoringIndex

ggsave(spearmanPlotLWildlifeScoringIndex, filename = "spearmanPlotLWildlifeScoringIndex.png", width = 7, height = 7, bg = "white")
```

### Rocks Index
```{r}
# Variable groups
algae_vars <- c("Biovolume")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filtered data
df_filtered <- dfIndex %>%
  select(RocksDustScoringIndex, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Function to extract correlations
get_correlation_data <- function(data, location_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")
  
  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Extract only Biovolume vs anion correlations
  pairs <- expand.grid(Algae = algae_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Algae, Anion],
      p_value = p_vals[Algae, Anion],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      RocksDustScoringIndex = location_name
    ) %>%
    ungroup()

  return(results)
}

# Apply across all RocksDustScoringIndex groups
cor_results <- df_filtered %>%
  group_split(RocksDustScoringIndex) %>%
  map_dfr(~get_correlation_data(.x, unique(.x$RocksDustScoringIndex)))

# Convert Anion to factor with desired order
cor_results$Anion <- factor(cor_results$Anion, levels = anion_vars)

# Plot
spearmanPlotLRocksDustScoringIndex <- ggplot(cor_results, aes(x = Anion, y = RocksDustScoringIndex, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank())

spearmanPlotLRocksDustScoringIndex

ggsave(spearmanPlotLRocksDustScoringIndex, filename = "spearmanPlotLRocksDustScoringIndex.png", width = 7, height = 7, bg = "white")
```

### Oceanic Influence Index
```{r}
# Variable groups
algae_vars <- c("Biovolume")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filtered data
df_filtered <- dfIndex %>%
  select(OceanicInfluenceScoringIndex, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Function to extract correlations
get_correlation_data <- function(data, location_name) {
  cor_matrix <- rcorr(as.matrix(data[, -1]), type = "spearman")
  
  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Extract only Biovolume vs anion correlations
  pairs <- expand.grid(Algae = algae_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[Algae, Anion],
      p_value = p_vals[Algae, Anion],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      OceanicInfluenceScoringIndex = location_name
    ) %>%
    ungroup()

  return(results)
}

# Apply across all OceanicInfluenceScoringIndex groups
cor_results <- df_filtered %>%
  group_split(OceanicInfluenceScoringIndex) %>%
  map_dfr(~get_correlation_data(.x, unique(.x$OceanicInfluenceScoringIndex)))

# Convert Anion to factor with desired order
cor_results$Anion <- factor(cor_results$Anion, levels = anion_vars)

# Plot
spearmanPlotLOceanicInfluenceScoringIndex <- ggplot(cor_results, aes(x = Anion, y = OceanicInfluenceScoringIndex, fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank())

spearmanPlotLOceanicInfluenceScoringIndex

ggsave(spearmanPlotLOceanicInfluenceScoringIndex, filename = "spearmanPlotLOceanicInfluenceScoringIndex.png", width = 7, height = 7, bg = "white")
```

## Biovolume & Anions (by Color)

```{r}
# Variable groups
algae_vars <- c("Biovolume")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Filtered data
df_filtered <- dfBio %>%
  filter(Location == "Antarctic Peninsula") %>% 
  filter(Color %in% c("green", "red")) %>% 
  select(Location, Color, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Function to extract correlations
get_correlation_data <- function(data, location_name, algae_name) {
  cor_matrix <- rcorr(as.matrix(data[, c(algae_vars, anion_vars)]), type = "spearman")
  
  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  # Extract only CellDensity vs anion correlations
  pairs <- expand.grid(AlgaeVar = algae_vars, Anion = anion_vars, stringsAsFactors = FALSE)

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[AlgaeVar, Anion],
      p_value = p_vals[AlgaeVar, Anion],
      significant = p_value < 0.05,
      label = paste0(round(Spearman_r, 2), "\n(p=", round(p_value, 3), ")"),
      Location = location_name,
      SnowColor = algae_name
    ) %>%
    ungroup()

  return(results)
}

# Apply across all Location x SnowColor groups
cor_results <- df_filtered %>%
  group_split(Location, Color) %>%
  map_dfr(~get_correlation_data(.x, unique(.x$Location), unique(.x$Color)))

# Convert Anion to factor for plot order
cor_results$Anion <- factor(cor_results$Anion, levels = anion_vars)

# Plot with SnowColor on y-axis nested within Location
spearmanPlotAPRGL <- ggplot(cor_results, aes(x = Anion, y = interaction(Location, SnowColor, sep = " - "), fill = Spearman_r)) +
  geom_tile() +
  geom_tile(data = filter(cor_results, significant), color = "black", linewidth = 0.75) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, name = "Spearman r") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank(),
        axis.text.y = element_text(size = 8))

spearmanPlotAPRGL

# Save
ggsave(spearmanPlotAPRGL, filename = "spearmanPlotAPRGL.png", width = 7, height = 6, bg = "white")
```