---
title: "Data Visualization - Cell Density & Volume"
author: "Ellie Ryan"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: flatly
    highlight: zenburn
    toc: true
    toc_float: true
    toc_collapsed: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Created:** 9/16/2025

**Last Modified:** 1/23/2026

```{r, message=FALSE, warning=FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(stringr)
library(kableExtra)
library(scales)
library(GGally)
library(patchwork)
library(Hmisc)
library(reshape2)
library(car)
library(purrr)
library(tibble)
library(FSA) 
```

```{r}
colorPal <- c(
  "clean" = "#769a9c",
  "red" = "#933410",
  "green" = "#326e35"
)

algaePal<- c(
  "clean" = "#a5bdc0",
  "algae" = "#735a69"
)

plot_theme <- theme(
  axis.title = element_text(size = 12, face = "bold"), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  plot.margin = margin(20, 20, 20, 20)
)

plot_themePub <- theme(
  text = element_text(family = "Helvetica"),
  axis.title.x = element_text(size = 10, margin = margin(t = 10)),  # t = top margin
  axis.title.y = element_text(size = 10, margin = margin(r = 10)),  # r = right margin
  axis.text = element_text(size = 10),
  axis.line.x = element_line(color = "black", linewidth = 0.5),
  axis.line.y = element_line(color = "black", linewidth = 0.5),,
  axis.ticks.x = element_line(color = "black", linewidth = 0.25),
  axis.ticks.y = element_line(color = "black", linewidth = 0.25),  # Add tick marks
  axis.ticks.length = unit(-0.15, "cm"),  # Length of tick marks
  panel.grid = element_blank(),
  panel.background = element_blank(),
  panel.border = element_blank(),
  legend.position = "top",
  legend.justification = c(0, 1),
  legend.text = element_text(size = 10),
  legend.title = element_text(size = 10),
  plot.title = element_text(size = 14, face = "bold"),
  plot.subtitle = element_text(size = 13, face = "italic"),
  plot.margin = margin(10, 10, 10, 10)
)

plot_theme2 <- theme(
  axis.title = element_text(size = 12, face = "bold"), # Axis labels: 12 pt, bold
  axis.text = element_text(size = 12), # Axis text: 12 pt
  axis.line = element_line(linewidth = 0.5),  # Thicker axis line
  legend.position = "top", # Position legend at the top
  legend.justification = c(0, 1), # Justify legend to top-left
  legend.text = element_text(size = 12), # Legend text: 12 pt
  legend.title = element_text(size = 12), 
  plot.title = element_text(size = 14, face = "bold"),  # Adjust plot title size
  plot.subtitle = element_text(size = 13, face = "italic"), 
  axis.text.x = element_text(angle = 45, hjust = 1), 
  plot.margin = margin(20, 20, 20, 20)
)

pairsplot_theme <- theme(
  text = element_text(size = 10),  # Adjust variable name text size
  axis.text = element_text(size = 8),  # Adjust axis text size
  plot.title = element_text(size = 12, face = "bold"),  # Adjust plot title size
  legend.text = element_text(size = 12))  # Adjust legend text size
```

# Overview

This is the updated document for cell density/volume analysis, following addition of the Petermann Island samples. This doc focuses specifically on cell density and cell volume, and does not consider cell biovolume or cell length for simplicity. The purpose of this `.Rmd` is to visualize cell density and volume by site and color, and obtain basic summary statistics associated with cell density in order to inform my next steps in analysis. Before proceeding, I will clean the dataset so that it features only the variables I am interested in for this specific analysis. In this dataset, 108 samples are used from the Antarctic Peninsula, following the removal of 6 outliers. These observations are broken down by `Color`, or the apparent snow color at time of sample collection. To keep things simple, there are only three levels to the `Color` category: red, green and clean. The breakdown of samples collected in the WAP is as follows: 

-   Antarctic Peninsula:
    -   108 samples
        -   43 red
        -   47 green
        -   18 clean
        
*NOTE: Fewer samples were used for analysis, since there are NA values for cell density/cell volumes for some samples.*
        
From outlier analysis conducted using figures generated in the first run-through of this script, the following outliers were identified for removal:

-   2025-02-19-SI-algae3 (green)
-   2025-02-24-DA-algae2 (red)
-   2025-02-24-DA-algae6 (green)
-   2025-03-07-NH-algae3 (clean)
-   2025-01-12-PI-algae10 (red)

See the Thesis Figures (Preliminary) doc for justification for removal of each.

## Data Wrangling

```{r}
df <- read.csv("CompiledDataPI.csv")
str(df)
```

```{r}
# Format date column 
df <- df %>%
  mutate(
    Date = as.Date(Date, format = "%m/%d/%y"))
```

```{r}
str(df)
```


```{r}
outliers <- c(
  "2025-02-19-SI-algae3",
  "2025-02-24-DA-algae2",
  "2025-02-24-DA-algae6",
  "2025-03-07-NH-algae3",
  "2025-01-12-PI-algae10")

df <- df %>% 
  filter(!SampleID %in% outliers)
```

### Cell Density Dataframe
I will keep the following variables in the `dfC` for this analysis:

-   `SampleID`
-   `Date`
-   `Location`
-   `Site`
-   `Color`
-   `CellDensity`

```{r}
dfC <- df %>% 
  select(c("SampleID", "Date", "Location", "Site", "Color", "CellDensity" )) %>% 
  na.omit(dfC) %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  relocate(Algae, .after = Color) %>% 
  mutate(Region = case_when(
    Site %in% c("Horseshoe Island", "Stonington Island") ~ "South", 
    TRUE ~ "North")) %>% 
  relocate(Region, .after = Site)

head(dfC)
```

### Cell Volume Dataframe
I will make a separate dataframe that calculates cell volume from the `CellLength` variable, since this variable has some NA values for samples with low cell density. There are 264 observations in the `dfC` dataframe with only `CellDensity` and 254 observations in the `dfCL` dataframe with both `CellDensity` and `CellLength` variables. I will remove rows where cell length was set to zero for non-zero counts (lengths were not measured for some low density counts). I am also going to add a new categorical variable, `Algae`, which generalizes the `Color` category by grouping red and green snow samples together as `algae` and clean samples as `clean`. 

Lastly, I will add a `CellVolume` variable. It is calculated as described below: 

**Biovolume:** This represents the volume of each cell, assuming a spherical shape for snow algae (µm^3^ or mm^3^). This method was first categorized by Hillebrand et al., 1999, *Biovolume Calculation for Pelagic and Benthic Microalgae*. The equation is as follows:

-   V~bio~ = 4/3πr^3^

Volume is in units of µm^3^.

```{r}
#Filter dataset and add algae variable
dfCL <- df %>% 
  select(c("SampleID", "Date", "Location", "Site", "Color", "CellDensity", "CellLength" )) %>% 
  na.omit(dfCL) %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  ))  %>% 
  relocate(Algae, .after = Color) %>% 
  mutate(Region = case_when(
    Site %in% c("Horseshoe Island", "Stonington Island") ~ "South", 
    TRUE ~ "North")) %>% 
  relocate(Region, .after = Site)

# Add volume variable
dfCL <- dfCL  %>% 
  mutate(
    # Calulate Volume (µm^3)
    Volume = (pi/6)*(CellLength^3)) %>% 
  relocate(Volume, .after = CellLength)

# Remove samples with zero values for cell length and nonzero values for cell density
dfCL <- dfCL %>% 
  filter(!(CellDensity > 0 & CellLength == 0))
  
```

### Cell Density & Anions Dataframe

I'll also make a dataframe with cell density and anion concentrations so I can assess this relationship using a pairs plot later on. I am going to combine NO2 and NO3 by summing them, to create a new variable, NO2NO3.

```{r}
dfCAnion <- df %>% 
  select(c("SampleID", "Date", "Location", "Site", "Color", CellDensity, Fo:SO42)) %>% 
  na.omit() %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ "clean"
  )) %>% 
  mutate(NO2NO3 = NO2 + NO3) %>% 
  relocate(Algae, .after = Color) %>% 
  select(-NO2, -NO3) %>% 
  mutate(Region = case_when(
    Site %in% c("Horseshoe Island", "Stonington Island") ~ "South", 
    TRUE ~ "North")) %>% 
  relocate(Region, .after = Site)
```

```{r}
dfCAnion_long <- dfCAnion %>% 
  pivot_longer(cols = Fo:NO2NO3,
               names_to = "Anion", 
               values_to = "Concentration")
```

### Cell Volume & Anions Dataframe
And I'll also make a dataframe with cell volume and anion concentrations.

```{r}

#Filter dataset and add algae variable
dfCLAnion <- df %>% 
  select(c("SampleID", "Date", "Location", "Site", "Color", "CellDensity", "CellLength",Fo:SO42)) %>% 
  na.omit(dfCLAnion) %>% 
  mutate(Algae = case_when(
    Color %in% c("red", "green") ~ "algae",
    Color == "clean" ~ 'clean'
  )) %>% 
  mutate(Region = case_when(
    Site %in% c("Horseshoe Island", "Stonington Island") ~ "South", 
    TRUE ~ "North")) %>% 
  relocate(Region, .after = Site)

# Add volume variable
dfCLAnion <- dfCLAnion  %>% 
  mutate(
    # Calulate Volume (µm^3)
    Volume = (pi/6)*(CellLength^3)) %>% 
  relocate(Volume, .after = CellLength)

# Remove samples with zero values for cell length and nonzero values for cell density
dfCLAnion <- dfCLAnion %>% 
  filter(!(CellDensity > 0 & CellLength == 0))

#Combine NO2 & NO3
dfCLAnion <- dfCLAnion %>% 
  mutate(NO2NO3 = NO2 + NO3) %>% 
  relocate(Algae, .after = Color) %>% 
  select(-NO2, -NO3)
```

```{r}
dfCLAnion_long <- dfCLAnion %>% 
  pivot_longer(cols = Fo:NO2NO3,
               names_to = "Anion", 
               values_to = "Concentration")
```

# Summary Statistics {.tabset}

## Cell Density (Site & Color)

Below is a summary stats table for cell density, grouped by `Color`. The `all` level of the `Color` classification includes red, green and clean snow observations.

```{r}
# Cell Density summary
dfCSummary <- dfC %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    mean = comma(round(mean(CellDensity, na.rm = TRUE), 2)),
    median = comma(round(median(CellDensity, na.rm = TRUE), 2)),
    max = comma(round(max(CellDensity, na.rm = TRUE), 2)),
    min = comma(round(min(CellDensity, na.rm = TRUE), 2)),
    sd = comma(round(sd(CellDensity, na.rm = TRUE), 2)),
    .groups = "drop"
  )

# Add "all" for each site
dfCSummaryAll <- dfC %>% 
  summarise(
    Color = "all",
    n = n(),
    mean = comma(round(mean(CellDensity, na.rm = TRUE), 2)),
    median = comma(round(median(CellDensity, na.rm = TRUE), 2)),
    max = comma(round(max(CellDensity, na.rm = TRUE), 2)),
    min = comma(round(min(CellDensity, na.rm = TRUE), 2)),
    sd = comma(round(sd(CellDensity, na.rm = TRUE), 2)),
    .groups = "drop"
  )

# Filter Antarctic red + green
dfCSummaryAPRG <- dfC %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Color = "red+green",
    n = n(),
    mean = comma(round(mean(CellDensity, na.rm = TRUE), 2)),
    median = comma(round(median(CellDensity, na.rm = TRUE), 2)),
    max = comma(round(max(CellDensity, na.rm = TRUE), 2)),
    min = comma(round(min(CellDensity, na.rm = TRUE), 2)),
    sd = comma(round(sd(CellDensity, na.rm = TRUE), 2))
  )

# Combine and order rows
dfCSummary <- bind_rows(dfCSummaryAll, dfCSummary, dfCSummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
dfCSummary %>% 
  kable("html", caption = "Summary Statistics for cell density (cells/mL) by Snow Color", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```


## Cell Volume (Site and Color)

```{r}
# Cell Volume 
dfCLSummary <- dfCL %>% 
  group_by(Color) %>% 
  summarise(
    n = n(),
    mean = comma(round(mean(Volume, na.rm = TRUE), 2)),
    median = comma(round(median(Volume, na.rm = TRUE), 2)),
    max = comma(round(max(Volume, na.rm = TRUE), 2)),
    min = comma(round(min(Volume, na.rm = TRUE), 2)),
    sd = comma(round(sd(Volume, na.rm = TRUE), 2)),
    .groups = "drop")

# Add "All" for each site 
dfCLSummaryAll <- dfCL %>% 
  summarise(
    Color = "all",
    n = n(),
    mean = comma(round(mean(Volume, na.rm = TRUE), 2)),
    median = comma(round(median(Volume, na.rm = TRUE), 2)),
    max = comma(round(max(Volume, na.rm = TRUE), 2)),
    min = comma(round(min(Volume, na.rm = TRUE), 2)),
    sd = comma(round(sd(Volume, na.rm = TRUE), 2)),
    .groups = "drop"
  )

# Filter Antarctic red + green
dfCLSummaryAPRG <- dfCL %>%
  filter(Color %in% c("red", "green")) %>%
  summarise(
    Color = "red+green",
    n = n(),
    mean = comma(round(mean(Volume, na.rm = TRUE), 2)),
    median = comma(round(median(Volume, na.rm = TRUE), 2)),
    max = comma(round(max(Volume, na.rm = TRUE), 2)),
    min = comma(round(min(Volume, na.rm = TRUE), 2)),
    sd = comma(round(sd(Volume, na.rm = TRUE), 2)),
    .groups = "drop")

# Combine and order rows
dfCLSummary <- bind_rows(dfCLSummaryAll, dfCLSummary, dfCLSummaryAPRG) %>% 
  arrange(match(Color, c("green", "red", "red+green", "clean", "all")))

# Format table with kableExtra
dfCLSummary %>% 
  kable("html", caption = "Summary Statistics for cell density (µm^3) by Snow Color", escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

# ANOVA/Kruskal Wallis Test
Before proceeding with an ANOVA test to assess differences in cell density/volume between colors, I am going to check to see if these variables follow the assumptions of normality and equal variance required to run an ANOVA. Cell density is below: 
```{r}
# Put your variables of interest into a vector
vars <- c("CellDensity")

# Loop over each variable
anovaAssumptionsResultsC <- map_dfr(vars, function(var) {
  
  # Fit ANOVA model
  formula <- as.formula(paste(var, "~ Color"))
  aov_model <- aov(formula, data = dfC) 
  
  # Extract assumption test results
  shapiro_p <- shapiro.test(residuals(aov_model))$p.value
  levene_p  <- leveneTest(formula, data = dfC)$`Pr(>F)`[1]
  
  tibble(
    Variable = var,
    shapiro_p = shapiro_p,
    levene_p = levene_p
  )
})

anovaAssumptionsResultsC
```

Cell volume is below:
```{r}
# Put your variables of interest into a vector
vars <- c("Volume")

# Loop over each variable
anovaAssumptionsResultsCL <- map_dfr(vars, function(var) {
  
  # Fit ANOVA model
  formula <- as.formula(paste(var, "~ Color"))
  aov_model <- aov(formula, data = dfCL) 
  
  # Extract assumption test results
  shapiro_p <- shapiro.test(residuals(aov_model))$p.value
  levene_p  <- leveneTest(formula, data = dfCL)$`Pr(>F)`[1]
  
  tibble(
    Variable = var,
    shapiro_p = shapiro_p,
    levene_p = levene_p
  )
})

anovaAssumptionsResultsCL
```

Neither meets assumptions of both normality and homogeneity of variance. Because of this, I am going to proceed with a Kruskal-Wallis test in lieu of an ANOVA. 

## Kruskal-Wallis Test {.tabset}
Below, I use the Kruskal-Wallis test to run a non-parametric comparison of cell density and cell volume across each location. Then, I follow up with a Dunn's post hoc test. I also look at differences in cell density/volume between different snow colors within each site. 

### Cell Density

```{r}
# Step 1: Kruskal-Wallis test for CellDensity
kruskal_modelAP <- kruskal.test(CellDensity ~ Color, data = dfC %>% filter(Location == "Antarctic Peninsula"))

kruskalResultsAP <- tibble(
  Variable = "CellDensity",
  kw_p = kruskal_modelAP$p.value
)

# Step 2: If significant, run Dunn's test
dunn_modelAP <- dunnTest(CellDensity ~ factor(Color), data = dfC %>% filter(Location == "Antarctic Peninsula"), method = "bonferroni")
dunn_tableAP <- dunn_modelAP$res

# Step 3: Prepare results for plotting (only if Dunn's test ran)
dunn_pairsAP <- dunn_tableAP %>%
    mutate(
      Variable = "CellDensity",
      significant = P.adj < 0.05,
      label = paste0("p=", round(P.adj, 3))
    )

```

#### Visualize Results
```{r}
dunnPlot_CellDensityAP <- ggplot(dunn_pairsAP, aes(x = Comparison, y = Variable, fill = P.adj)) +
    geom_tile() +
    geom_tile(data = filter(dunn_pairsAP, significant), color = "black", linewidth = 0.75) +
    geom_text(aes(label = label), size = 3) +
    scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.title = element_blank(),
      panel.grid = element_blank()
    )
  
dunnPlot_CellDensityAP
  
ggsave(dunnPlot_CellDensityAP, filename = "dunnPlot_CellDensityAP.png",
       width = 7, height = 6, bg = "white")
```

### Volume 

```{r}
# Step 1: Kruskal-Wallis test for Volume
kruskal_modelVAP <- kruskal.test(Volume ~ Color, data = dfCL %>% filter(Location == "Antarctic Peninsula"))

kruskalResultsVAP <- tibble(
  Variable = "Volume",
  kw_p = kruskal_modelVAP$p.value
)

# Step 2: If significant, run Dunn's test
dunn_modelVAP <- dunnTest(Volume ~ factor(Color), data = dfCL %>% filter(Location == "Antarctic Peninsula"), method = "bonferroni")
dunn_tableVAP <- dunn_modelVAP$res

# Step 3: Prepare results for plotting (only if Dunn's test ran)
dunn_pairsVAP <- dunn_tableVAP %>%
    mutate(
      Variable = "Volume",
      significant = P.adj < 0.05,
      label = paste0("p=", round(P.adj, 3))
    )

```

#### Visualize Results
```{r}
dunnPlot_VolumeVAP <- ggplot(dunn_pairsVAP, aes(x = Comparison, y = Variable, fill = P.adj)) +
    geom_tile() +
    geom_tile(data = filter(dunn_pairsVAP, significant), color = "black", linewidth = 0.75) +
    geom_text(aes(label = label), size = 3) +
    scale_fill_gradient(low = "tomato", high = "lightblue", name = "Adj. p-value") +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.title = element_blank(),
      panel.grid = element_blank()
    )
  
dunnPlot_VolumeVAP
  
ggsave(dunnPlot_VolumeVAP, filename = "dunnPlot_VolumeVAP.png",
       width = 7, height = 6, bg = "white")
```

# Box Plots {.tabset}

## Cell Density (Region & Color)

```{r}
boxplotDensityRegion <- dfC %>%
  ggplot(aes(x = Region, y = CellDensity, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.9, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Density (cells/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotDensityRegion

ggsave(boxplotDensityRegion, filename = "boxplotDensityRegion.png", width = 6, height = 4, bg = "white")
```

## Cell Density (Color)

```{r}
boxplotDensity <- dfC %>%
  ggplot(aes(x = Color, y = CellDensity, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.8, 
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Density (cells/mL)",
    fill = NULL,
    color = NULL
  ) +
  scale_y_sqrt() +
  theme_minimal() +
  plot_theme

boxplotDensity

ggsave(boxplotDensity, filename = "boxplotDensity.png", width = 6, height = 4, bg = "white")
```


## Cell Volume (Color)

```{r}
boxplotVolume <- dfCL %>%
  filter(Volume != 0) %>% 
  ggplot(aes(x = Color, y = Volume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.9,
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Volume (μm^3)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme

boxplotVolume

ggsave(boxplotVolume, filename = "boxplotVolume.png", width = 6, height = 4, bg = "white")
```

## Cell Volume (Region & Color)
```{r}
boxplotVolumeRegion <- dfCL %>%
  filter(Volume != 0) %>% 
  ggplot(aes(x = Region, y = Volume, fill = Color, color = Color)) +
  geom_boxplot(alpha = 0.8,
               width = 0.9,
               position = position_dodge2(width = 0.8, preserve = "single")) +
  scale_fill_manual(values = colorPal) +
  scale_color_manual(values = colorPal) +
  labs(
    x = NULL,
    y = "Cell Volume (μm^3)",
    fill = NULL,
    color = NULL
  ) +
  theme_minimal() +
  plot_theme

boxplotVolumeRegion

ggsave(boxplotVolumeRegion, filename = "boxplotVolumeRegion.png", width = 6, height = 4, bg = "white")
```

# Pairs Plots - Cell Density/Volume and Anions {.tabset}

```{r, warning = FALSE, message = FALSE}
corplotAP <- dfCLAnion %>% 
  ggpairs(  
  columns = c("CellDensity", "Volume", "Fo", "Cl", "NO2NO3", "PO43", "SO42"), 
  aes(color = Color, fill = Color, alpha = 0.8),
  diag = list(continuous = wrap("barDiag")),
  lower = list(continuous = wrap("points", na.rm = TRUE)),
  upper = list(continuous = wrap("cor", size = 3, na.rm = TRUE, method = "spearman"))
) + 
  theme_minimal() +
  pairsplot_theme + 
  scale_color_manual(values = colorPal) + 
  scale_fill_manual(values = colorPal) 
   

corplotAP

ggsave(corplotAP, filename = "corplotAP.png", width = 7, height = 7, bg = "white")
```


# Cell Density vs. Cell Length {.tabset}

## All sites 
I am going to compare cell density vs. cell length across sites to see whether any trends appear here. For this plot, I kept the samples where cell length was zero.

```{r}
scatterCDCL <- dfCL %>% 
  ggplot(aes(x = CellLength, y = CellDensity, color = Color)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", linewidth = 0.8) +
  scale_color_manual(values = colorPal) +
  scale_y_sqrt() +
  theme_minimal() + 
  plot_theme + 
  labs(x = "Cell Diameter (μm)",
       y = "Cell Density (cells/mL)",
       color = NULL)

scatterCDCL 

ggsave(scatterCDCL, filename = "scatterCDCL.png", width = 7, height = 4, bg = "white")
```


# Spearman's Correlation {.tabset}

### Antarctic Peninsula
```{r}
APcor_densityVolume <- dfCL %>%
  filter(Location == "Antarctic Peninsula") %>%
  { cor.test(~ CellDensity + Volume, data = . , method = "spearman") }

APcor_densityVolume
```

```{r}
APcor_densityVolumeR <- dfCL %>%
  filter(Location == "Antarctic Peninsula") %>%
  filter(Color == "red") %>% 
  { cor.test(~ CellDensity + Volume, data = . , method = "spearman") }

APcor_densityVolumeR
```

```{r}
APcor_densityVolumeG <- dfCL %>%
  filter(Location == "Antarctic Peninsula") %>%
  filter(Color == "green") %>% 
  { cor.test(~ CellDensity + Volume, data = . , method = "spearman") }

APcor_densityVolumeG
```


# Heat Maps - Cell Density/Volume & Anions {.tabset}

## Cell Density

```{r}
# Variable groups
algae_vars <- c("CellDensity")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Function to extract correlations
get_correlation_data <- function(data, location_name, algae_name) {

  cor_matrix <- rcorr(
    as.matrix(data[, c(algae_vars, anion_vars)]),
    type = "spearman"
  )

  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  pairs <- expand.grid(
    AlgaeVar = algae_vars,
    Anion = anion_vars,
    stringsAsFactors = FALSE
  )

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[AlgaeVar, Anion],
      p_value = p_vals[AlgaeVar, Anion],
      significant = p_value < 0.05,
      label = paste0(
        round(Spearman_r, 2),
        "\n(p=", round(p_value, 3), ")"
      ),
      Location = location_name,
      SnowColor = algae_name
    ) %>%
    ungroup()

  return(results)
}

# Filtered datasets
# Red + Green only
df_rg <- dfCAnion %>%
  filter(
    Location == "Antarctic Peninsula",
    Color %in% c("red", "green")
  ) %>%
  select(Location, Color, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# ALL samples (includes clean)
df_all <- dfCAnion %>%
  filter(Location == "Antarctic Peninsula") %>%
  select(Location, Color, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Correlations
# Red + Green (separate rows)
cor_rg <- df_rg %>%
  group_split(Location, Color) %>%
  map_dfr(~get_correlation_data(
    .x,
    unique(.x$Location),
    unique(.x$Color)
  ))

# Algae pooled (red + green)
cor_algae <- df_rg %>%
  group_by(Location) %>%
  group_split() %>%
  map_dfr(~get_correlation_data(
    .x,
    unique(.x$Location),
    "algae"
  ))

# All samples pooled (red + green + clean)
cor_all <- df_all %>%
  group_by(Location) %>%
  group_split() %>%
  map_dfr(~get_correlation_data(
    .x,
    unique(.x$Location),
    "all"
  ))

# Combine results
cor_results_full <- bind_rows(
  cor_rg,
  cor_algae,
  cor_all
)

# Factor ordering for clean y-axis
cor_results_full$SnowColor <- factor(
  cor_results_full$SnowColor,
  levels = c("red", "green", "algae", "all")
)

cor_results_full$Anion <- factor(
  cor_results_full$Anion,
  levels = anion_vars
)

# Plot
spearmanPlotAPRG <- ggplot(
  cor_results_full,
  aes(
    x = Anion,
    y = interaction(Location, SnowColor, sep = " - "),
    fill = Spearman_r
  )
) +
  geom_tile() +
  geom_tile(
    data = filter(cor_results_full, significant),
    color = "black",
    linewidth = 0.75
  ) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0,
    name = "Spearman r"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    axis.text.y = element_text(size = 8)
  )

spearmanPlotAPRG

# Save
ggsave(
  spearmanPlotAPRG,
  filename = "spearmanPlotAPRG.png",
  width = 9,
  height = 7,
  bg = "white"
)
```

## Cell Volume

```{r}

# Variable groups
algae_vars <- c("Volume")
anion_vars <- c("Cl", "NO2NO3", "SO42", "PO43", "Fo")

# Function to extract correlations
get_correlation_data <- function(data, location_name, algae_name) {

  cor_matrix <- rcorr(
    as.matrix(data[, c(algae_vars, anion_vars)]),
    type = "spearman"
  )

  r_vals <- cor_matrix$r
  p_vals <- cor_matrix$P

  pairs <- expand.grid(
    AlgaeVar = algae_vars,
    Anion = anion_vars,
    stringsAsFactors = FALSE
  )

  results <- pairs %>%
    rowwise() %>%
    mutate(
      Spearman_r = r_vals[AlgaeVar, Anion],
      p_value = p_vals[AlgaeVar, Anion],
      significant = p_value < 0.05,
      label = paste0(
        round(Spearman_r, 2),
        "\n(p=", round(p_value, 3), ")"
      ),
      Location = location_name,
      SnowColor = algae_name
    ) %>%
    ungroup()

  return(results)
}

# Filtered datasets
# Red + Green only
df_rg <- dfCLAnion %>%
  filter(
    Location == "Antarctic Peninsula",
    Color %in% c("red", "green")
  ) %>%
  select(Location, Color, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# ALL samples (includes clean)
df_all <- dfCLAnion %>%
  filter(Location == "Antarctic Peninsula") %>%
  select(Location, Color, all_of(c(algae_vars, anion_vars))) %>%
  na.omit()

# Correlations
# Red + Green (separate rows)
cor_rg <- df_rg %>%
  group_split(Location, Color) %>%
  map_dfr(~get_correlation_data(
    .x,
    unique(.x$Location),
    unique(.x$Color)
  ))

# Algae pooled (red + green)
cor_algae <- df_rg %>%
  group_by(Location) %>%
  group_split() %>%
  map_dfr(~get_correlation_data(
    .x,
    unique(.x$Location),
    "algae"
  ))

# All samples pooled (red + green + clean)
cor_all <- df_all %>%
  group_by(Location) %>%
  group_split() %>%
  map_dfr(~get_correlation_data(
    .x,
    unique(.x$Location),
    "all"
  ))

# Combine results
cor_results_full <- bind_rows(
  cor_rg,
  cor_algae,
  cor_all
)

# Factor ordering for clean y-axis
cor_results_full$SnowColor <- factor(
  cor_results_full$SnowColor,
  levels = c("red", "green", "algae", "all")
)

cor_results_full$Anion <- factor(
  cor_results_full$Anion,
  levels = anion_vars
)

# Plot
spearmanPlotVAPRG <- ggplot(
  cor_results_full,
  aes(
    x = Anion,
    y = interaction(Location, SnowColor, sep = " - "),
    fill = Spearman_r
  )
) +
  geom_tile() +
  geom_tile(
    data = filter(cor_results_full, significant),
    color = "black",
    linewidth = 0.75
  ) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0,
    name = "Spearman r"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    axis.text.y = element_text(size = 8)
  )

spearmanPlotVAPRG

# Save
ggsave(
  spearmanPlotVAPRG,
  filename = "spearmanPlotVAPRG.png",
  width = 9,
  height = 7,
  bg = "white"
)
```






